### 2016-10-28 nips2016/nips2016.tex
nips v1 from notes (2016-04-29)

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage{nips2016}
%\usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Convolutional Neural Networks for Graphs}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Michaël Defferrard \\ %\thanks{further info} \\
  %Department of Electrical Engineering \\
  %École Polytechnique Fédérale de Lausanne (EPFL) \\
  EPFL \\
  Lausanne, Switzerland \\
  \texttt{michael.defferrard@epfl.ch} \\
  \And % \AND
  Xavier Bresson \\
  EPFL \\
  Lausanne, Switzerland \\
  \texttt{xavier.bresson@epfl.ch} \\
  \And
  Pierre Vandergheynst \\
  EPFL \\
  Lausanne, Switzerland \\
  \texttt{pierre.vandergheynst@epfl.ch} \\
}

\usepackage[acronym]{glossaries}
\newacronym{SVD}{SVD}{Singular Value Decomposition}
\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
\newacronym{PSD}{PSD}{positive semidefinite}

\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\spn}{span}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\W}{\mathcal{W}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Xh}{\hat{X}}
\newcommand{\Yh}{\hat{Y}}
\newcommand{\st}{\ \text{s.t.} \,}
\newcommand{\norm}[1]{\left\| #1 \right\|}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
This paper introduces Graph Neural Network (GNN), a novel framework to process
data on irregular, i.e.~non-Euclidean domains modelized as graphs. Even recent
developments in Deep Learning who use graph embeddings don't make use of the
graph structure per se.
\end{abstract}

\section{Introduction}\label{introduction}

Large networks allow to learn complex relationships. Although they
overfit easily if the training set is small because there is a lot of
parameters to learn.

Convolutional neural networks (CNN) offer a great reduction of
parameters on regular Euclidean domain. This work proposes to generalize
it to the discrete model of manifolds, graphs. Such an approach was
proposed by \citep{bruna_spectral_2013, henaff_deep_2015}, we will
refine it with proper graph signal processing tools.

This paper introduces a framework to use CNN with signals defined on
weighted graphs. Our work leverages spectral methods, which are much
faster than a spatial approach would be, in the same sense as doing
convolutions as multiplication in the Fourier domain (with the help of
the FFT) is much faster than multiplying patches on an image.

Another option to do clustering or regression on data graphs is to first
embed the graph and then use the embedding as additional features.

\section{Related work}\label{related-work}

Comparison with fully connected NN (Reuters used in
\citep{srivastava_dropout_2014})

\begin{itemize}
\item
  + less parameters via convolution
\item
  ? accuracy
\end{itemize}

Comparison with first generation graph CNN \citep{henaff_deep_2015}.

\begin{itemize}
\item
  + better accuracy
\item
  = same number of parameters
\item
  - graph construction vs supervised estimation
\item
  Theoretical:

  \begin{itemize}
  \item
    + proper spectral graph signal processing tools
  \item
    + kNN graph (sparse weight matrix) vs fully connected graph. SVD
    complexity for Fourier basis is O(n\^{}3), we avoid it.
  \item
    + METIS vs naive agglomeration
  \item
    + Chebyshev polynomials vs splines to approximate filters Filtering
    from \(O(n^2)\) to \(O(|E| K)\) where \(K\) is the polynomial order
    approximately \(O(n)\) if the weight matrix is sparse, i.e.~grows
    linearly with \(N\) This addresses the first limitation mentioned in
    \citep{henaff_deep_2015}.
  \end{itemize}
\end{itemize}

\section{Model}\label{model}

\subsection{Spectral Graph Theory}\label{spectral-graph-theory}

A graph \(\G = (\V, \E, \W)\) is defined by a set \(V\) of \(|V| = M\)
nodes and a set \(E\) of weighted edges. The connectivity of the graph
is captured by the adjacency matrix \(W \in \R^{M \times M}\) which
entry \(W_{i,j}\) denotes the weight of the edge \((v_i, v_j) \in E\)
which connects the vertex \(v_i \in V\) to \(v_j \in V\). It is set to
\(0\) if the vertices are not connected, i.e. \((v_i, v_j) \notin E\).
Assuming an undirected graph, \(W\) is a symmetric matrix. A graph
signal is any signal \(x \in \R^M\) defined on the vertices of \(G\).

The combinatorial graph Laplacian is defined as
\[ \L^c := D - W \in \R^{M \times M}, \] where \(D \in \R^{M \times M}\)
is the diagonal degree matrix defined as \(D_{i,i} := \sum_j W_{i,j}\).
Note that it is a difference operator such that
\[ (\L^c x)_i = \sum_j W_{i,j} (x_i - x_j). \] The normalized graph
Laplacian is defined as
\[ \L^n := I_M - D^{-1/2} W D^{-1/2} \in \R^{M \times M}, \] where
\(I_M \in \R^{M \times M}\) is the identity matrix. Finally, the
random-walk graph Laplacian is defined as
\[ \L^{rw} := I_M - D^{-1} W \in \R^{M \times M}. \] Note that this work
is independent of the chosen graph Laplacian \(\L\), which can be any of
\(\L^c\), \(\L^n\) or \(\L^{rw}\).

In analogy to the real line Fourier Transform, a Fourier basis is given
by the eigenvectors of the Laplacian
\[ \L u_i = \lambda_i u_i \st \norm{u_i}_2=1,
\ u_i \in \R^M, \ i = 0, \ldots, M-1, \] with their associated
eigenvalues \(\lambda_i\). Assuming the graph is connected, we may order
the vector \(\lambda := [\lambda_0, \ldots, \lambda_{M-1}]^T \in \R^M\)
of eigenvalues such that
\[ 0 = \lambda_{min} = \lambda_0 < \lambda_1 \leq \ldots \leq
\lambda_{M-1} = \lambda_{max}. \] As the Laplacian is a real symmetric
and \gls{PSD} matrix, the eigenvalues are real and positive, and the
eigenvectors are orthonormal. The Laplacian is indeed diagonalized by
the Fourier basis \(U := [u_0, \ldots, u_{M-1}] \in \R^{M \times M}\)
such that
\begin{equation} \L = U \Lambda U^T, \label{eq:lap_diag}\end{equation}
where \(\Lambda := \diag(\lambda) \in \R^{M \times M}\) denotes a
diagonal matrix of eigenvalues and \(U^T\) is the matrix transpose of
\(U\). See \citep{chung_spectral_1997} for details on spectral graph
theory.

The Graph Fourier Transform \(\hat{x} \in \R^M\) of any graph signal
\(x\) is given by \[ \hat{x} = U^T x =
[\langle u_0, x \rangle, \ldots, \langle u_{M-1}, x \rangle]^T, \] where
\(\langle \cdot , \cdot \rangle\) denotes the standard \(\ell_2\) inner
product \citep{shuman_emerging_2013}. The inverse transform is then
given by \[
x = U \hat{x}. \]

It follows that any signal \(\hat{x}\) can be filtered in the spectral
domain by \[ \hat{y} = g_\theta(\Lambda) \hat{x}, \] where the operator
\(g_\theta(\Lambda) \in \R^{M \times M}\), a matrix function, yields a
diagonal matrix of \(M\) Fourier coefficients parametrized by
\(\theta\). The filtering operation can equivalently take place in the
vertex domain as \[ y = U g_\theta(\Lambda) U^T x = g_\theta(\L) x \]
where the operator
\(g_\theta(\L) = g_\theta(U \Lambda U^T) = U g_\theta(\Lambda) U^T \in \R^{M \times M}\)
is akin to a convolution with the filter \(Ug_\theta(\Lambda)\) in the
vertex domain. Note that the filtering operation is defined as a
multiplication in the spectral domain because a convolution cannot be
defined in the vertex domain \citep{shuman_emerging_2013}.

\subsection{Graph Filter Learning}\label{graph-filter-learning}

Given a graph \(G\), a set
\(X = [x_0, \ldots, x_{N-1}] \in \R^{M \times N}\) of \(N\) source graph
signals of dimensionality \(M\) and their associated set
\(Y = [y_0, \ldots, y_{N-1}] \in \R^{M \times N}\) of target signals, we
want to learn the parameters \(\theta \in \R^M\) of a graph filter
\(g_\theta(\Lambda) := \diag(\theta)\) such as to minimize the convex
mean square error \begin{equation} L =
\frac{1}{N} \sum_{i=0}^{N-1} \norm{ g_\theta(\L) x_i - y_i }_2^2 =
\frac{1}{N} \norm{ U \diag(\theta) U^T X - Y }_F^2, \label{eq:loss}\end{equation}
where \(\norm{\cdot}_2^2\) denotes the squared \(\ell_2\) norm and
\(\norm{\cdot}_F^2\) the squared Frobenius norm.

Note that, while being the most flexible definition of
\(g_\theta(\Lambda)\), the independence of the Fourier coefficients
\(\theta\) from the eigenvalues \(\Lambda\) does omit all information
about frequencies. As we know that the lower frequencies are more
important for clustering {[}ref NCut{]}, we could have designed a
parametric filter with less parameters.

Rewriting eq.~\ref{eq:loss} in the spectral domain while decomposing it
w.r.t. the scalar coefficients \(\theta_i\) gives \[ L =
\frac{1}{N} \norm{ \diag(\theta) U^T X - U^T Y }_F^2 =
\frac{1}{N} \sum_{i=0}^{M-1} \norm{\theta_i\Xh_{i,\cdot}-\Yh_{i,\cdot}}_2^2, \]
where \(\Xh = U^TX \in \R^{M \times N}\) and
\(\Yh = U^TY \in \R^{M \times N}\) are the spectral representations of
the signals \(X\) and \(Y\). The gradient for each coefficient is then
given by \[ \nabla_{\theta_i} L =
\frac{2}{N} ( \theta_i \Xh_{i,\cdot} - \Yh_{i,\cdot} ) \Xh^T_{\cdot,i} \]
and can be rewritten in a vectorized form as
\begin{equation} \nabla_{\theta} L =
\frac{2}{N} \diag \left( (\diag(\theta) \Xh - \Yh) \Xh^T \right) =
\frac{2}{N} \left( ( \theta 1_N^T \odot \Xh - \Yh ) \odot \Xh \right) 1_N,
\label{eq:gradient}\end{equation} where \(1_N\) denotes a unit vector of
length \(N\) and \(\odot\) the element-wise Hadamard product. The second
form avoids the computation of the useless off-diagonal elements.

A closed-form solution is given by the optimality condition
\(\nabla_{\theta}L=0\) such that
\begin{equation} \theta^* = \argmin_\theta L =
(\Xh \odot \Yh) 1_N \oslash (\Xh \odot \Xh) 1_N, \label{eq:direct}\end{equation}
where \(\oslash\) denotes an element-wise division. Note that this
method is impractical for large \(M\) and \(N\) (sufficiently large for
\(X\) and \(Y\) to not fit in memory). A \gls{SGD} based on
eq.~\ref{eq:gradient} is then necessary.

There are however two major computational drawbacks to this optimization
process: 1) the Fourier transform \(\Xh=U^TX\) of a set of signals costs
\(O(M^2N)\) operations and 2) the eigenvalue decomposition
\(\L = U \Lambda U^T\) costs \(O(M^3)\). The total cost of filtering is
thus \(O(M^2 \max(M,N))\), a problem already stated in
\citep{henaff_deep_2015}. We propose to overcome these computational
issues using efficient numerical approximations such as the Chebyshev
polynomials or the Lanczos algorithm.

\subsubsection{Chebyshev
Parametrization}\label{chebyshev-parametrization}

The main idea, to avoid the Fourier basis, is twofold: 1) parametrize
the filter, in the spectral domain, as a truncated Chebyshev expansion
and 2) recursively compute the Chebyshev polynomials from the Laplacian.
This approximate method for spectral graph filtering was first proposed
in \citep{hammond_wavelets_2011} for a fast wavelet transform on graphs.

Recall that the Chebyshev polynomial \(T_k(x)\) of order \(k\) may be
generated by the stable recurrence relation
\(T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)\) with \(T_0 = 1\) and
\(T_1 = x\). These polynomials form an orthogonal basis for
\(L^2([-1,1], dy / \sqrt{1-y^2})\), the Hilbert space of square
integrable functions with respect to the measure \(dy/\sqrt{1-y^2}\).

The graph filter \(g_\theta(\Lambda)\) can thus be constructed from the
truncated expansion
\[ g_\theta(\Lambda) := \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\Lambda}) \]
of polynomial order \(K-1\), where the parameter \(\theta \in \R^K\),
\(K \ll M\), is a vector of Chebyshev coefficients and
\(T_k(\tilde{\Lambda}) \in \R^{M \times M}\) is the Chebyshev polynomial
of order \(k\) evaluated at
\(\tilde{\Lambda} := 2 \Lambda / \lambda_{max} - I_M \in \R^{M \times M}\),
a diagonal matrix of scaled eigenvalues (so that they lie in
\([-1,1]\)). While reducing the number of parameters from \(M\) to
\(K\), this parametrization enforces smoothness in the spectral domain,
which translates to localization in the vertex domain. It can indeed be
shown that \((\L^k)_{i,j}=0\) if the shortest-path between vertices
\(v_i\) and \(v_j\) is longer than \(k\) edges
\citep{hammond_wavelets_2011}, which limits the influence of a
\(K^\text{th}\) order filter to \(K\) hopes. This is often a desired
property, e.g.~to learn local features for classification.

To avoid the Fourier basis \(U\), we express the polynomials as
functions of the Laplacian \(\L\) instead of its eigenvalues \(\Lambda\)
using eq.~\ref{eq:lap_diag}, such that the filtering operator in the
vertex domain is given by
\begin{equation} g_\theta(\L) = U g_\theta(\Lambda) U^T =
\sum_{k=0}^{K-1} U \theta_k T_k(\tilde{\Lambda}) U^T =
\sum_{k=0}^{K-1} \theta_k T_k(\tilde{\L}), \label{eq:chebyshev}\end{equation}
where \(T_k(\tilde{\L}) \in \R^{M \times M}\) is the Chebyshev
polynomial of order \(k\) evaluated at the scaled Laplacian
\(\tilde{\L} := 2 \L / \lambda_{max} - I_M\). Note that the spectrum of
the normalized Laplacian is bounded by \(2\)
\citep{chung_spectral_1997}, such that the scaling can simply be
\(\tilde{\L} = \L - I_M\), tolerating some imprecision in the
approximation due to the loss of a fraction \(2-\lambda_{max}\) of the
domain.

Inserting eq.~\ref{eq:chebyshev} into eq.~\ref{eq:loss} we obtain
\begin{equation} L =
\frac{1}{N} \norm{ \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\L}) X - Y }_F^2 =
\frac{1}{N} \norm{ \bar{X} \theta - \bar{y} }_2^2, \label{eq:loss_c}\end{equation}
where \(\bar{y} \in \R^{MN}\) is the vectorized matrix \(Y\) and the
\(k^\text{th}\) column of \(\bar{X} \in \R^{MN \times K}\) is the
vectorized matrix
\(\tilde{X}_k := T_k(\tilde{\L}) X \in \R^{M \times N}\). The gradient
is then given by \begin{equation} \nabla_\theta L =
\frac{2}{N} \bar{X}^T (\bar{X} \theta - \bar{y}). \label{eq:gradient_c}\end{equation}

While the system \(\bar{X} \theta - \bar{y}\) is largely over-determined
as \(K \ll MN\), a closed-form solution of eq.~\ref{eq:loss_c} is given
by the optimality condition \(\nabla_\theta L = 0\) so that
\begin{equation} \theta^* = \argmin_\theta L = \bar{X}^+ \bar{y}, \label{eq:direct_c}\end{equation}
where \(\bar{X}^+ = (\bar{X}^T\bar{X})^{-1} \bar{X}^T\) is the
pseudo-inverse of \(\bar{X}\). The computation of the inverse is fast
and stable for small \(K\). Alternative approximate solution methods for
larger order \(K\) include the computation of the pseudo-inverse with
\gls{SVD}, the use of a least-square solver or a gradient descent scheme
based on eq.~\ref{eq:gradient_c}.

Using the recurrence
\[ \tilde{X}_k = 2\tilde{\L} \tilde{X}_{k-1} - \tilde{X}_{k-2} \] with
\(\tilde{X}_0 = X\) and \(\tilde{X}_1 = \tilde{\L} X\), the computation
of \(\bar{X}\) given \(X\) costs \(O(K|E|N) = O(KMN)\) operations if the
number of edges \(|E|\) is proportional to the number of nodes \(M\),
e.g.~for kNN graphs. As the cost of the product \(\bar{X}\theta\) is
similar, the entire filtering operation has a computational cost of
\(O(KMN)\) whereas the straightforward implementation using the Fourier
basis has a cost of \(O(M^2\max(M,N)\). When applying multiple filters
to the same set of signals, as is the case with \gls{SGD}, one may save
computations by storing and querying \(\bar{X}\) instead of \(X\), at
the expense of \(K\) times the memory.

\subsubsection{Lanczos Parametrization}\label{lanczos-parametrization}

Another applicable parametrization is based on the Lanczos algorithm, an
adaptation of the power iteration to find the largest or smallest
eigenvalues and corresponding eigenvectors of a linear system. It was
first introduced for fast graph filtering in
\citep{susnjara_accelerated_2015}.

The algorithm, described in
\citep{gallopoulos_efficient_1992, susnjara_accelerated_2015},
constructs an orthonormal basis
\(V = [v_0, \ldots, v_{K-1}] \in \R^{M \times K}\) of the Krylov
subspace \(\mathcal{K}_K(\L,x) = \spn\{ x, \L x, \ldots, \L^{K-1} x \}\)
and a tri-diagonal matrix \(H = V^T \L V \in \R^{K \times K}\) with a
computational cost of \(O(K |E|)\). Note that for large
\(K \gtrapprox 30\), the original iterative algorithm may loose the
basis orthogonality such that a necessary orthogonalization step will
increase the complexity \citep{susnjara_accelerated_2015}. Filtering the
signal \(x\) with \(g_\theta(\L)\) can then be approximated by an order
\(K-1\) polynomial as \begin{equation} y = g_\theta(\L) x \approx
V g_\theta(H) V^T x = V Q g_\theta(\Sigma) Q^T V^T x \label{eq:lanczos}\end{equation}
where \(Q \Sigma Q^T = H\) is the eigendecomposition of \(H\). There
exist fast methods for the eigendecomposition of symmetric tri-diagonal
matrices {[}{]}. As for the Chebyshev approximation, this construction
enforces smoothness in the spectral domain.

Inserting eq.~\ref{eq:lanczos} with a parametrized filter
\(g_\theta(\Sigma) := \diag(\theta)\), \(\theta \in \R^K\), \(K \ll M\),
into eq.~\ref{eq:loss} gives \[ L = \frac{1}{N} \sum_{n=0}^{N-1}
\norm{V_n Q_n \diag(\theta) Q_n^T V_n^T x_n - y_n}_2^2 =
\frac{1}{N} \norm{\diag(\theta) \hat{X} - \hat{Y}}_F^2 \] where
\(\hat{X} = [\hat{x}_0, \ldots, \hat{x}_{N-1}] \in \R^{K \times N}\),
\(\hat{Y} = [\hat{y}_0, \ldots, \hat{y}_{N-1}] \in \R^{K \times N}\) and
\(\hat{x}_n = Q_n^T V_n^T x_n \in \R^K\),
\(\hat{y}_n = Q_n^T V_n^T y_n \in \R^K\) are approximate representations
of the signals \(x_n\), \(y_n\) in the orthonormal basis
\(V_n Q_n \in \R^{M \times K}\) of \(\mathcal{K}_K(\L,x_n)\). Similarly
to eq.~\ref{eq:loss}, the gradient is given by eq.~\ref{eq:gradient} and
a closed-form solution by eq.~\ref{eq:direct}.

The expression eq.~\ref{eq:lanczos} can be simplified by setting
\(v_0 := x / \norm{x}_2\) (the first basis vector can be set to an
arbitrary unit length vector) such that \(V^T x = \norm{x}_2 e_1\) and
\[ V g_\theta(H) V^T x =
\norm{x}_2 V Q \diag(\theta) Q^T e_1 =
\norm{x}_2 V Q \diag(q) \theta \] where \(e_1 \in \R^K\) denotes the
first unit vector and \(q = Q^T e_1 \in \R^K\) is the first row of
\(Q\). Inserting into eq.~\ref{eq:loss} gives \[ L =
\frac{1}{N} \sum_{n=0}^{N-1} \norm{\tilde{X}_n \theta - y_n}_2^2 =
\frac{1}{N} \sum_{k=0}^{K-1} \norm{\bar{X}_{\cdot,k} \theta_k - \bar{y}}_2^2 =
\frac{1}{N} \norm{\bar{X} \theta - \bar{y} }_2^2 \] where
\(\bar{y} \in \R^{NM}\) is the vectorized matrix \(Y\) and
\(\bar{X} := [\tilde{X}_0, \ldots, \tilde{X}_{N-1}]^T \in \R^{NM \times K}\)
is a stack of \(N\) matrices
\(\tilde{X}_n := \norm{x_n}_2 V_n Q_n \diag(q_n) \in \R^{M \times K}\)
where \(V_n\), \(Q_n\) and \(q_n\) are derived from \(x_n\). The second
form (in terms of independent coefficients \(\theta_k\)) is valid
because \(\bar{X}\) is orthogonal, i.e \(\bar{X}^T \bar{X}\) is a
diagonal matrix, so that the solution eq.~\ref{eq:direct_c} can be
written as \begin{equation} \theta^* = \argmin_\theta L =
\underbrace{\sum_{n=0}^{N-1} \Big( \norm{x_n}_2 \diag(q_n) \Big)^{-2}}
_{(\bar{X}^T\bar{X})^{-1}} \bar{X}^T \bar{y} \label{eq:direct_l}\end{equation}
where
\(\bar{X}^T \bar{y} = [\langle \bar{X}_{\cdot,0}, \bar{y} \rangle, \ldots, \langle \bar{X}_{\cdot,K-1}, \bar{y} \rangle]^T \in \R^K\)
is the projection of \(\bar{y}\) onto the \(K\) basis vectors
\(\bar{X}_{\cdot,k} \in \R^{MN}\). While this closed-form evaluation is
fast and stable, a gradient descent scheme can be used with
eq.~\ref{eq:gradient_c}.

While the time and space complexities are similar, this method has two
advantages over the Chebyshev approximation: 1) it does not require the
normalization of the Laplacian spectrum (thus the estimation of
\(\lambda_{max}\)) and 2) the optimization is easier as the parameters
\(\theta_k\) are independent of each other thanks to the orthogonal
basis. By easier we mean that 1) the closed-form solution
eq.~\ref{eq:direct_l} does not involve the inversion of a large matrix
and 2) \gls{SGD} converges to a better solution with less iterations.
See \citep{susnjara_accelerated_2015} for a discussion of the
approximation quality and a comparison with the Chebyshev approximation.

\subsection{Graph coarsening}\label{graph-coarsening}

The graph partitioning problem has been widely studied. Applications
include VLSI design, load balancing for parallel computations, network
analysis, and optimal scheduling. The goal is to partition the vertices
of a graph into a certain number of disjoint sets of approximately the
same size so that a cut metric is minimized. This problem is NP-complete
even for several restricted classes of graphs, and there is no constant
factor approximation algorithm for general graphs {[}Bui and Jones
1992{]}. While notable developments in exact algorithms and heuristics
have been done, only the introduction of the general-purpose multilevel
methods during the 1990s has provided a breakthrough in efficiency and
quality {[}Safro 2009 2012{]}. Graph coarsening, where the problem
instance is gradually mapped to smaller ones to reduce the original
complexity, i.e., the graph underlying the problem is compressed, is the
first step of these multilevel methods.

Existing multilevel algorithms for combinatorial optimization problems
(such as k-partitioning, linear ordering, clustering, and segmentation)
can be divided into two classes: contraction-based schemes and algebraic
multigrid (AMG)-inspired schemes.

Alternative standard algorithms for graph coarsening include the Kron
reduction {[}{]}, Label Propagation {[}{]}, spectral clustering {[}{]}
or a sampling {[}Nath uncertainty{]}. A standard approach for graph
coarsening is the maximal match, as used by METIS for graph partitioning
{[}?{]}.

The assumption behind METIS is that a (near) optimal partitioning on a
coarser graph implies a good partitioning in the finer graph. In
general, it only holds true when the degree of nodes in the graph is
bounded by a constant {[}Karypis 95{]}. Some real-life graphs,
e.g.~social networks or WWW, are right-skewed, i.e.~there exists hub
vertices which have very large degrees. These graphs follow a power-law
degree distribution.

\section{Applications}\label{applications}

\subsection{Regression}\label{regression}

Solving a linear regression problem on graph signals is akin to minimize
eq.~\ref{eq:loss} where \(Y\) is the dependent variable and \(X\) the
independent observations.

\subsection{Representation learning}\label{representation-learning}

Representation learning is the problem of finding an appropriate (to
some criteria) representation \(y_i\) of a sample \(x_i\) defined as
\[ \min_{A,Y} \norm{ AX - Y }_F^2 + \tau P(Y) \] where
\(A \in \R^{M \times M}\) is some operator and \(P(Y)\) is some prior on
\(Y\). Take \(P(Z)=\norm{Z}_F^2\) and you get a Thikhonov prior. Take
\(P(Y) = \norm{Y}_1\) and you get sparse coding (with dictionary
learning).

\subsection{Classification}\label{classification}

There are dependent vectors
\(X = \{x_i\}_{i=0}^{N-1} \in \R^{M \times N}\) we want to predict from
observations \(Y = \{y_i\}_{i=0}^{N-1} \in \R^{M \times N}\).

In the classification case, \(y_i \in \N\) is a discrete number which
represents a class, e.g. \(y_i=-1\) indicates that sample \(x_i\)
belongs to class 1 and \(y_i=+1\) to class 2.

\subsubsection{Representation learning}\label{representation-learning-1}

Three ways to reduce the number of parameters:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Graph coarsening
\item
  No weights: loss of spatiality (convolutional only)
\item
  Non-linear activation function
\end{enumerate}

But \(\theta\) and \(w\) are redundant, so
\[ L = \norm{\overline{\bar{X} \theta}^T 1_{FM} - y}_2^2 \]
\[ \bar{X} \in \R^{NM \times K} \] \[ \theta \in \R^{K \times F} \]

\[ \nabla_\theta L = \bar{X}^T (\bar{X} \theta - z) \]
\[ \nabla_\theta L = (\overline{\bar{X} \theta}^T 1_{FM} - y) \]

\[ \bar{X} 1_{FM} (\overline{\bar{X} \theta}^T 1_{FM} - y) \]
\[ (\bar{X} 1_{FM} \overline{\bar{X}}^T 1_{FM})^{-1} \bar{X} 1_{FM} y \]

\subsubsection{Linear classification}\label{linear-classification}

Empirical risk minimization (statistical learning).

The \(\ell_2\) regularized least-square model, also known as ridge
regression, is defined as \begin{equation} L =
\frac{1}{N} \sum_{i=0}^{N-1} |f_w(x_i) - y_i|^2 + \tau \norm{f}_\H^2,
\label{eq:rls}\end{equation} where \(f_w \in \H\) is some classification
function parametrized by \(w\) and \(\H\) is an Hilbert space.

Introducing the linear classifier \(f(x) = w^T x\) into eq.~\ref{eq:rls}
gives \[ L = \frac{1}{N} \norm{X^T w - y}_2^2  + \tau \norm{w}_2^2, \]
where \(w \in \R^M\) are the weights to learn. Note that there is no
bias because the data can be centered around zero (we know the lower and
upper bounds on luminosity values).

Gradient
\begin{equation} \nabla_w L = \frac{2}{N} X (X^T w - y) + 2 \tau w \label{eq:gradient_rls}\end{equation}

Optimality condition \(\nabla L = 0\)
\begin{equation} w^* = \argmin_w \nabla L =
\left(X X^T + \tau N I_M \right)^{-1} X y \label{eq:sol_rls}\end{equation}

\subsubsection{Feature learning}\label{feature-learning}

Introducing graph filter learning as a scheme for feature learning gives
the objective \[ L = \frac{1}{N} \sum_{i=0}^{N-1}
\Big| f_w \Big( A_\theta(\L) x_i \Big) - y_i \Big|^2 +
\tau_R \norm{f_w}_\H^2, \] where the filter bank
\(A_\theta(\L) = \bigcup_{i=0}^{F-1} g_{\theta_i}(\L) \in \R^{FM \times M}\)
is a block-diagonal matrix composed of \(F\) filters
\(g_{\theta_i}(\L)\) which extract different features.

Using a linear classifier and the Lanczos parametrization introduced in
\ref{lanczos-parametrization} gives \begin{equation} L =
\frac{1}{N} \norm{\overline{\bar{X} \theta}^T w - y}_2^2 +
\tau_R \norm{w}_2^2, \label{eq:rls_gfl}\end{equation} where
\(w \in \R^{FM}\) are the weights to learn,
\(\bar{X} \in \R^{NM \times K}\) is the Lanczos basis described in
\ref{lanczos-parametrization}, \(\theta \in \R^{K \times F}\) are the
\(K\) coefficients of the \(F\) filters and
\(\overline{\bar{X} \theta} \in \R^{FM \times N}\) is a rearrangement of
the matrix \(\bar{X} \theta \in \R^{NM \times F}\). Note that there is
now \(F\) times more weigts to learn, i.e. \(w \in \R^{FM}\).

\(\bar{X} \theta \in \R^{NM \times F}\)

\[ \frac{2}{N} \bar{X} w (\overline{\bar{X} \theta}^T w - y) \]

\[ (\bar{X} w \overline{\bar{X}}^T w)^{-1} \bar{X}wy \]

\subsubsection{Splitting scheme}\label{splitting-scheme}

The splitting scheme \[ L =
\frac{1}{N} \norm{Z^T w - y}_2^2 +
\frac{\tau_F}{N} \norm{\overline{\bar{X} \theta} - Z}_F^2 +
\tau_R \norm{w}_2^2 \] may be used to minimize the non-convex objective
eq.~\ref{eq:rls_gfl} for \(\theta\) and \(w\), where
\(Z = [z_0, \ldots, z_{N-1}] \in \R^{FM \times N}\) are the signals in
feature space. Solving the overall non-convex problem requires to
iteratively solve the following three convex sub-problems:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The gradient and solution to \(\min_\theta L\) are given by
  eq.~\ref{eq:gradient_c} and eq.~\ref{eq:direct_l}, substituting
  \(\bar{y}\) by \(\bar{Z} \in \R^{NM \times F}\), a rearrangement of
  \(Z\).
\item
  The gradient of \(\min_Z L\) is \[ \nabla_Z L =
  \frac{2}{N} w (w^T Z - y^T) + \frac{2\tau_F}{N} (Z - \overline{\bar{X} \theta}) \]
  and the closed-form solution is \[ Z^* = \argmin_Z L =
  \left( \tau_F I_{FM} + w w^T \right)^{-1}
  \left( \tau_F \overline{\bar{X} \theta} + w y^T \right). \]
\item
  The gradient and solution to \(\min_w L\) are given by
  eq.~\ref{eq:gradient_rls} and eq.~\ref{eq:sol_rls}, substituting \(X\)
  by \(Z\).
\end{enumerate}

\subsubsection{Graph coarsening}\label{graph-coarsening-1}

To reduce again the number of parameters in \(w\), the graph should be
coarsened to trade feature resolution at the price of vertex resolution.

\subsubsection{Kernel methods}\label{kernel-methods}

The same can be done for kernel methods.

We want to learn a classification function
\(f(x) = \sum_i \K(x,x_i) \alpha_i \in \R\), where \(\K(\cdot,\cdot)\)
is some kernel function, which maps a sample \(x\) to a label \(y\). The
loss function to minimize over the training set is then given by
\[ L = \sum_i |f(x_i) - y_i|^2 + \tau \norm{f}_{H_\K}^2 =
\norm{ \K \alpha - y }_2^2 + \tau \alpha^T \K \alpha, \] where
\(\K_{i,j} = \K(x_i, x_j)\) is the kernel function evaluated at all
training samples and \(\alpha \in \R^N\) are the parameters.

\[ \nabla_\alpha L = \K (\K \alpha - y) + \tau \K \alpha \] \(\K\) is
symmetric.

\[ \alpha^* = \argmin_\alpha L = (\K + \tau I_N)^{-1} y \]

\subsubsection{Feature learning}\label{feature-learning-1}

A general formulation of the loss for a classification problem is given
by \[ \min_{\theta_A, \theta_h} \left\{ L = \frac{1}{N} \sum_{i=0}^{N-1}
\left| h_{\theta_h}(A_{\theta_A}(\L) x_i) - y_i \right|^2 +
\tau_A \Omega_A(A_{\theta_A}) + \tau_h \Omega_h(h_{\theta_h}) \right\} \]
where \(A\) is an operator which transforms the input signals into a
suitable, i.e.~learned, representation and \(h\) is a classification
function. \(\Omega_A\) and \(\Omega_h\) are regularizations, \(\tau_A\)
and \(\tau_h\) are hyper-parameters.

The SVM classification function \(h \in H_\K\) is defined as
\[ h_{\theta_h}(x) = \sum_{i=0}^{N-1} \alpha_i \K(x, x_i) + \beta,
\ \theta_h = (\alpha_i, \beta), \] where \(\K(\cdot,\cdot)\) is some
kernel function. It is regularized by
\[ \Omega_h = \norm{h}_{H_\K}^2. \]

In our graph filters learning setting, the mapping
\(A \in \R^{FM \times M}\) is given by the filter bank
\[ A_{\theta_A}(\L) = \bigcup_{f=0}^{F-1} g_{\theta_f}(\L),
\ \theta_A = (\theta_f), \] a block-diagonal matrix composed of \(F\)
filters, which extracts multiple features from the input signal.

While the optimization complexity would increase a lot if we were to
minimize for the model complexity \(K\), we can easily include a penalty
of the form
\[ \Omega_A = \sum_f \sum_i |\theta_f(\lambda_i) w(\lambda_i)|^2 \]
where \(w(\lambda_i) \sim 1 / \lambda_i\).

Learning complexity: there is \(FK\) parameters in \(\theta_A\) and
\(N+1\) parameters in \(\theta_h\).

\subsection{Going deep}\label{going-deep}

Define convolutional neural networks on irregular domains by learning
multiple layers of graph filters.

Image of the architecture.

\section{Experiments}\label{experiments}

\subsection{Filter learning}\label{filter-learning}

Example from the Juypiter notebook.

Ground truth filter. where \(c_\ell = g(\lambda_\ell)\) is the evaluated
filter.

(randomly generated) graph signals and a filter \(g(\lambda)\) for some
\(t\), we want to learn the filter coefficients \(C\), a diagonal
matrix, such that \[\min_C \sum_{i=1}^N \norm{ U^T C U x_i - y_i }_2^2\]
where \(U\) are the eigenvectors of the graph Laplacian \(L\) and
\(y_i = x_i * g + \varepsilon, \varepsilon \sim \mathcal{N}(0,\epsilon)\)
is the target signal (the filtered signal \(x_i\) with additive Gaussian
noise).

\subsubsection{Non-parametrized filter}\label{non-parametrized-filter}

\subsubsection{Spline Parametrization}\label{spline-parametrization}

The authors of \citep{bruna_spectral_2013, henaff_deep_2015} learned
graph filters parametrized by cubic splines.

\subsubsection{Chebyshev
Parametrization}\label{chebyshev-parametrization-1}

\subsubsection{Lanczos Parametrization}\label{lanczos-parametrization-1}

\subsection{Classification}\label{classification-1}

\subsubsection{Sampled MNIST}\label{sampled-mnist}

\subsubsection{Text documents}\label{text-documents}

\section{Discussion}\label{discussion}

\section{Conclusion and Outlook}\label{conclusion-and-outlook}

That was awesome.

\subsubsection*{Acknowledgments}

Thanks ...

\section*{References}

\end{document}

### 2016-10-28 nips2016/nips2016.tex
nips v2: change line terminators (2016-05-13)

@@
 
-\title{Convolutional Neural Networks for Graphs}
+\title{Convolutional Neural Networks on Graphs\\ with Fast Spectral Filtering}
 
@@
 \end{abstract}
+
+%%% XAV %%%
+\section{Story}
+(1) Several DL approaches to perform DL on networks s.a.
+bruna, johnson.
+It has become an essential problem to go beyond existing DL techniques.
+Our main contribution is to define DL on graphs using rigorous math framework based on signal processing on graphs tools.
+\\
+(2) We introduce efficient computational techniques that scales w.r.t. E and not $n^2$, and simultaneously offer control of spatial local filters\\
+(3) We test w.r.t. fully connected and bruna on mnist, 20news, rcv1\\
+(4) Such framework offers the necessary basis to be developed in several directions s.a. johnson, natural graphs, etc.\\
+%%% XAV %%%
+
+%%% XAV %%%
+\section{Outline}
+(1) Introduction \& Motivation\\
+(2) Signal Processing on Graphs\\
+(3) CNNs on Graphs\\
+(4) Fast Graph filtering\\
+(5) Graph Coarsening with Kron Reduction\\
+(6) Experiments\\
+%%% XAV %%%
+
+%%% XAV %%%
+\section{Contributions}
+$\bullet$ Spectral extension of CNNs using signal graph processing paradigm \cite{shuman_emerging_2013} for shift-invariance property on graphs.\\
+$\bullet$ Fast graph filtering using Chebyshev and Lanczos techniques (no costly EVD required).\\
+$\bullet$ Exact control of spatial filtering size for optimal local feature extraction on graphs.\\
+$\bullet$ Less (?) parameters to learn than original CNNs.\\
+$\bullet$ Experiments on unstructured text document data with classification of RCV1 (Reuters), 20news, WEBKB4.\\
+%%% XAV %%%
 

### 2016-10-28 nips2016/nips2016.tex
nips v3: submitted version, mostly written by Xavier (2016-05-20)

--- 2016-10-28 nips2016/nips2016.tex
@@
 \usepackage{microtype}      % microtypography
+\usepackage{subfigure}
+\usepackage{graphicx}
+\usepackage{setspace}
+\usepackage{color}
 
-\title{Convolutional Neural Networks for Graphs}
+\title{Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering}
 
@@
 \begin{abstract}
-This paper introduces Graph Neural Network (GNN), a novel framework to process
-data on irregular, i.e.~non-Euclidean domains modelized as graphs. Even recent
-developments in Deep Learning who use graph embeddings don't make use of the
-graph structure per se.
+Convolutional neural networks have greatly improved state-of-the-art performances in computer vision and speech analysis tasks, due to its high ability to extract multiple levels of representations of data.
+In this work, we are interested in generalizing convolutional neural networks from low-dimensional regular grids, where image, video and speech are represented, to high-dimensional irregular domains, such as social networks, biological graphs like gene regulatory and brain connectivity networks, telecommunication networks, or words' embedding.
+We present a formulation of convolutional neural networks on graphs in the context of the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design fast localized convolutional filters on graphs.
+Numerical experiments on MNIST and 20NEWS demonstrate the ability of the system to learn local stationary features on graphs, as long as the graph is well-constructed.
 \end{abstract}
 
-\section{Introduction}\label{introduction}
+%Previous works have been developed along this today essential line of research, but they lack of mathematical foundations and efficient learning of filters.
 
-Large networks allow to learn complex relationships. Although they
-overfit easily if the training set is small because there is a lot of
-parameters to learn.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.eps}
+%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
+\caption{Architecture of the proposed CNNs on graphs.
+Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at layer $l$, $G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
+\label{fig:}
+\end{figure}
 
-Convolutional neural networks (CNN) offer a great reduction of
-parameters on regular Euclidean domain. This work proposes to generalize
-it to the discrete model of manifolds, graphs. Such an approach was
-proposed by \citep{bruna_spectral_2013, henaff_deep_2015}, we will
-refine it with proper graph signal processing tools.
+\section{Introduction}
+\vspace{-0.2cm}
+Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST} offer an efficient architecture to extract highly meaningful statistical patterns in large-scale and high-dimensional datasets.
+The key success of CNNs is its advanced ability to learn local stationary structures and compose them to form multi-scale hierarchical patterns.
+Precisely, CNNs extract the local stationarity property of the input data or signals by revealing local features that are shared across the data domain.
+These similar features are identified with localized convolutional filters or kernels, which are learned from the data.
+Convolutional filters are shift- or translation-invariant filters, meaning they are able to recognize identical features independently of their spatial locations.
+Localized kernels or compactly supported filters refer to filters that extract local features independently of the input data size, with a support size that can be much smaller than the input size.
+\\
+The compositional local stationarity property of CNNs can be efficiently implemented on regular grids, like image and sound domains, as convolutional and downsampling operators are mathematically well-defined on such discretized Euclidean spaces.
+This has led to the breakthrough of CNNs in image, video, and sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on low-dimensional regular lattices.
+But not all data lie on regular grids.
+User data on social networks, gene data on biological regulatory networks, log data on telecommunication networks, or text documents on words' embedding are important examples of data lying on irregular or non-Euclidean domains.
+Unfortunately, CNNs cannot be directly applied to such complex domains.\\
+A generalization of CNNs to irregular grids is not direct as the basic convolutional and downsampling operators, necessary to extract compositional local stationarity features, are not mathematically clear on these domains.
+This makes this extension mathematically challenging, both in theory and in implementation.
+The goal of this work is to leverage the recent works of \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs,art:HammondVandergheynstGribonval11GraphWav,art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG,art:ShumanRicaudVandergheynst16Local} to introduce an efficient implementation of CNNs on irregular domains represented here as graphs.
+Graphs are universal representations of heterogeneous pairwise relationships between possibly high-dimensional data.
+Graphs can encode complex geometric structures, and can be studied with strong mathematical tools such as graph spectral theory, a blend between graph theory and harmonic analysis.\\
+The major bottleneck of generalizing CNNs to graphs is the design of fast and localized filters on graphs, that are the core elements to extract local stationarity structures.
+Our major goal is to overcome this barrier.
+We will make use of recent tools developed in the context of signal processing on graphs (SPG) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to precisely control the spatial support of graph filters using the so-called polynomial kernel localization \cite{art:ShumanRicaudVandergheynst16Local}, and to perform an efficient backward propagation technique using the Chebyshev technique \cite{art:HenaffBrunaLeCun15DLgraphs}.
+Eventually, a generic and efficient implementation of CNNs on graphs may pave the way to the development of a potentially large class of deep learning techniques for graph-based data.
 
-This paper introduces a framework to use CNN with signals defined on
-weighted graphs. Our work leverages spectral methods, which are much
-faster than a spatial approach would be, in the same sense as doing
-convolutions as multiplication in the Fourier domain (with the help of
-the FFT) is much faster than multiplying patches on an image.
+\section{Related Works}
+\vspace{-0.2cm}
+{\bf CNNs on non-Euclidean domains.} Extending the success of CNNs to non-Euclidean domains is essential to boost the power analysis of data lying on complex networks like biological, social or telecommunication networks.
+This is a recent line of works, with a preliminary approach coming from the local reception fields in \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF}, and applied successfully to object recognition in computer vision.
+The main idea is to unsupervisedly learn groups of similar features.
+This is highly useful to automatically choose a limited number of connections between two successive layers.
+This approach is however not able to extract similar groups over the data domain, as it is not clear how to define convolutional filters on these groups.
+\\
+In \cite{art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, the authors proposed a generalization of CNNs on graphs using graph counterparts of grid convolution and downsampling.
+They were able to learn shift-invariant or convolutional filters on graphs using the convolution theorem \cite{book:Mallat99wavelets} that defines convolutional operators as linear operators that diagonalize in the Fourier basis (represented by the eigenvectors of the Laplacian operator).
+They also designed a strategy to learn the graph structure from the data, and apply their model on image recognition, text categorization and bioinformatics.
+This promising approach is however limited in implementation as forward and backward evaluations have complexity $O(n^2)$ due to the matrix multiplication operations that operate the graph Fourier transforms.
+Besides, there is no strong control over the support size of the convolutional kernels, which is essential to design localized filters to extract local stationary features, independent of the input size.
+\\
+The authors in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,art:MasciBoscainiBronsteinVandergheynst15ShapeNet} introduced ShapeNet, a generalization of CNNs to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces.
+They used geodesic polar coordinates to define mesh patches and formulated a deep learning architecture.
+They obtained state-of-the-art results for shape recognition tasks.\\
+Finally, the papers \cite{pro:ChenChengMallat14deepHaar,pro:RustamovGuibas14deepHaar} investigated the construction Haar wavelet transforms on graphs using a deep hierarchical architecture.
+They applied these deep wavelets on graphs to object recognition on sphere, and to sparse reconstruction of faces.
+\\
+{\bf Signal Processing on Graphs.} SPG is an emerging field that aims at bridging the gap between signal processing techniques like wavelet analysis and graph theory such as spectral graph analysis \cite{art:BelkinNiyogi05LaplaBeltrami,art:VonLuxburg07Tutorial}.
+The main goal is to generalize fundamental analysis operations for signals from regular grids to irregular structures embodying by graphs.
+We refer the reader to \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} for an introduction of the field.
+Standard operations on grids such as convolution, translation, filtering, dilatation, modulation, downsampling do not extend directly to graphs and thus require new mathematical definitions while keeping the original intuitive concepts.
+In this context, the authors of \cite{art:HammondVandergheynstGribonval11GraphWav,art:CoifmanLafon06DifMap,pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet operators on graphs.
+In \cite{art:ShumanFarajiVandergheynst16PyramTrans,art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform mutli-scale pyramid transforms on graphs.
+The works of \cite{pro:TsitsveroBarbarossa15Uncert,pro:PasdeloupAlamiGriponRabbat15Uncert,art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles on graphs, and showed that intuitive concepts may be lost, but can also produce enhanced localization principles for signals on graphs.
+In \cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to carry out lasso-based signal regularization on graphs, and studied the intertwined relationships between smoothness and sparsity on graphs.
+In \cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors investigated compressed sensing recovery conditions for graph spectral signal analysis.
+\\
+Beyond this work, our long-term goal is to make use of SPG tools as much as possible to formulate mathematically well-designed techniques of CNNs on graphs.
+This paper introduces the mathematical and computational foundations of such models, while future developments may consider enhanced graph filters like graph wavelet operators, spectral graph coarsening techniques, and improved localization properties.
 
-Another option to do clustering or regression on data graphs is to first
-embed the graph and then use the embedding as additional features.
+\section{Contributions}
+\vspace{-0.2cm}
+{\textbf {\textit {(1)}}} Learning deep fully connected networks is almost computationally untractable for large data domain.
+Fortunately, for data satisfying stationarity and hierarchical compositionality structure, CNNs have proven to be remarkably powerful to identify highly informative features using low learning complexity.
+For example, learning a single layer with $n$ input coordinates and $m<n$ outputs require $O(n^2)$ complexity to learn fully connected networks.
+It also requires the complexity $O(n)$ to learn any arbitrary filter, and $O(S)$ for localized filters, where $S$ is the size of the filter support, independent of the input size $n$.
+One of our main contributions is to stay with the same low complexity for learning graph convolutional filters.
+We will show it is actually possible to keep exactly the same complexity as standard CNNs, i.e.
+$O(S)$ for localized graph filters.\\
+{\textbf {\textit {(2)}}} The backpropagation learning algorithm requires to run convolutional operations, which can be efficiently implemented by FFT on grids with complexity $O(n\log n)$.
+Convolution on graphs can also be implemented by matrix multiplications representing the forward and backward graph Fourier transforms, but it is expensive as the complexity is $O(n^2)$.
+As noticed in \cite{art:HenaffBrunaLeCun15DLgraphs}, this complexity is a strong bottleneck to deal with large-scale problems.
+We will greatly improve this complexity by reducing to $O(E.S)$, where $E$ is the number of edges in the given graph, using the Chebyshev technique from SPG \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
+Besides, as most real-world graphs are highly sparse, we have $E\ll n^2$, and therefore a low complexity for the matrix multiplication operations.
+Eventually, the Chebyshev computations can be naturally distributed on GPUs, as we did in our experiments.\\
+{\textbf {\textit {(3)}}} We introduce an efficient pooling strategy for CNNs on graphs, that also makes use of distributed computing architecture like GPUs.
+The pooling approach consists on two stages.
+First, a graph coarsening procedure based on METIS \cite{art:DhillonGuanKulis07Graclus,art:KarypisKumar98Metis} is successively applied to the original graph to produce coarser versions of the graph.
+Second, at each coarsening level, we sequentially order all nodes such that two successive indexed nodes are the parent nodes of the child node in the next coarser level.
+This node indexation structure in memory satisfies parallel architectures.
+\\
+{\textbf {\textit {(4)}}} We regard this proposed work as the funding paper that bridges the gap between two powerful fields, namely deep learning and signal processing on graphs.
+In the future, we expect to further benefit from tools in SPG to enhance the formulation of CNNs on graphs.
+\\
+{\textbf {\textit {(5)}}} We revisit the Euclidean case in the numerical experiments with the standard MNIST image classification problem.
+The model is able to learn localized convolutional filters and achieve 98.65\% of success rate.
+More interestingly, we apply our model to the text categorization problem on 20NEWS.
+We also demonstrate that our techniques is able to extract local stationary features, as it improves by 2.5\% the classification accuracy when no convolutional are present.
+We observe that the quality of the graph construction is strongly linked to the quality of the convolutional layers.
+Finally, we demonstrate the computational performance of our model over \cite{art:HenaffBrunaLeCun15DLgraphs}.
 
-\section{Related work}\label{related-work}
+\section{Learning Localized Graph Filters}
+\vspace{-0.2cm}
+In this section, we explore how to design efficient localized convolutional filters on graphs.
 
-Comparison with fully connected NN (Reuters used in
-\citep{srivastava_dropout_2014})
+\subsection{Signal Processing on Graphs and Localized Filters}
+\vspace{-0.2cm}
+{\bf Graph Fourier Transform.} We are interested in processing signals defined on undirected and connected graphs $G=(V,E,W)$, where $V$ is the set of vertices with $|V|=n$, $E$ is the set of edges and $W$ is the weighted adjacency matrix encoding the connection weights between two vertices.
+A signal $x:V\rightarrow\mathbb{R}$ defined on the nodes of the graph may be regarded as a vector $x\in\mathbb{R}^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node.
+An essential operator in spectral graph analysis is the graph Laplacian \cite{book:Chung97Spectral}, which unnormalized definition is $L=D-W\in\mathbb{R}^{n\times n}$ with $D\in\mathbb{R}^{n\times n}$ is the diagonal degree matrix where $D_{ii}=\sum_jW_{ij}$, and its normalized definition is $L=I_n-D^{-1/2}WD^{-1/2}$ where $I_n$ is the identity matrix.
+As $L$ is a real symmetric positive semidefinite matrix, it has a complete set of orthonormal eigenvectors $\{u_l\}_{l=0}^{n-1}\in\mathbb{R}^n$, known as the graph Fourier modes, and their associated real nonnegative eigenvalues $\{\lambda_l\}_{l=0}^{n-1}$, which are the frequencies of the graph system.
+Equivalently, linear operator $L$ diagonalizes the Fourier basis matrix $U=[u_0,...,u_{n-1}]\in\mathbb{R}^{n\times n}$ such that $L=U\Lambda U^T$ where $\Lambda=\diag(\lambda_l)$.
+We can now define the graph Fourier transform (GFT) and its inverse \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
+The GFT of a spatial signal $x\in\mathbb{R}^n$ is $\hat{x}=U^Tx\in\mathbb{R}^n$, and the inverse GFT is $x=U\hat{x}$.
+Using GFT we can now generate fundamental operations of signal processing on graphs such as convolution and filtering.
+\\
+{\bf Convolutional Filters.} The graph convolutional operator is simply defined in the Fourier space such as $f\ast_G g=U((U^Tf) \odot (U^Tg))$, where $\odot$ the element-wise Hadamard product.
+Then the convolutional filtering operation on graphs is defined as $y=g_\theta(L)x\in\mathbb{R}^n$, where $g_\theta(L)=g_\theta(U\Lambda U^T)=Ug_\theta(\Lambda) U^T\in\mathbb{R}^{n\times n}$ where $g_\theta(\Lambda)=\diag(\theta)$ and $\theta\in\mathbb{R}^n$ is the filter expressed in the Fourier space.
+Remember that convolutional filters, i.e.
+translation-invariant filters on graphs, can only be defined in the Fourier domain as convolution operators are not defined in the vertex space.
+\\
+{\bf Localized Convolutional Filters.} Learning convolutional filters on graphs boils down to learn spectral multipliers $\theta$ of complexity $O(n)$, the size of the input data.
+Our goal is to reduce this complexity to $O(S)$, the support size of the filter in the vertex space, exactly as in standard CNNs.
+For this, we will make use of a property called polynomial localization \cite{art:ShumanRicaudVandergheynst16Local}.
+Let us represent $\theta=\{\theta_{\{{\lambda_l}\}_{l=0}^n}\}\in\mathbb{R}^n$ as a $K$-th differentiable polynomial function: $\theta_{\lambda_l}^K(a)=\sum_{k=0}^{k-1}a_k \lambda_l^k$, where $a=\{a_k\}_{k=0}^{K-1}\in\mathbb{R}^K$.
+The polynomial localization theorem \cite{art:ShumanRicaudVandergheynst16Local} shows that $(T_ig_{\theta^K})_j=(\delta_i \ast_G (Ug_{\theta^K}))_j=0$ for $d_G(i,j)>K$ where $d_G$ is the shortest distance on graph, meaning that the support of the associated spatial filter is exactly $K$-localized, all values are zero after $K$ hopes.
+Consequently, spectral filters represented by $K$-order polynomials are spatially localized with support $S=K$, implying the low learning complexity $O(K)$ for each filter.
+We notice that the authors \cite{art:HenaffBrunaLeCun15DLgraphs} also used a $K$-order polynomials represented by splines, but they did not formalize the quantitive relationship between filter support size and polynomial order.
+Finally, it is also possible to estimate the spatial decaying profile of the filter \cite{art:ShumanRicaudVandergheynst16Local}: for any K-localized filter centered at vertex $i$, we have the profile $|(T_ig_{\theta^K})(j)|\leq\big[\frac{2\sqrt{n}}{K!}(\frac{\lambda_{max}}{4})^K \sup_{\lambda}|(Ug_{\theta^K})^K(\lambda)|\big]$, which may be useful for graph filter analysis.
 
-\begin{itemize}
-\item
-  + less parameters via convolution
-\item
-  ? accuracy
-\end{itemize}
+\subsection{Fast Filter Learning with Chebyshev Technique}
+\vspace{-0.2cm}
+Let $y$ be a signal $x$ filtered on the graph: $y=g_\theta(L)x=U(g_\theta(\Lambda) (U^Tx))$.
+Learning a spectral filter can be reduced to a complexity $O(K)$ by polynomial parametrization.
+However, the numerical computation of $y$ still requires two full matrix multiplications by $U^T,U$ with a high complexity of $O(n^2)$, and these matrix multiplications must be done several times in the backpropagation process.
+This is a computational bottleneck, identified in \cite{art:HenaffBrunaLeCun15DLgraphs}, with an additional cost of $O(n^3)$ for computing a full eigenvalue decomposition of $L$ to get $U$.
+We can overcome this computational issue using a Chebyshev technique as proposed in SPG \cite{art:HammondVandergheynstGribonval11GraphWav}.
+We will also explore in the future the Lanczos technique \cite{art:SusnjaraPerraudinKressnerVandergheynst15Lanczos}, which seems attractive because of the independence of the coefficients.
+\\
+We first recall the fundamental recurrence property of Chebyshev polynomials.
+Let the Chebyshev polynomial \(T_k(x)\) of order \(k\) be generated by the recurrence relation \(T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)\) with \(T_0 = 1\) and
+\(T_1 = x\).
+These polynomials form an orthogonal basis.
+Then, the K-compact spectral filter $g_{\theta^K}(\Lambda)$ can be defined as $g_{\theta^K}(\Lambda)=\sum_{k=0}^{K-1}\theta_k T_k(\tilde{\Lambda})$ of order \(k\) evaluated at
+\(\tilde{\Lambda} := 2 \Lambda / \lambda_{max} - I_n \in \R^{n \times n}\), a diagonal matrix of scaled eigenvalues such that they lie in \([-1,1]\).
+Consequently, the filtering operation is defined as $y=g_{\theta^K}(L)x=Ug_{\theta^K}(\Lambda) U^Tx=\sum_{k=0}^{K-1} U \theta_k T_k(\tilde{\Lambda}) U^Tx=\sum_{k=0}^{K-1} \theta_k T_k(\tilde{L}) x$, where \(T_k(\tilde{L}) \in \R^{n \times n}\) is the Chebyshev polynomial of order \(k\) evaluated at the scaled Laplacian
+\(\tilde{L} := 2 L / \lambda_{max} - I_n\).
+If we denote $X_k:=T_k(\tilde{\Lambda})x$ and using the recurrence equation, we can generate $X_k=2\tilde{L}X_{k-1}-X_{k-2}$ with \(X_0 = x\) and \(X_1 = \tilde{L} x\).
+As $L$ is sparse, it is straightforward to see that all matrix multiplications are done between sparse matrices, greatly decreasing the computational complexity for graph filtering from $O(n^2)$ to $O(E.K)$.
+Finally, matrix multiplications can be efficiently carried out in a distributive way, allowing an efficient GPU implementation of Chebyshev filtering.
 
-Comparison with first generation graph CNN \citep{henaff_deep_2015}.
+\section{Graph Coarsening and Fast Pooling}
+\vspace{-0.2cm}
+Local stationarity property of data is extracted via localized convolutional kernels.
+In this section, we are interested to extract the multi-scale hierarchical composition property of data.
+This is efficiently achieved via grid subsampling or downsampling and pooling, which trades spatial resolution with feature resolution reducing the learning complexity without compromising the system performances.
+In the case of standard CNNs for data lying on regular grids, the subsampling operation is mathematically sound.
+The translation operator commutes with the downsampling operator, hence not affecting the translation invariance property of the system at different scales of grid resolutions.\\
+How to reduce grid size in the case of graphs?
+We need to construct meaningful neighbourhoods on graphs where similar features may be clustered together.
+This naturally boils down to compute a multi-scale clustering of the graph that preserves local geometric structures.
+Unfortunately, it is well-known that graph clustering is a NP-hard problem \cite{art:BuiJonesGraphPartNPhard}, and approximations must be explored.
+Several clustering techniques exist like spectral clustering \cite{art:VonLuxburg07Tutorial}.
+However, we are more interested in clustering techniques that can reduce the size of the graph by a factor 2, which offers a precise control on the coarsening process.
+This is essential to keep control on the reduction factor to be able to extract fine hierarchical compositional structures.
+Algebraic multigrid techniques on graphs \cite{art:RonSafroBrandt11MultigridGraph} can be used in this context.\\
+In this work, we make use of the downsampling or coarsening technique introduced in METIS \cite{art:DhillonGuanKulis07Graclus,art:KarypisKumar98Metis} which has shown to be extremely efficient at clustering a large variety of graphs.
+METIS uses a greedy algorithm to compute successive coarser versions of a given graph.
+Its greedy rule consists, at any given coarsening level, in matching two nodes that locally maximizes a balanced cut energy like normalized cut \cite{art:ShiMalik00NCut}, which are excellent clustering energies.
+Precisely, we randomly pick an unmarked vertex $i$ and match it with one of its unmarked neighbors $j$ that maximizes the local normalized cut $\frac{W_{ij}}{d_i}+\frac{W_{ij}}{d_j}$.
+The two matched vertices are then marked.
+The matching is repeated until all nodes have been explored.
+This is an extremely fast coarsening technique, which divides by almost a factor 2 the size of the nodes from one level to the next coarser level.
+The coarsening factor is not exactly as there exist a few singletons, non-matched nodes.
+We can however control the reduction factor and set it to exactly 2 by simply adding artificial disconnected nodes to the singletons.
+We then order the nodes such that two successive indexed nodes are the parent nodes of the child node in the next coarser level.
+The major reason to do this is to benefit from parallel hardware architectures like GPUs to considerably speed up the pooling operations, which must be done several times during backprogation.\\
+In a future work, we will explore the use of SPG by using a Kron reduction technique \cite{art:ShumanFarajiVandergheynst16PyramTrans}.
+The Kron graph coarsening technique has the main advantage to preserve the ordering of the Laplacian spectrum, and may be able to commute with the graph convolution operator.
+If proven, this would be a very essential mathematical property of CNNs on graphs, exactly as regular grid downsampling in standard CNNs.
 
-\begin{itemize}
-\item
-  + better accuracy
-\item
-  = same number of parameters
-\item
-  - graph construction vs supervised estimation
-\item
-  Theoretical:
+%\newpage
 
-  \begin{itemize}
-  \item
-    + proper spectral graph signal processing tools
-  \item
-    + kNN graph (sparse weight matrix) vs fully connected graph. SVD
-    complexity for Fourier basis is O(n\^{}3), we avoid it.
-  \item
-    + METIS vs naive agglomeration
-  \item
-    + Chebyshev polynomials vs splines to approximate filters Filtering
-    from \(O(n^2)\) to \(O(|E| K)\) where \(K\) is the polynomial order
-    approximately \(O(n)\) if the weight matrix is sparse, i.e.~grows
-    linearly with \(N\) This addresses the first limitation mentioned in
-    \citep{henaff_deep_2015}.
-  \end{itemize}
-\end{itemize}
+\section{Numerical Experiments}
+\vspace{-0.4cm}
+We performed the numerical experiments with TensorFlow, an open source software library for deep learning developed by the Google Brain Team, which features various backends, notable CUDA for GPU computing.
 
-\section{Model}\label{model}
+\subsection{Revisiting Euclidean CNNs on MNIST}
+\vspace{-0.2cm}
+Our first experiment aims at reproducing the performance of standard CNNs on the highly popular MNIST benchmark dataset \cite{pro:LeCunBottouBengioHaffner98MNIST}.
+This is an important sanity check for our proposed generalized model.
+It should be able to recover the original object classification rate as the convolutional layers are also designed to extract local stationary features on any graph, including the 2D grid graph on which images reside.
+\\
+MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28\times 28$, such that data points lie on a space of $784$ dimensions.
+TensorFlow reports an accuracy of 99.20\% on MNIST with a LeNet5-like CNN architecture.
+Our proposed CNNs on graphs with a graph of $8$-NN (nearest neigbors) of a 2D Euclidean grid achieves a precision of 98.65\%, as reported in Table \ref{tab1}.
 
-\subsection{Spectral Graph Theory}\label{spectral-graph-theory}
+\begin{table*}[h!]
+ \centering
+{\small
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+ Algorithms & Accuracy  \\
+\hline
+Linear SVM & 91.76  \\
+Softmax & 92.36  \\
+CNNs [LeNet5-like] & 99.20  \\
+graph CNNs: CN32-P4-CN64-P4-FC512-softmax & 98.65 \\
+\hline
+\end{tabular}
+}
+\caption{Classification results on MNIST.} 
+\label{tab1}
+\end{table*}
 
-A graph \(\G = (\V, \E, \W)\) is defined by a set \(V\) of \(|V| = M\)
-nodes and a set \(E\) of weighted edges. The connectivity of the graph
-is captured by the adjacency matrix \(W \in \R^{M \times M}\) which
-entry \(W_{i,j}\) denotes the weight of the edge \((v_i, v_j) \in E\)
-which connects the vertex \(v_i \in V\) to \(v_j \in V\). It is set to
-\(0\) if the vertices are not connected, i.e. \((v_i, v_j) \notin E\).
-Assuming an undirected graph, \(W\) is a symmetric matrix. A graph
-signal is any signal \(x \in \R^M\) defined on the vertices of \(G\).
+\noindent
+We report in Table \ref{tab2} the speedup offered by GPUs.
+A GPU implementation of graph CNNs is 8x faster than a CPU implementation, i.e.
+the same order of speedup than standard CNNs, as it only uses matrix multiplications which are efficiently implemented in CUDA BLAS.
 
-The combinatorial graph Laplacian is defined as
-\[ \L^c := D - W \in \R^{M \times M}, \] where \(D \in \R^{M \times M}\)
-is the diagonal degree matrix defined as \(D_{i,i} := \sum_j W_{i,j}\).
-Note that it is a difference operator such that
-\[ (\L^c x)_i = \sum_j W_{i,j} (x_i - x_j). \] The normalized graph
-Laplacian is defined as
-\[ \L^n := I_M - D^{-1/2} W D^{-1/2} \in \R^{M \times M}, \] where
-\(I_M \in \R^{M \times M}\) is the identity matrix. Finally, the
-random-walk graph Laplacian is defined as
-\[ \L^{rw} := I_M - D^{-1} W \in \R^{M \times M}. \] Note that this work
-is independent of the chosen graph Laplacian \(\L\), which can be any of
-\(\L^c\), \(\L^n\) or \(\L^{rw}\).
+\begin{table*}[h!]
+ \centering
+{\small
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+ Architecture & Time  & Speedup \\
+\hline
+CNNs with CPU & 210ms & -\\
+CNNs with GPU NVIDIA K40c & 31ms & 6.77x\\
+graph CNNs with CPU & 200ms & -\\
+graph CNNs  with GPU NVIDIA K40c & 25ms & 8.00x\\
+\hline
+\end{tabular}
+}
+\caption{Computational time for processing a mini-batch of 100 images.} 
+\label{tab2}
+\end{table*}
 
-In analogy to the real line Fourier Transform, a Fourier basis is given
-by the eigenvectors of the Laplacian
-\[ \L u_i = \lambda_i u_i \st \norm{u_i}_2=1,
-\ u_i \in \R^M, \ i = 0, \ldots, M-1, \] with their associated
-eigenvalues \(\lambda_i\). Assuming the graph is connected, we may order
-the vector \(\lambda := [\lambda_0, \ldots, \lambda_{M-1}]^T \in \R^M\)
-of eigenvalues such that
-\[ 0 = \lambda_{min} = \lambda_0 < \lambda_1 \leq \ldots \leq
-\lambda_{M-1} = \lambda_{max}. \] As the Laplacian is a real symmetric
-and \gls{PSD} matrix, the eigenvalues are real and positive, and the
-eigenvectors are orthonormal. The Laplacian is indeed diagonalized by
-the Fourier basis \(U := [u_0, \ldots, u_{M-1}] \in \R^{M \times M}\)
-such that
-\begin{equation} \L = U \Lambda U^T, \label{eq:lap_diag}\end{equation}
-where \(\Lambda := \diag(\lambda) \in \R^{M \times M}\) denotes a
-diagonal matrix of eigenvalues and \(U^T\) is the matrix transpose of
-\(U\). See \citep{chung_spectral_1997} for details on spectral graph
-theory.
+\noindent
+We now discuss the influence of the graph quality over the performance of the CNNs on graphs model.
+Convolutional layers aims at extracting local stationary features, or repetitive feature patches over the image domain.
+This implies 2 requirements to be successful.
+First, the data must satisfy this property, which we know to be true for MNIST.
+Second, the graph must be well constructed to be able to extract local stationary features.
+In the case of images, a simple $k$-NN graph of the grid is good enough, and the value of $k$ does not have a strong influence on the results.
+However, let us destroy the graph structure by using a random graph, then the accuracy drops to 95.39\%, which is the order of a simple fully connected graph with a softmax classifier, that is when the convolutional layers are not used.
+Table \ref{tab3} reports the results.
 
-The Graph Fourier Transform \(\hat{x} \in \R^M\) of any graph signal
-\(x\) is given by \[ \hat{x} = U^T x =
-[\langle u_0, x \rangle, \ldots, \langle u_{M-1}, x \rangle]^T, \] where
-\(\langle \cdot , \cdot \rangle\) denotes the standard \(\ell_2\) inner
-product \citep{shuman_emerging_2013}. The inverse transform is then
-given by \[
-x = U \hat{x}. \]
+\begin{table*}[h!]
+ \centering
+{\small
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+ Architecture & $G_1=$ 2D Euclidean grid  & $G_2=$ Random graph \\
+\hline
+graph CNNs: CN32-softmax & 97.40 & 96.88\\
+graph CNNs: CN32-P4-CN64-P4-FC512-softmax & 98.65 & 95.39 \\
+\hline
+\end{tabular}
+}
+\caption{Influence of the graph quality over CNNs on graphs.} 
+\label{tab3}
+\end{table*}
 
-It follows that any signal \(\hat{x}\) can be filtered in the spectral
-domain by \[ \hat{y} = g_\theta(\Lambda) \hat{x}, \] where the operator
-\(g_\theta(\Lambda) \in \R^{M \times M}\), a matrix function, yields a
-diagonal matrix of \(M\) Fourier coefficients parametrized by
-\(\theta\). The filtering operation can equivalently take place in the
-vertex domain as \[ y = U g_\theta(\Lambda) U^T x = g_\theta(\L) x \]
-where the operator
-\(g_\theta(\L) = g_\theta(U \Lambda U^T) = U g_\theta(\Lambda) U^T \in \R^{M \times M}\)
-is akin to a convolution with the filter \(Ug_\theta(\Lambda)\) in the
-vertex domain. Note that the filtering operation is defined as a
-multiplication in the spectral domain because a convolution cannot be
-defined in the vertex domain \citep{shuman_emerging_2013}.
+%\newpage
 
-\subsection{Graph Filter Learning}\label{graph-filter-learning}
+\subsection{Text Categorization with the 20NEWS Dataset}
+\vspace{-0.2cm}
+For this section, we consider the text categorization task with the 20NEWS dataset which consists of 18,846 text documents associated with 20 classes \cite{art:Joachims9620NEWS}.
+The number of unique words in this corpus is 93,953, so the dimensionality of these data points is around 94k.
+We extract 10,000 words that most occur in the corpus.\\
+The original full dataset is split into a training set of 11,314 documents and a test set of 7,532 documents.
+Our goal is to classify the test set.
+We remind that standard CNNs cannot be directly applied to this problem.
+A natural alternative would be to use a deep fully connected networks, but they are computationally expensive.
+To decrease the computational learning complexity of fully connected networks, the authors of \cite{art:HenaffBrunaLeCun15DLgraphs} suggested to carry out dropout regularization \cite{art:SrivastavaHintonKrizhevskySutskeverSalakhutdinov14dropout} on these networks.
+Table \ref{tab4a} reports the classification performances of a set of algorithms on 20NEWS, and the proposed graph CNNs with one convolutional layer of 32 filters and a softmax classifier.
 
-Given a graph \(G\), a set
-\(X = [x_0, \ldots, x_{N-1}] \in \R^{M \times N}\) of \(N\) source graph
-signals of dimensionality \(M\) and their associated set
-\(Y = [y_0, \ldots, y_{N-1}] \in \R^{M \times N}\) of target signals, we
-want to learn the parameters \(\theta \in \R^M\) of a graph filter
-\(g_\theta(\Lambda) := \diag(\theta)\) such as to minimize the convex
-mean square error \begin{equation} L =
-\frac{1}{N} \sum_{i=0}^{N-1} \norm{ g_\theta(\L) x_i - y_i }_2^2 =
-\frac{1}{N} \norm{ U \diag(\theta) U^T X - Y }_F^2, \label{eq:loss}\end{equation}
-where \(\norm{\cdot}_2^2\) denotes the squared \(\ell_2\) norm and
-\(\norm{\cdot}_F^2\) the squared Frobenius norm.
+\begin{table*}[h!]
+ \centering
+{\small
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+ Algorithms & Word2vec features  \\
+\hline
+Linear SVM & 65.90  \\
+Multinomial NB & 68.51 \\
+Softmax & 66.28 \\
+FC + softmax + dropout & 64.64 \\
+FC + FC + softmax + dropout & 65.76 \\
+graph CNNs: CN32-softmax & 68.26 \\
+\hline
+\end{tabular}
+}
+\caption{Classification accuracies on 20NEWS with features learned by word2vec \cite{pro:MikolovChenCorradoDean13word2vec}.} 
+\label{tab4a}
+\end{table*}
 
-Note that, while being the most flexible definition of
-\(g_\theta(\Lambda)\), the independence of the Fourier coefficients
-\(\theta\) from the eigenvalues \(\Lambda\) does omit all information
-about frequencies. As we know that the lower frequencies are more
-important for clustering {[}ref NCut{]}, we could have designed a
-parametric filter with less parameters.
+\noindent
+As text documents do not naturally lie on a graph, it is here necessary to construct a graph of words' embedding.
+This is a very essential step as the graph quality will have a strong impact on the ability of convolutional layers to identify local stationary features present in the data.
+We also observe that local stationarity is an assumption that is made about the data, and the CNNs are fundamentally designed on this data hypothesis.
+However, we believe this stationarity assumption to be present in many real-world data, as structured data are usually composed of repetitive small patterns at different scales of observation.
+We investigate here 5 graphs of words: a graph of learned embedding of the corpus using word2vec \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of normalized word counts across each document,
+a graph of pre-trained embedding of words \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of approximate nearest neighbors (ANN) using LSHForest on the learned embedding \cite{pro:MikolovChenCorradoDean13word2vec}, and a random graph.
+The classification results are reported in Table \ref{tab4b}.
+It shows that the graph $G_1$ learned from the training set is the best at capturing the local stationarity property of text documents, improving the fully connected networks performance from 65.76\%, reported Table \ref{tab4a}, to 68.26\%.
+It is worth noticing that the approximate $k$-NN graph $G_4$ constructed from the best learned features is almost as worse as a random graph, meaning that it is a better strategy to learn a good low-dimensional words' embedding first and then construct an exact $k$-NN graph from this embedding.
+As the graph quality has an impact on the performance of the classification system, a next step would be to learn a better graph as already pointed in \cite{art:HenaffBrunaLeCun15DLgraphs}.
+A natural and future approach would be to alternate learning of the NN weights and learning of the graph of data, as a virtuous circle.
 
-Rewriting eq.~\ref{eq:loss} in the spectral domain while decomposing it
-w.r.t. the scalar coefficients \(\theta_i\) gives \[ L =
-\frac{1}{N} \norm{ \diag(\theta) U^T X - U^T Y }_F^2 =
-\frac{1}{N} \sum_{i=0}^{M-1} \norm{\theta_i\Xh_{i,\cdot}-\Yh_{i,\cdot}}_2^2, \]
-where \(\Xh = U^TX \in \R^{M \times N}\) and
-\(\Yh = U^TY \in \R^{M \times N}\) are the spectral representations of
-the signals \(X\) and \(Y\). The gradient for each coefficient is then
-given by \[ \nabla_{\theta_i} L =
-\frac{2}{N} ( \theta_i \Xh_{i,\cdot} - \Yh_{i,\cdot} ) \Xh^T_{\cdot,i} \]
-and can be rewritten in a vectorized form as
-\begin{equation} \nabla_{\theta} L =
-\frac{2}{N} \diag \left( (\diag(\theta) \Xh - \Yh) \Xh^T \right) =
-\frac{2}{N} \left( ( \theta 1_N^T \odot \Xh - \Yh ) \odot \Xh \right) 1_N,
-\label{eq:gradient}\end{equation} where \(1_N\) denotes a unit vector of
-length \(N\) and \(\odot\) the element-wise Hadamard product. The second
-form avoids the computation of the useless off-diagonal elements.
+\begin{table*}[h!]
+ \centering
+{\small
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+ Architecture & $G_1$= learned  & $G_2$= normalized  & $G_3$= pre-trained & $G_4$= ANN  &  $G_5$= random   \\
+  & graph \cite{pro:MikolovChenCorradoDean13word2vec} & word count graph & graph  &  graph  & graph \\
+\hline
+CN32-softmax & 68.26 & 67.50 & 66.98 & 67.86 & 67.75  \\
+\hline
+\end{tabular}
+}
+\caption{Classification accuracies of graph CNNs with different graph constructions on 20NEWS.} 
+\label{tab4b}
+\end{table*}
 
-A closed-form solution is given by the optimality condition
-\(\nabla_{\theta}L=0\) such that
-\begin{equation} \theta^* = \argmin_\theta L =
-(\Xh \odot \Yh) 1_N \oslash (\Xh \odot \Xh) 1_N, \label{eq:direct}\end{equation}
-where \(\oslash\) denotes an element-wise division. Note that this
-method is impractical for large \(M\) and \(N\) (sufficiently large for
-\(X\) and \(Y\) to not fit in memory). A \gls{SGD} based on
-eq.~\ref{eq:gradient} is then necessary.
+\subsection{Comparison with Non-Parametric Filters and Henaff-Bruna-LeCun \cite{art:HenaffBrunaLeCun15DLgraphs}}
+\vspace{-0.2cm}
+The advantages of our graph CNNs over \cite{art:HenaffBrunaLeCun15DLgraphs} are listed below:\\
+$\bullet$ Improved learning complexity from quadratic to linear: In \cite{art:HenaffBrunaLeCun15DLgraphs}, graph convolutional operations need $O(n^2)$ operations, while our technique requires $O(E.K)$ operations, where $E$ is the number of edges and $K$ is the polynomial order of the spectral filters.
+As we consider $k$-NN graphs, $E=n.k$ and the complexity becomes linear $O(n)$ w.r.t.
+the input size $n$, while  \cite{art:HenaffBrunaLeCun15DLgraphs} stays quadratic.
+To give some figures, in our experiments, we have for the first layer, $n=10^4, k=8, K=32$, so the total number of operations is $E.K=10^4.8.32 \simeq 3.10^5$ and $n^2=10^8$.
+Figure \ref{fig2} compares the computational times.\\
+$\bullet$  No eigenvalue decomposition required: The spectral graph CNNs \cite{art:HenaffBrunaLeCun15DLgraphs} require to compute a full eigenvalue decomposition (EVD) to compute the Fourier basis.
+This has a complexity of $0(n^3)$ with $n=10^{4}$, which is memory expensive.
+Our technique does not require an EVD, it only requires multiplications between the graph Laplacian sparse matrix and vectors.
+\\
+$\bullet$ Improved memory efficiency: \cite{art:HenaffBrunaLeCun15DLgraphs} needs to store in memory the Fourier basis of size $O(n^2)$, this can greatly limit the performance of GPU computing.
+Our technique requires to store the sparse Laplacian of size $O(E)=O(n)$ in the case of k-NN graphs.
 
-There are however two major computational drawbacks to this optimization
-process: 1) the Fourier transform \(\Xh=U^TX\) of a set of signals costs
-\(O(M^2N)\) operations and 2) the eigenvalue decomposition
-\(\L = U \Lambda U^T\) costs \(O(M^3)\). The total cost of filtering is
-thus \(O(M^2 \max(M,N))\), a problem already stated in
-\citep{henaff_deep_2015}. We propose to overcome these computational
-issues using efficient numerical approximations such as the Chebyshev
-polynomials or the Lanczos algorithm.
+%\begin{table*}[h!]
+ %\centering
+%{\small
+%\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+%\hline
+% Nb of words & Non-Param filters  & Splines filters \cite{art:HenaffBrunaLeCun15DLgraphs} & Chebyshev filters [Our model] \\
+%\hline
+%1k words & 25ms & 25ms & 18.75ms \\
+%2k words & 57.5ms & 58.75ms & 28.75ms \\
+%4k words & 170ms & 170ms & 50ms \\
+%6k words & 348.75ms & 348.75ms & 73.75ms \\
+%8k words & 578.75ms & 580.0ms & 95.0ms \\
+%\hline
+%\end{tabular}
+%}
+%\caption{Comparison of computational times between non-parametric filters (CN32+softmax), spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} (CN32+softmax and $K=5$) and Chebyshev-based filters (CN32+softmax and $K=5$) on 20NEWS.
+We consider a mini-batch size of 100 data, and computations are carried out on an NVIDIA GPU K40c.}
+%\label{tab5}
+%\end{table*}
 
-\subsubsection{Chebyshev
-Parametrization}\label{chebyshev-parametrization}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=5cm]{images/speed.png}
+\caption{Plots of computational times w.r.t.
+the number of selected words in 20NEWS for non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters.
+The architecture is CN32+softmax with $K=5$ for splines and Chebyshev.
+We consider a mini-batch size of 100 data, and computations are carried out on an NVIDIA GPU K40c.}
+\label{fig2}
+\end{figure}
 
-The main idea, to avoid the Fourier basis, is twofold: 1) parametrize
-the filter, in the spectral domain, as a truncated Chebyshev expansion
-and 2) recursively compute the Chebyshev polynomials from the Laplacian.
-This approximate method for spectral graph filtering was first proposed
-in \citep{hammond_wavelets_2011} for a fast wavelet transform on graphs.
+\noindent
+Observe that non-parametric filters are most costly as they require $O(n)$ parameters to learn, and they are unable to learn localized filters, unlike \cite{art:HenaffBrunaLeCun15DLgraphs} and our model.
+This is particularly illustrated on Table \ref{tab6}, where non-parametric filters provide less accurate results.
+Besides, our model outperforms the classification accuracy of \cite{art:HenaffBrunaLeCun15DLgraphs} on MNIST.
 
-Recall that the Chebyshev polynomial \(T_k(x)\) of order \(k\) may be
-generated by the stable recurrence relation
-\(T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)\) with \(T_0 = 1\) and
-\(T_1 = x\). These polynomials form an orthogonal basis for
-\(L^2([-1,1], dy / \sqrt{1-y^2})\), the Hilbert space of square
-integrable functions with respect to the measure \(dy/\sqrt{1-y^2}\).
+\begin{table*}[h!]
+ \centering
+{\small
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+ Architecture & Accuracy  \\
+\hline
+CN10 (Non-Param filters) + softmax & 95.75 \\
+CN10 (Spline filters)  + softmax & 97.26 \\
+CN10 (Chebyshev filters)  + softmax & 97.48 \\
+CN32-P4-CN64-P4-FC512-softmax  (Non-Param filters) & 96.28 \\
+CN32-P4-CN64-P4-FC512-softmax  (Spline filters) & 97.15 \\
+CN32-P4-CN64-P4-FC512-softmax  (Chebyshev filters) & 98.65 \\
+\hline
+\end{tabular}
+}
+\caption{Comparison of precisions between non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters [our model] on MNIST.
+For spline and Chebyshev filters, we picked $K=20$.}
+\label{tab6}
+\end{table*}
 
-The graph filter \(g_\theta(\Lambda)\) can thus be constructed from the
-truncated expansion
-\[ g_\theta(\Lambda) := \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\Lambda}) \]
-of polynomial order \(K-1\), where the parameter \(\theta \in \R^K\),
-\(K \ll M\), is a vector of Chebyshev coefficients and
-\(T_k(\tilde{\Lambda}) \in \R^{M \times M}\) is the Chebyshev polynomial
-of order \(k\) evaluated at
-\(\tilde{\Lambda} := 2 \Lambda / \lambda_{max} - I_M \in \R^{M \times M}\),
-a diagonal matrix of scaled eigenvalues (so that they lie in
-\([-1,1]\)). While reducing the number of parameters from \(M\) to
-\(K\), this parametrization enforces smoothness in the spectral domain,
-which translates to localization in the vertex domain. It can indeed be
-shown that \((\L^k)_{i,j}=0\) if the shortest-path between vertices
-\(v_i\) and \(v_j\) is longer than \(k\) edges
-\citep{hammond_wavelets_2011}, which limits the influence of a
-\(K^\text{th}\) order filter to \(K\) hopes. This is often a desired
-property, e.g.~to learn local features for classification.
+\noindent
+Eventually, we plot in Figure  \ref{fig1} the curves of accuracy and energy loss w.r.t. the number of iterations for the three types of filters.
 
-To avoid the Fourier basis \(U\), we express the polynomials as
-functions of the Laplacian \(\L\) instead of its eigenvalues \(\Lambda\)
-using eq.~\ref{eq:lap_diag}, such that the filtering operator in the
-vertex domain is given by
-\begin{equation} g_\theta(\L) = U g_\theta(\Lambda) U^T =
-\sum_{k=0}^{K-1} U \theta_k T_k(\tilde{\Lambda}) U^T =
-\sum_{k=0}^{K-1} \theta_k T_k(\tilde{\L}), \label{eq:chebyshev}\end{equation}
-where \(T_k(\tilde{\L}) \in \R^{M \times M}\) is the Chebyshev
-polynomial of order \(k\) evaluated at the scaled Laplacian
-\(\tilde{\L} := 2 \L / \lambda_{max} - I_M\). Note that the spectrum of
-the normalized Laplacian is bounded by \(2\)
-\citep{chung_spectral_1997}, such that the scaling can simply be
-\(\tilde{\L} = \L - I_M\), tolerating some imprecision in the
-approximation due to the loss of a fraction \(2-\lambda_{max}\) of the
-domain.
+\begin{figure}[h!]
+\centering
+\subfigure[]{\includegraphics[width=6cm]{images/acc.png}}
+\hspace{0.5cm}
+\subfigure[]{\includegraphics[width=6cm]{images/loss.png}}
+\caption{Plots of accuracy (a) and energy loss (b) w.r.t. the number of iterations.}
+\label{fig1}
+\end{figure}
 
-Inserting eq.~\ref{eq:chebyshev} into eq.~\ref{eq:loss} we obtain
-\begin{equation} L =
-\frac{1}{N} \norm{ \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\L}) X - Y }_F^2 =
-\frac{1}{N} \norm{ \bar{X} \theta - \bar{y} }_2^2, \label{eq:loss_c}\end{equation}
-where \(\bar{y} \in \R^{MN}\) is the vectorized matrix \(Y\) and the
-\(k^\text{th}\) column of \(\bar{X} \in \R^{MN \times K}\) is the
-vectorized matrix
-\(\tilde{X}_k := T_k(\tilde{\L}) X \in \R^{M \times N}\). The gradient
-is then given by \begin{equation} \nabla_\theta L =
-\frac{2}{N} \bar{X}^T (\bar{X} \theta - \bar{y}). \label{eq:gradient_c}\end{equation}
+\section{Conclusion and Future Work}
+\vspace{-0.2cm}
+We have introduced an efficient implementation of CNNs on non-Euclidean manifolds represented by graphs using tools from SPG.
+Experiments have shown the ability of the model to extract the local stationary features through the graph convolutional layers.
+Future works will investigate two directions.
+First, we will explore applications of this generic model to important fields where the data naturally lie on non-artificially constructed graphs such as social networks, biological networks, or telecommunication networks.
+It makes actually more sense to apply this generalized CNN framework to natural network-based data, rather than artificially created networks which quality may vary as seen in the experiments.
+Second, we will also improve this framework with tools developed in SGP, including graph wavelet operators \cite{art:HammondVandergheynstGribonval11GraphWav,art:CoifmanLafon06DifMap,pro:GavishNadlerCoifman10GraphHaar,pro:ChenChengMallat14deepHaar,pro:RustamovGuibas14deepHaar}, spectral graph coarsening techniques \cite{art:ShumanFarajiVandergheynst16PyramTrans}, and improved localization properties \cite{pro:TsitsveroBarbarossa15Uncert,pro:PasdeloupAlamiGriponRabbat15Uncert,art:PerraudinRicaudShumanVandergheynst16Uncert}.
 
-While the system \(\bar{X} \theta - \bar{y}\) is largely over-determined
-as \(K \ll MN\), a closed-form solution of eq.~\ref{eq:loss_c} is given
-by the optimality condition \(\nabla_\theta L = 0\) so that
-\begin{equation} \theta^* = \argmin_\theta L = \bar{X}^+ \bar{y}, \label{eq:direct_c}\end{equation}
-where \(\bar{X}^+ = (\bar{X}^T\bar{X})^{-1} \bar{X}^T\) is the
-pseudo-inverse of \(\bar{X}\). The computation of the inverse is fast
-and stable for small \(K\). Alternative approximate solution methods for
-larger order \(K\) include the computation of the pseudo-inverse with
-\gls{SVD}, the use of a least-square solver or a gradient descent scheme
-based on eq.~\ref{eq:gradient_c}.
+\newpage
 
-Using the recurrence
-\[ \tilde{X}_k = 2\tilde{\L} \tilde{X}_{k-1} - \tilde{X}_{k-2} \] with
-\(\tilde{X}_0 = X\) and \(\tilde{X}_1 = \tilde{\L} X\), the computation
-of \(\bar{X}\) given \(X\) costs \(O(K|E|N) = O(KMN)\) operations if the
-number of edges \(|E|\) is proportional to the number of nodes \(M\),
-e.g.~for kNN graphs. As the cost of the product \(\bar{X}\theta\) is
-similar, the entire filtering operation has a computational cost of
-\(O(KMN)\) whereas the straightforward implementation using the Fourier
-basis has a cost of \(O(M^2\max(M,N)\). When applying multiple filters
-to the same set of signals, as is the case with \gls{SGD}, one may save
-computations by storing and querying \(\bar{X}\) instead of \(X\), at
-the expense of \(K\) times the memory.
-
-\subsubsection{Lanczos Parametrization}\label{lanczos-parametrization}
-
-Another applicable parametrization is based on the Lanczos algorithm, an
-adaptation of the power iteration to find the largest or smallest
-eigenvalues and corresponding eigenvectors of a linear system. It was
-first introduced for fast graph filtering in
-\citep{susnjara_accelerated_2015}.
-
-The algorithm, described in
-\citep{gallopoulos_efficient_1992, susnjara_accelerated_2015},
-constructs an orthonormal basis
-\(V = [v_0, \ldots, v_{K-1}] \in \R^{M \times K}\) of the Krylov
-subspace \(\mathcal{K}_K(\L,x) = \spn\{ x, \L x, \ldots, \L^{K-1} x \}\)
-and a tri-diagonal matrix \(H = V^T \L V \in \R^{K \times K}\) with a
-computational cost of \(O(K |E|)\). Note that for large
-\(K \gtrapprox 30\), the original iterative algorithm may loose the
-basis orthogonality such that a necessary orthogonalization step will
-increase the complexity \citep{susnjara_accelerated_2015}. Filtering the
-signal \(x\) with \(g_\theta(\L)\) can then be approximated by an order
-\(K-1\) polynomial as \begin{equation} y = g_\theta(\L) x \approx
-V g_\theta(H) V^T x = V Q g_\theta(\Sigma) Q^T V^T x \label{eq:lanczos}\end{equation}
-where \(Q \Sigma Q^T = H\) is the eigendecomposition of \(H\). There
-exist fast methods for the eigendecomposition of symmetric tri-diagonal
-matrices {[}{]}. As for the Chebyshev approximation, this construction
-enforces smoothness in the spectral domain.
-
-Inserting eq.~\ref{eq:lanczos} with a parametrized filter
-\(g_\theta(\Sigma) := \diag(\theta)\), \(\theta \in \R^K\), \(K \ll M\),
-into eq.~\ref{eq:loss} gives \[ L = \frac{1}{N} \sum_{n=0}^{N-1}
-\norm{V_n Q_n \diag(\theta) Q_n^T V_n^T x_n - y_n}_2^2 =
-\frac{1}{N} \norm{\diag(\theta) \hat{X} - \hat{Y}}_F^2 \] where
-\(\hat{X} = [\hat{x}_0, \ldots, \hat{x}_{N-1}] \in \R^{K \times N}\),
-\(\hat{Y} = [\hat{y}_0, \ldots, \hat{y}_{N-1}] \in \R^{K \times N}\) and
-\(\hat{x}_n = Q_n^T V_n^T x_n \in \R^K\),
-\(\hat{y}_n = Q_n^T V_n^T y_n \in \R^K\) are approximate representations
-of the signals \(x_n\), \(y_n\) in the orthonormal basis
-\(V_n Q_n \in \R^{M \times K}\) of \(\mathcal{K}_K(\L,x_n)\). Similarly
-to eq.~\ref{eq:loss}, the gradient is given by eq.~\ref{eq:gradient} and
-a closed-form solution by eq.~\ref{eq:direct}.
-
-The expression eq.~\ref{eq:lanczos} can be simplified by setting
-\(v_0 := x / \norm{x}_2\) (the first basis vector can be set to an
-arbitrary unit length vector) such that \(V^T x = \norm{x}_2 e_1\) and
-\[ V g_\theta(H) V^T x =
-\norm{x}_2 V Q \diag(\theta) Q^T e_1 =
-\norm{x}_2 V Q \diag(q) \theta \] where \(e_1 \in \R^K\) denotes the
-first unit vector and \(q = Q^T e_1 \in \R^K\) is the first row of
-\(Q\). Inserting into eq.~\ref{eq:loss} gives \[ L =
-\frac{1}{N} \sum_{n=0}^{N-1} \norm{\tilde{X}_n \theta - y_n}_2^2 =
-\frac{1}{N} \sum_{k=0}^{K-1} \norm{\bar{X}_{\cdot,k} \theta_k - \bar{y}}_2^2 =
-\frac{1}{N} \norm{\bar{X} \theta - \bar{y} }_2^2 \] where
-\(\bar{y} \in \R^{NM}\) is the vectorized matrix \(Y\) and
-\(\bar{X} := [\tilde{X}_0, \ldots, \tilde{X}_{N-1}]^T \in \R^{NM \times K}\)
-is a stack of \(N\) matrices
-\(\tilde{X}_n := \norm{x_n}_2 V_n Q_n \diag(q_n) \in \R^{M \times K}\)
-where \(V_n\), \(Q_n\) and \(q_n\) are derived from \(x_n\). The second
-form (in terms of independent coefficients \(\theta_k\)) is valid
-because \(\bar{X}\) is orthogonal, i.e \(\bar{X}^T \bar{X}\) is a
-diagonal matrix, so that the solution eq.~\ref{eq:direct_c} can be
-written as \begin{equation} \theta^* = \argmin_\theta L =
-\underbrace{\sum_{n=0}^{N-1} \Big( \norm{x_n}_2 \diag(q_n) \Big)^{-2}}
-_{(\bar{X}^T\bar{X})^{-1}} \bar{X}^T \bar{y} \label{eq:direct_l}\end{equation}
-where
-\(\bar{X}^T \bar{y} = [\langle \bar{X}_{\cdot,0}, \bar{y} \rangle, \ldots, \langle \bar{X}_{\cdot,K-1}, \bar{y} \rangle]^T \in \R^K\)
-is the projection of \(\bar{y}\) onto the \(K\) basis vectors
-\(\bar{X}_{\cdot,k} \in \R^{MN}\). While this closed-form evaluation is
-fast and stable, a gradient descent scheme can be used with
-eq.~\ref{eq:gradient_c}.
-
-While the time and space complexities are similar, this method has two
-advantages over the Chebyshev approximation: 1) it does not require the
-normalization of the Laplacian spectrum (thus the estimation of
-\(\lambda_{max}\)) and 2) the optimization is easier as the parameters
-\(\theta_k\) are independent of each other thanks to the orthogonal
-basis. By easier we mean that 1) the closed-form solution
-eq.~\ref{eq:direct_l} does not involve the inversion of a large matrix
-and 2) \gls{SGD} converges to a better solution with less iterations.
-See \citep{susnjara_accelerated_2015} for a discussion of the
-approximation quality and a comparison with the Chebyshev approximation.
-
-\subsection{Graph coarsening}\label{graph-coarsening}
-
-The graph partitioning problem has been widely studied. Applications
-include VLSI design, load balancing for parallel computations, network
-analysis, and optimal scheduling. The goal is to partition the vertices
-of a graph into a certain number of disjoint sets of approximately the
-same size so that a cut metric is minimized. This problem is NP-complete
-even for several restricted classes of graphs, and there is no constant
-factor approximation algorithm for general graphs {[}Bui and Jones
-1992{]}. While notable developments in exact algorithms and heuristics
-have been done, only the introduction of the general-purpose multilevel
-methods during the 1990s has provided a breakthrough in efficiency and
-quality {[}Safro 2009 2012{]}. Graph coarsening, where the problem
-instance is gradually mapped to smaller ones to reduce the original
-complexity, i.e., the graph underlying the problem is compressed, is the
-first step of these multilevel methods.
-
-Existing multilevel algorithms for combinatorial optimization problems
-(such as k-partitioning, linear ordering, clustering, and segmentation)
-can be divided into two classes: contraction-based schemes and algebraic
-multigrid (AMG)-inspired schemes.
-
-Alternative standard algorithms for graph coarsening include the Kron
-reduction {[}{]}, Label Propagation {[}{]}, spectral clustering {[}{]}
-or a sampling {[}Nath uncertainty{]}. A standard approach for graph
-coarsening is the maximal match, as used by METIS for graph partitioning
-{[}?{]}.
-
-The assumption behind METIS is that a (near) optimal partitioning on a
-coarser graph implies a good partitioning in the finer graph. In
-general, it only holds true when the degree of nodes in the graph is
-bounded by a constant {[}Karypis 95{]}. Some real-life graphs,
-e.g.~social networks or WWW, are right-skewed, i.e.~there exists hub
-vertices which have very large degrees. These graphs follow a power-law
-degree distribution.
-
-\section{Applications}\label{applications}
-
-\subsection{Regression}\label{regression}
-
-Solving a linear regression problem on graph signals is akin to minimize
-eq.~\ref{eq:loss} where \(Y\) is the dependent variable and \(X\) the
-independent observations.
-
-\subsection{Representation learning}\label{representation-learning}
-
-Representation learning is the problem of finding an appropriate (to
-some criteria) representation \(y_i\) of a sample \(x_i\) defined as
-\[ \min_{A,Y} \norm{ AX - Y }_F^2 + \tau P(Y) \] where
-\(A \in \R^{M \times M}\) is some operator and \(P(Y)\) is some prior on
-\(Y\). Take \(P(Z)=\norm{Z}_F^2\) and you get a Thikhonov prior. Take
-\(P(Y) = \norm{Y}_1\) and you get sparse coding (with dictionary
-learning).
-
-\subsection{Classification}\label{classification}
-
-There are dependent vectors
-\(X = \{x_i\}_{i=0}^{N-1} \in \R^{M \times N}\) we want to predict from
-observations \(Y = \{y_i\}_{i=0}^{N-1} \in \R^{M \times N}\).
-
-In the classification case, \(y_i \in \N\) is a discrete number which
-represents a class, e.g. \(y_i=-1\) indicates that sample \(x_i\)
-belongs to class 1 and \(y_i=+1\) to class 2.
-
-\subsubsection{Representation learning}\label{representation-learning-1}
-
-Three ways to reduce the number of parameters:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Graph coarsening
-\item
-  No weights: loss of spatiality (convolutional only)
-\item
-  Non-linear activation function
-\end{enumerate}
-
-But \(\theta\) and \(w\) are redundant, so
-\[ L = \norm{\overline{\bar{X} \theta}^T 1_{FM} - y}_2^2 \]
-\[ \bar{X} \in \R^{NM \times K} \] \[ \theta \in \R^{K \times F} \]
-
-\[ \nabla_\theta L = \bar{X}^T (\bar{X} \theta - z) \]
-\[ \nabla_\theta L = (\overline{\bar{X} \theta}^T 1_{FM} - y) \]
-
-\[ \bar{X} 1_{FM} (\overline{\bar{X} \theta}^T 1_{FM} - y) \]
-\[ (\bar{X} 1_{FM} \overline{\bar{X}}^T 1_{FM})^{-1} \bar{X} 1_{FM} y \]
-
-\subsubsection{Linear classification}\label{linear-classification}
-
-Empirical risk minimization (statistical learning).
-
-The \(\ell_2\) regularized least-square model, also known as ridge
-regression, is defined as \begin{equation} L =
-\frac{1}{N} \sum_{i=0}^{N-1} |f_w(x_i) - y_i|^2 + \tau \norm{f}_\H^2,
-\label{eq:rls}\end{equation} where \(f_w \in \H\) is some classification
-function parametrized by \(w\) and \(\H\) is an Hilbert space.
-
-Introducing the linear classifier \(f(x) = w^T x\) into eq.~\ref{eq:rls}
-gives \[ L = \frac{1}{N} \norm{X^T w - y}_2^2  + \tau \norm{w}_2^2, \]
-where \(w \in \R^M\) are the weights to learn. Note that there is no
-bias because the data can be centered around zero (we know the lower and
-upper bounds on luminosity values).
-
-Gradient
-\begin{equation} \nabla_w L = \frac{2}{N} X (X^T w - y) + 2 \tau w \label{eq:gradient_rls}\end{equation}
-
-Optimality condition \(\nabla L = 0\)
-\begin{equation} w^* = \argmin_w \nabla L =
-\left(X X^T + \tau N I_M \right)^{-1} X y \label{eq:sol_rls}\end{equation}
-
-\subsubsection{Feature learning}\label{feature-learning}
-
-Introducing graph filter learning as a scheme for feature learning gives
-the objective \[ L = \frac{1}{N} \sum_{i=0}^{N-1}
-\Big| f_w \Big( A_\theta(\L) x_i \Big) - y_i \Big|^2 +
-\tau_R \norm{f_w}_\H^2, \] where the filter bank
-\(A_\theta(\L) = \bigcup_{i=0}^{F-1} g_{\theta_i}(\L) \in \R^{FM \times M}\)
-is a block-diagonal matrix composed of \(F\) filters
-\(g_{\theta_i}(\L)\) which extract different features.
-
-Using a linear classifier and the Lanczos parametrization introduced in
-\ref{lanczos-parametrization} gives \begin{equation} L =
-\frac{1}{N} \norm{\overline{\bar{X} \theta}^T w - y}_2^2 +
-\tau_R \norm{w}_2^2, \label{eq:rls_gfl}\end{equation} where
-\(w \in \R^{FM}\) are the weights to learn,
-\(\bar{X} \in \R^{NM \times K}\) is the Lanczos basis described in
-\ref{lanczos-parametrization}, \(\theta \in \R^{K \times F}\) are the
-\(K\) coefficients of the \(F\) filters and
-\(\overline{\bar{X} \theta} \in \R^{FM \times N}\) is a rearrangement of
-the matrix \(\bar{X} \theta \in \R^{NM \times F}\). Note that there is
-now \(F\) times more weigts to learn, i.e. \(w \in \R^{FM}\).
-
-\(\bar{X} \theta \in \R^{NM \times F}\)
-
-\[ \frac{2}{N} \bar{X} w (\overline{\bar{X} \theta}^T w - y) \]
-
-\[ (\bar{X} w \overline{\bar{X}}^T w)^{-1} \bar{X}wy \]
-
-\subsubsection{Splitting scheme}\label{splitting-scheme}
-
-The splitting scheme \[ L =
-\frac{1}{N} \norm{Z^T w - y}_2^2 +
-\frac{\tau_F}{N} \norm{\overline{\bar{X} \theta} - Z}_F^2 +
-\tau_R \norm{w}_2^2 \] may be used to minimize the non-convex objective
-eq.~\ref{eq:rls_gfl} for \(\theta\) and \(w\), where
-\(Z = [z_0, \ldots, z_{N-1}] \in \R^{FM \times N}\) are the signals in
-feature space. Solving the overall non-convex problem requires to
-iteratively solve the following three convex sub-problems:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  The gradient and solution to \(\min_\theta L\) are given by
-  eq.~\ref{eq:gradient_c} and eq.~\ref{eq:direct_l}, substituting
-  \(\bar{y}\) by \(\bar{Z} \in \R^{NM \times F}\), a rearrangement of
-  \(Z\).
-\item
-  The gradient of \(\min_Z L\) is \[ \nabla_Z L =
-  \frac{2}{N} w (w^T Z - y^T) + \frac{2\tau_F}{N} (Z - \overline{\bar{X} \theta}) \]
-  and the closed-form solution is \[ Z^* = \argmin_Z L =
-  \left( \tau_F I_{FM} + w w^T \right)^{-1}
-  \left( \tau_F \overline{\bar{X} \theta} + w y^T \right). \]
-\item
-  The gradient and solution to \(\min_w L\) are given by
-  eq.~\ref{eq:gradient_rls} and eq.~\ref{eq:sol_rls}, substituting \(X\)
-  by \(Z\).
-\end{enumerate}
-
-\subsubsection{Graph coarsening}\label{graph-coarsening-1}
-
-To reduce again the number of parameters in \(w\), the graph should be
-coarsened to trade feature resolution at the price of vertex resolution.
-
-\subsubsection{Kernel methods}\label{kernel-methods}
-
-The same can be done for kernel methods.
-
-We want to learn a classification function
-\(f(x) = \sum_i \K(x,x_i) \alpha_i \in \R\), where \(\K(\cdot,\cdot)\)
-is some kernel function, which maps a sample \(x\) to a label \(y\). The
-loss function to minimize over the training set is then given by
-\[ L = \sum_i |f(x_i) - y_i|^2 + \tau \norm{f}_{H_\K}^2 =
-\norm{ \K \alpha - y }_2^2 + \tau \alpha^T \K \alpha, \] where
-\(\K_{i,j} = \K(x_i, x_j)\) is the kernel function evaluated at all
-training samples and \(\alpha \in \R^N\) are the parameters.
-
-\[ \nabla_\alpha L = \K (\K \alpha - y) + \tau \K \alpha \] \(\K\) is
-symmetric.
-
-\[ \alpha^* = \argmin_\alpha L = (\K + \tau I_N)^{-1} y \]
-
-\subsubsection{Feature learning}\label{feature-learning-1}
-
-A general formulation of the loss for a classification problem is given
-by \[ \min_{\theta_A, \theta_h} \left\{ L = \frac{1}{N} \sum_{i=0}^{N-1}
-\left| h_{\theta_h}(A_{\theta_A}(\L) x_i) - y_i \right|^2 +
-\tau_A \Omega_A(A_{\theta_A}) + \tau_h \Omega_h(h_{\theta_h}) \right\} \]
-where \(A\) is an operator which transforms the input signals into a
-suitable, i.e.~learned, representation and \(h\) is a classification
-function. \(\Omega_A\) and \(\Omega_h\) are regularizations, \(\tau_A\)
-and \(\tau_h\) are hyper-parameters.
-
-The SVM classification function \(h \in H_\K\) is defined as
-\[ h_{\theta_h}(x) = \sum_{i=0}^{N-1} \alpha_i \K(x, x_i) + \beta,
-\ \theta_h = (\alpha_i, \beta), \] where \(\K(\cdot,\cdot)\) is some
-kernel function. It is regularized by
-\[ \Omega_h = \norm{h}_{H_\K}^2. \]
-
-In our graph filters learning setting, the mapping
-\(A \in \R^{FM \times M}\) is given by the filter bank
-\[ A_{\theta_A}(\L) = \bigcup_{f=0}^{F-1} g_{\theta_f}(\L),
-\ \theta_A = (\theta_f), \] a block-diagonal matrix composed of \(F\)
-filters, which extracts multiple features from the input signal.
-
-While the optimization complexity would increase a lot if we were to
-minimize for the model complexity \(K\), we can easily include a penalty
-of the form
-\[ \Omega_A = \sum_f \sum_i |\theta_f(\lambda_i) w(\lambda_i)|^2 \]
-where \(w(\lambda_i) \sim 1 / \lambda_i\).
-
-Learning complexity: there is \(FK\) parameters in \(\theta_A\) and
-\(N+1\) parameters in \(\theta_h\).
-
-\subsection{Going deep}\label{going-deep}
-
-Define convolutional neural networks on irregular domains by learning
-multiple layers of graph filters.
-
-Image of the architecture.
-
-\section{Experiments}\label{experiments}
-
-\subsection{Filter learning}\label{filter-learning}
-
-Example from the Juypiter notebook.
-
-Ground truth filter. where \(c_\ell = g(\lambda_\ell)\) is the evaluated
-filter.
-
-(randomly generated) graph signals and a filter \(g(\lambda)\) for some
-\(t\), we want to learn the filter coefficients \(C\), a diagonal
-matrix, such that \[\min_C \sum_{i=1}^N \norm{ U^T C U x_i - y_i }_2^2\]
-where \(U\) are the eigenvectors of the graph Laplacian \(L\) and
-\(y_i = x_i * g + \varepsilon, \varepsilon \sim \mathcal{N}(0,\epsilon)\)
-is the target signal (the filtered signal \(x_i\) with additive Gaussian
-noise).
-
-\subsubsection{Non-parametrized filter}\label{non-parametrized-filter}
-
-\subsubsection{Spline Parametrization}\label{spline-parametrization}
-
-The authors of \citep{bruna_spectral_2013, henaff_deep_2015} learned
-graph filters parametrized by cubic splines.
-
-\subsubsection{Chebyshev
-Parametrization}\label{chebyshev-parametrization-1}
-
-\subsubsection{Lanczos Parametrization}\label{lanczos-parametrization-1}
-
-\subsection{Classification}\label{classification-1}
-
-\subsubsection{Sampled MNIST}\label{sampled-mnist}
-
-\subsubsection{Text documents}\label{text-documents}
-
-\section{Discussion}\label{discussion}
-
-\section{Conclusion and Outlook}\label{conclusion-and-outlook}
-
-That was awesome.
-
-\subsubsection*{Acknowledgments}
-
-Thanks ...
-
-\section*{References}
+%\vspace{-0.15cm} in .bbl
+\bibliographystyle{plain}
+{\setstretch{0}
+{\small
+%\bibliography{bib_nips16}
+\bibliography{bib_nips16_short}
+}
+}
 

### 2016-10-28 nips2016/nips2016.tex
nips v4: small edition by Xavier (2016-05-24)

@@
 Its greedy rule consists, at any given coarsening level, in matching two nodes that locally maximizes a balanced cut energy like normalized cut \cite{art:ShiMalik00NCut}, which are excellent clustering energies.
-Precisely, we randomly pick an unmarked vertex $i$ and match it with one of its unmarked neighbors $j$ that maximizes the local normalized cut $\frac{W_{ij}}{d_i}+\frac{W_{ij}}{d_j}$.
+Precisely, we follow \cite{art:BressonHuLaurentSzlamvonBrecht14Reseed} by randomly picking an unmarked vertex $i$ and match it with one of its unmarked neighbors $j$ that maximizes the local normalized cut $\frac{W_{ij}}{d_i}+\frac{W_{ij}}{d_j}$.
 The two matched vertices are then marked.
@@
 \hline
- Algorithms & Accuracy  \\
+ Algorithm & Accuracy  \\
 \hline
@@
 \hline
- Algorithms & Word2vec features  \\
+ Algorithm & Word2vec features  \\
 \hline
@@
 We investigate here 5 graphs of words: a graph of learned embedding of the corpus using word2vec \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of normalized word counts across each document,
-a graph of pre-trained embedding of words \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of approximate nearest neighbors (ANN) using LSHForest on the learned embedding \cite{pro:MikolovChenCorradoDean13word2vec}, and a random graph.
+a graph of pre-trained embedding of words \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of approximate nearest neighbors (ANN) using LSHForest \cite{pro:BawaCondieGanesan05LSHForest} on the learned embedding \cite{pro:MikolovChenCorradoDean13word2vec}, and a random graph.
 The classification results are reported in Table \ref{tab4b}.
@@
 $\bullet$  No eigenvalue decomposition required: The spectral graph CNNs \cite{art:HenaffBrunaLeCun15DLgraphs} require to compute a full eigenvalue decomposition (EVD) to compute the Fourier basis.
-This has a complexity of $0(n^3)$ with $n=10^{4}$, which is memory expensive.
+This has a complexity of $O(n^3)$ with $n=10^{4}$, which is memory expensive.
 Our technique does not require an EVD, it only requires multiplications between the graph Laplacian sparse matrix and vectors.
@@
 \caption{Plots of computational times w.r.t.
-the number of selected words in 20NEWS for non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters.
+$n$, the number of selected words in 20NEWS, for non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters [our model].
 The architecture is CN32+softmax with $K=5$ for splines and Chebyshev.
@@
 \noindent
-Observe that non-parametric filters are most costly as they require $O(n)$ parameters to learn, and they are unable to learn localized filters, unlike \cite{art:HenaffBrunaLeCun15DLgraphs} and our model.
+Observe that non-parametric filters are costly as they require $O(n)$ parameters to learn, and they are unable to learn localized filters, unlike \cite{art:HenaffBrunaLeCun15DLgraphs} and our model.
 This is particularly illustrated on Table \ref{tab6}, where non-parametric filters provide less accurate results.
@@
 }
-\caption{Comparison of precisions between non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters [our model] on MNIST.
+\caption{Comparison of precisions between non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters on MNIST.
 For spline and Chebyshev filters, we picked $K=20$.}
@@
 \subfigure[]{\includegraphics[width=6cm]{images/loss.png}}
-\caption{Plots of accuracy (a) and energy loss (b) w.r.t. the number of iterations.}
+\caption{Plots of accuracy (a) and energy loss (b) w.r.t. the number of iterations for the first 2000 iterations.}
 \label{fig1}
@@
 {\small
-%\bibliography{bib_nips16}
-\bibliography{bib_nips16_short}
+\bibliography{refs}
+%\bibliography{bib_nips16_short}
 }

### 2016-10-29 nips2016/nips2016.tex
nips v5: large edition by Michaël (2016-06-17)

--- 2016-10-28 nips2016/nips2016.tex
@@
 
-\usepackage{nips2016}
-%\usepackage[final]{nips_2016}
+%\usepackage{nips_2016}
+\usepackage[final]{nips2016}
 
@@
 \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
+\usepackage[english]{babel}
+\usepackage[colorlinks=true,allcolors=blue]{hyperref}       % hyperlinks
+%\usepackage{url}            % simple URL typesetting
+%\usepackage{booktabs}       % professional-quality tables
 \usepackage{amsfonts}       % blackboard math symbols
-\usepackage{amssymb}
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
+%\usepackage{amssymb}
+\usepackage{xfrac}       % compact symbols for 1/2, etc.
+%\usepackage{microtype}      % microtypography
 \usepackage{subfigure}
 \usepackage{graphicx}
+%\usepackage{tabularx}
+\usepackage{booktabs}
+%\usepackage{multirow}
+%\usepackage{svg}
 \usepackage{setspace}
@@
 
-\title{Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering}
+% Use less space
+%\usepackage{titlesec}		% space before / after headings
+%\titlespacing\section{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
+%\titlespacing\subsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
+%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
+%\usepackage{parskip} \setlength{\parskip}{5pt}
+\setlength{\bibsep}{0.2pt}
 
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
+\title{Convolutional Neural Networks on Graphs\\
+with Fast Localized Spectral Filtering}
 
@@
   %École Polytechnique Fédérale de Lausanne (EPFL) \\
-  EPFL \\
-  Lausanne, Switzerland \\
+  EPFL, Lausanne, Switzerland \\
   \texttt{michael.defferrard@epfl.ch} \\
@@
   Xavier Bresson \\
-  EPFL \\
-  Lausanne, Switzerland \\
+  EPFL, Lausanne, Switzerland \\
   \texttt{xavier.bresson@epfl.ch} \\
-  \And
+  \And % \AND
   Pierre Vandergheynst \\
-  EPFL \\
-  Lausanne, Switzerland \\
+  EPFL, Lausanne, Switzerland \\
   \texttt{pierre.vandergheynst@epfl.ch} \\
@@
 \usepackage[acronym]{glossaries}
+\newacronym{CNN}{CNN}{Convolutional Neural Network}
 \newacronym{SVD}{SVD}{Singular Value Decomposition}
@@
 \DeclareMathOperator*{\spn}{span}
-\renewcommand{\L}{\mathcal{L}}
 \newcommand{\G}{\mathcal{G}}
-\renewcommand{\H}{\mathcal{H}}
-\newcommand{\K}{\mathcal{K}}
 \newcommand{\V}{\mathcal{V}}
 \newcommand{\E}{\mathcal{E}}
-\newcommand{\W}{\mathcal{W}}
+\newcommand{\bO}{\mathcal{O}}
 \newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Xh}{\hat{X}}
-\newcommand{\Yh}{\hat{Y}}
-\newcommand{\st}{\ \text{s.t.} \,}
-\newcommand{\norm}[1]{\left\| #1 \right\|}
+\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
+\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
+\newcommand{\secref}[1]{Section~\ref{sec:#1}}
+\newcommand{\eqnref}[1]{(\ref{eq:#1})}
+\newcommand{\bruna}{art:BrunaZarembaSzlamLeCun13DLgraphs,
+art:HenaffBrunaLeCun15DLgraphs}
+
+\newcommand{\todo}[1]{{\color{red} #1 }}
 
 \begin{document}
-% \nipsfinalcopy is no longer used
 
@@
 \begin{abstract}
-Convolutional neural networks have greatly improved state-of-the-art performances in computer vision and speech analysis tasks, due to its high ability to extract multiple levels of representations of data.
-In this work, we are interested in generalizing convolutional neural networks from low-dimensional regular grids, where image, video and speech are represented, to high-dimensional irregular domains, such as social networks, biological graphs like gene regulatory and brain connectivity networks, telecommunication networks, or words' embedding.
-We present a formulation of convolutional neural networks on graphs in the context of the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design fast localized convolutional filters on graphs.
-Numerical experiments on MNIST and 20NEWS demonstrate the ability of the system to learn local stationary features on graphs, as long as the graph is well-constructed.
+
+	Convolutional neural networks (CNNs) have greatly improved state-of-the-art
+	performances in a number of fields, notably computer vision and natural
+	language processing. In this work, we are interested in generalizing the
+	formulation of CNNs from low-dimensional regular Euclidean domains, where
+	image (2D), video (3D) and audio (1D) are represented, to high-dimensional
+	irregular domains represented by graphs. This paper introduces a formulation
+	of CNNs on graphs in the context of spectral graph theory. We borrow the
+	fundamental tools from the emerging field of signal processing on graphs,
+	which provides the necessary mathematical background and efficient numerical
+	schemes to design localized graph filters which are fast to evaluate.
+	Similarly to classical CNNs, the computational complexity as well as the
+	number of parameters to learn are independent of the data dimensionality.
+	Numerical experiments on MNIST and 20NEWS demonstrate the ability of the
+	model to learn local and stationary features on graphs, as long as the graph
+	is well-constructed. This work can be thought as the foundation of a second
+	generation of CNNs on graphs built on spectral graph theory, being
+	mathematically well defined, bringing better test performance and being
+	faster to evaluate and train than the first generation introduced by Bruna,
+	LeCun et al.
+
+	%there is nothing to lose by going to the graph domain
+	%except the rotation invariance
+
+	%Graphs can either be the natural domain of the data, such as social networks,
+	%biological graphs like gene regulatory and brain connectivity networks,
+	%telecommunication networks, or constructed from the data, such as similarity
+	%graphs.
+
+	%not a generalization because it provides more invariance, but we don't want
+	%to say it.
+	%Our formulation is not a direct generalization of classical CNNs, in the
+	%sense that it is not a spatial construction.
 \end{abstract}
@@
 \centering
-\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.eps}
+\includegraphics[width=\textwidth]{images/illustrationCNNgraphs.eps}
 %\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
-\caption{Architecture of the proposed CNNs on graphs.
-Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at layer $l$, $G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
+\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
+coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
 \label{fig:}
@@
 \section{Introduction}
-\vspace{-0.2cm}
-Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST} offer an efficient architecture to extract highly meaningful statistical patterns in large-scale and high-dimensional datasets.
-The key success of CNNs is its advanced ability to learn local stationary structures and compose them to form multi-scale hierarchical patterns.
-Precisely, CNNs extract the local stationarity property of the input data or signals by revealing local features that are shared across the data domain.
-These similar features are identified with localized convolutional filters or kernels, which are learned from the data.
-Convolutional filters are shift- or translation-invariant filters, meaning they are able to recognize identical features independently of their spatial locations.
-Localized kernels or compactly supported filters refer to filters that extract local features independently of the input data size, with a support size that can be much smaller than the input size.
-\\
-The compositional local stationarity property of CNNs can be efficiently implemented on regular grids, like image and sound domains, as convolutional and downsampling operators are mathematically well-defined on such discretized Euclidean spaces.
-This has led to the breakthrough of CNNs in image, video, and sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on low-dimensional regular lattices.
-But not all data lie on regular grids.
-User data on social networks, gene data on biological regulatory networks, log data on telecommunication networks, or text documents on words' embedding are important examples of data lying on irregular or non-Euclidean domains.
-Unfortunately, CNNs cannot be directly applied to such complex domains.\\
-A generalization of CNNs to irregular grids is not direct as the basic convolutional and downsampling operators, necessary to extract compositional local stationarity features, are not mathematically clear on these domains.
-This makes this extension mathematically challenging, both in theory and in implementation.
-The goal of this work is to leverage the recent works of \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs,art:HammondVandergheynstGribonval11GraphWav,art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG,art:ShumanRicaudVandergheynst16Local} to introduce an efficient implementation of CNNs on irregular domains represented here as graphs.
-Graphs are universal representations of heterogeneous pairwise relationships between possibly high-dimensional data.
-Graphs can encode complex geometric structures, and can be studied with strong mathematical tools such as graph spectral theory, a blend between graph theory and harmonic analysis.\\
-The major bottleneck of generalizing CNNs to graphs is the design of fast and localized filters on graphs, that are the core elements to extract local stationarity structures.
-Our major goal is to overcome this barrier.
-We will make use of recent tools developed in the context of signal processing on graphs (SPG) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to precisely control the spatial support of graph filters using the so-called polynomial kernel localization \cite{art:ShumanRicaudVandergheynst16Local}, and to perform an efficient backward propagation technique using the Chebyshev technique \cite{art:HenaffBrunaLeCun15DLgraphs}.
-Eventually, a generic and efficient implementation of CNNs on graphs may pave the way to the development of a potentially large class of deep learning techniques for graph-based data.
 
-\section{Related Works}
-\vspace{-0.2cm}
-{\bf CNNs on non-Euclidean domains.} Extending the success of CNNs to non-Euclidean domains is essential to boost the power analysis of data lying on complex networks like biological, social or telecommunication networks.
-This is a recent line of works, with a preliminary approach coming from the local reception fields in \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF}, and applied successfully to object recognition in computer vision.
-The main idea is to unsupervisedly learn groups of similar features.
-This is highly useful to automatically choose a limited number of connections between two successive layers.
-This approach is however not able to extract similar groups over the data domain, as it is not clear how to define convolutional filters on these groups.
-\\
-In \cite{art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, the authors proposed a generalization of CNNs on graphs using graph counterparts of grid convolution and downsampling.
-They were able to learn shift-invariant or convolutional filters on graphs using the convolution theorem \cite{book:Mallat99wavelets} that defines convolutional operators as linear operators that diagonalize in the Fourier basis (represented by the eigenvectors of the Laplacian operator).
-They also designed a strategy to learn the graph structure from the data, and apply their model on image recognition, text categorization and bioinformatics.
-This promising approach is however limited in implementation as forward and backward evaluations have complexity $O(n^2)$ due to the matrix multiplication operations that operate the graph Fourier transforms.
-Besides, there is no strong control over the support size of the convolutional kernels, which is essential to design localized filters to extract local stationary features, independent of the input size.
-\\
-The authors in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,art:MasciBoscainiBronsteinVandergheynst15ShapeNet} introduced ShapeNet, a generalization of CNNs to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces.
-They used geodesic polar coordinates to define mesh patches and formulated a deep learning architecture.
-They obtained state-of-the-art results for shape recognition tasks.\\
-Finally, the papers \cite{pro:ChenChengMallat14deepHaar,pro:RustamovGuibas14deepHaar} investigated the construction Haar wavelet transforms on graphs using a deep hierarchical architecture.
-They applied these deep wavelets on graphs to object recognition on sphere, and to sparse reconstruction of faces.
-\\
-{\bf Signal Processing on Graphs.} SPG is an emerging field that aims at bridging the gap between signal processing techniques like wavelet analysis and graph theory such as spectral graph analysis \cite{art:BelkinNiyogi05LaplaBeltrami,art:VonLuxburg07Tutorial}.
-The main goal is to generalize fundamental analysis operations for signals from regular grids to irregular structures embodying by graphs.
-We refer the reader to \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} for an introduction of the field.
-Standard operations on grids such as convolution, translation, filtering, dilatation, modulation, downsampling do not extend directly to graphs and thus require new mathematical definitions while keeping the original intuitive concepts.
-In this context, the authors of \cite{art:HammondVandergheynstGribonval11GraphWav,art:CoifmanLafon06DifMap,pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet operators on graphs.
-In \cite{art:ShumanFarajiVandergheynst16PyramTrans,art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform mutli-scale pyramid transforms on graphs.
-The works of \cite{pro:TsitsveroBarbarossa15Uncert,pro:PasdeloupAlamiGriponRabbat15Uncert,art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles on graphs, and showed that intuitive concepts may be lost, but can also produce enhanced localization principles for signals on graphs.
-In \cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to carry out lasso-based signal regularization on graphs, and studied the intertwined relationships between smoothness and sparsity on graphs.
-In \cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors investigated compressed sensing recovery conditions for graph spectral signal analysis.
-\\
-Beyond this work, our long-term goal is to make use of SPG tools as much as possible to formulate mathematically well-designed techniques of CNNs on graphs.
-This paper introduces the mathematical and computational foundations of such models, while future developments may consider enhanced graph filters like graph wavelet operators, spectral graph coarsening techniques, and improved localization properties.
+Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST}
+offer an efficient architecture to extract highly meaningful statistical
+patterns in large-scale and high-dimensional datasets. The key success of CNNs
+is its advanced ability to learn local stationary structures and compose them to
+form multi-scale hierarchical patterns. Precisely, CNNs extract the local
+stationarity property of the input data or signals by revealing local features
+that are shared across the data domain. These similar features are identified
+with localized convolutional filters or kernels, which are learned from the
+data. Convolutional filters are shift- or translation-invariant filters, meaning
+they are able to recognize identical features independently of their spatial
+locations. Localized kernels or compactly supported filters refer to filters
+that extract local features independently of the input data size, with a support
+size that can be much smaller than the input size.
 
-\section{Contributions}
-\vspace{-0.2cm}
-{\textbf {\textit {(1)}}} Learning deep fully connected networks is almost computationally untractable for large data domain.
-Fortunately, for data satisfying stationarity and hierarchical compositionality structure, CNNs have proven to be remarkably powerful to identify highly informative features using low learning complexity.
-For example, learning a single layer with $n$ input coordinates and $m<n$ outputs require $O(n^2)$ complexity to learn fully connected networks.
-It also requires the complexity $O(n)$ to learn any arbitrary filter, and $O(S)$ for localized filters, where $S$ is the size of the filter support, independent of the input size $n$.
-One of our main contributions is to stay with the same low complexity for learning graph convolutional filters.
-We will show it is actually possible to keep exactly the same complexity as standard CNNs, i.e.
-$O(S)$ for localized graph filters.\\
-{\textbf {\textit {(2)}}} The backpropagation learning algorithm requires to run convolutional operations, which can be efficiently implemented by FFT on grids with complexity $O(n\log n)$.
-Convolution on graphs can also be implemented by matrix multiplications representing the forward and backward graph Fourier transforms, but it is expensive as the complexity is $O(n^2)$.
-As noticed in \cite{art:HenaffBrunaLeCun15DLgraphs}, this complexity is a strong bottleneck to deal with large-scale problems.
-We will greatly improve this complexity by reducing to $O(E.S)$, where $E$ is the number of edges in the given graph, using the Chebyshev technique from SPG \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
-Besides, as most real-world graphs are highly sparse, we have $E\ll n^2$, and therefore a low complexity for the matrix multiplication operations.
-Eventually, the Chebyshev computations can be naturally distributed on GPUs, as we did in our experiments.\\
-{\textbf {\textit {(3)}}} We introduce an efficient pooling strategy for CNNs on graphs, that also makes use of distributed computing architecture like GPUs.
-The pooling approach consists on two stages.
-First, a graph coarsening procedure based on METIS \cite{art:DhillonGuanKulis07Graclus,art:KarypisKumar98Metis} is successively applied to the original graph to produce coarser versions of the graph.
-Second, at each coarsening level, we sequentially order all nodes such that two successive indexed nodes are the parent nodes of the child node in the next coarser level.
-This node indexation structure in memory satisfies parallel architectures.
-\\
-{\textbf {\textit {(4)}}} We regard this proposed work as the funding paper that bridges the gap between two powerful fields, namely deep learning and signal processing on graphs.
-In the future, we expect to further benefit from tools in SPG to enhance the formulation of CNNs on graphs.
-\\
-{\textbf {\textit {(5)}}} We revisit the Euclidean case in the numerical experiments with the standard MNIST image classification problem.
-The model is able to learn localized convolutional filters and achieve 98.65\% of success rate.
-More interestingly, we apply our model to the text categorization problem on 20NEWS.
-We also demonstrate that our techniques is able to extract local stationary features, as it improves by 2.5\% the classification accuracy when no convolutional are present.
-We observe that the quality of the graph construction is strongly linked to the quality of the convolutional layers.
-Finally, we demonstrate the computational performance of our model over \cite{art:HenaffBrunaLeCun15DLgraphs}.
+The compositional local stationarity property of CNNs can be efficiently
+implemented on regular grids, like image and sound domains, as convolutional and
+downsampling operators are mathematically well-defined on such discretized
+Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
+sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on
+low-dimensional regular lattices. But not all data lie on regular grids. User
+data on social networks, gene data on biological regulatory networks, log data
+on telecommunication networks, or text documents on words' embedding are
+important examples of data lying on irregular or non-Euclidean domains.
+Unfortunately, CNNs cannot be directly applied to such complex domains.
 
-\section{Learning Localized Graph Filters}
-\vspace{-0.2cm}
-In this section, we explore how to design efficient localized convolutional filters on graphs.
+A generalization to irregular grids is not direct as the building blocks of
+CNNs, the convolution and downsampling operators, are not mathematically clear
+on these domains. This makes this extension challenging, both theoretically and
+implementation-wise. This work leverages the recent works of
+\cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
+art:BrunaZarembaSzlamLeCun13DLgraphs, art:HenaffBrunaLeCun15DLgraphs,
+art:HammondVandergheynstGribonval11GraphWav,
+art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to introduce an
+efficient implementation of CNNs on irregular domains represented here as
+graphs. Graphs are universal representations of heterogeneous pairwise
+relationships between possibly high-dimensional data. Graphs can encode complex
+geometric structures, and can be studied with strong mathematical tools such as
+spectral graph theory, a blend between graph theory and harmonic analysis.
 
-\subsection{Signal Processing on Graphs and Localized Filters}
-\vspace{-0.2cm}
-{\bf Graph Fourier Transform.} We are interested in processing signals defined on undirected and connected graphs $G=(V,E,W)$, where $V$ is the set of vertices with $|V|=n$, $E$ is the set of edges and $W$ is the weighted adjacency matrix encoding the connection weights between two vertices.
-A signal $x:V\rightarrow\mathbb{R}$ defined on the nodes of the graph may be regarded as a vector $x\in\mathbb{R}^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node.
-An essential operator in spectral graph analysis is the graph Laplacian \cite{book:Chung97Spectral}, which unnormalized definition is $L=D-W\in\mathbb{R}^{n\times n}$ with $D\in\mathbb{R}^{n\times n}$ is the diagonal degree matrix where $D_{ii}=\sum_jW_{ij}$, and its normalized definition is $L=I_n-D^{-1/2}WD^{-1/2}$ where $I_n$ is the identity matrix.
-As $L$ is a real symmetric positive semidefinite matrix, it has a complete set of orthonormal eigenvectors $\{u_l\}_{l=0}^{n-1}\in\mathbb{R}^n$, known as the graph Fourier modes, and their associated real nonnegative eigenvalues $\{\lambda_l\}_{l=0}^{n-1}$, which are the frequencies of the graph system.
-Equivalently, linear operator $L$ diagonalizes the Fourier basis matrix $U=[u_0,...,u_{n-1}]\in\mathbb{R}^{n\times n}$ such that $L=U\Lambda U^T$ where $\Lambda=\diag(\lambda_l)$.
-We can now define the graph Fourier transform (GFT) and its inverse \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
-The GFT of a spatial signal $x\in\mathbb{R}^n$ is $\hat{x}=U^Tx\in\mathbb{R}^n$, and the inverse GFT is $x=U\hat{x}$.
-Using GFT we can now generate fundamental operations of signal processing on graphs such as convolution and filtering.
-\\
-{\bf Convolutional Filters.} The graph convolutional operator is simply defined in the Fourier space such as $f\ast_G g=U((U^Tf) \odot (U^Tg))$, where $\odot$ the element-wise Hadamard product.
-Then the convolutional filtering operation on graphs is defined as $y=g_\theta(L)x\in\mathbb{R}^n$, where $g_\theta(L)=g_\theta(U\Lambda U^T)=Ug_\theta(\Lambda) U^T\in\mathbb{R}^{n\times n}$ where $g_\theta(\Lambda)=\diag(\theta)$ and $\theta\in\mathbb{R}^n$ is the filter expressed in the Fourier space.
-Remember that convolutional filters, i.e.
-translation-invariant filters on graphs, can only be defined in the Fourier domain as convolution operators are not defined in the vertex space.
-\\
-{\bf Localized Convolutional Filters.} Learning convolutional filters on graphs boils down to learn spectral multipliers $\theta$ of complexity $O(n)$, the size of the input data.
-Our goal is to reduce this complexity to $O(S)$, the support size of the filter in the vertex space, exactly as in standard CNNs.
-For this, we will make use of a property called polynomial localization \cite{art:ShumanRicaudVandergheynst16Local}.
-Let us represent $\theta=\{\theta_{\{{\lambda_l}\}_{l=0}^n}\}\in\mathbb{R}^n$ as a $K$-th differentiable polynomial function: $\theta_{\lambda_l}^K(a)=\sum_{k=0}^{k-1}a_k \lambda_l^k$, where $a=\{a_k\}_{k=0}^{K-1}\in\mathbb{R}^K$.
-The polynomial localization theorem \cite{art:ShumanRicaudVandergheynst16Local} shows that $(T_ig_{\theta^K})_j=(\delta_i \ast_G (Ug_{\theta^K}))_j=0$ for $d_G(i,j)>K$ where $d_G$ is the shortest distance on graph, meaning that the support of the associated spatial filter is exactly $K$-localized, all values are zero after $K$ hopes.
-Consequently, spectral filters represented by $K$-order polynomials are spatially localized with support $S=K$, implying the low learning complexity $O(K)$ for each filter.
-We notice that the authors \cite{art:HenaffBrunaLeCun15DLgraphs} also used a $K$-order polynomials represented by splines, but they did not formalize the quantitive relationship between filter support size and polynomial order.
-Finally, it is also possible to estimate the spatial decaying profile of the filter \cite{art:ShumanRicaudVandergheynst16Local}: for any K-localized filter centered at vertex $i$, we have the profile $|(T_ig_{\theta^K})(j)|\leq\big[\frac{2\sqrt{n}}{K!}(\frac{\lambda_{max}}{4})^K \sup_{\lambda}|(Ug_{\theta^K})^K(\lambda)|\big]$, which may be useful for graph filter analysis.
+The major bottleneck of generalizing CNNs to graphs, and the primary goal of
+this work, is the definition of localized filters which are efficient to
+evaluate and learn.
+%that are the core elements to extract local stationarity structures.
+We will make use of recent tools developed in the context of graph signal
+processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
+to precisely control the spatial support of graph filters as well as to allow
+fast evaluations. Eventually, a generic and efficient implementation of CNNs for
+graph signals may pave the way to the development of a potentially large class
+of deep learning techniques for graph-based data.
 
-\subsection{Fast Filter Learning with Chebyshev Technique}
-\vspace{-0.2cm}
-Let $y$ be a signal $x$ filtered on the graph: $y=g_\theta(L)x=U(g_\theta(\Lambda) (U^Tx))$.
-Learning a spectral filter can be reduced to a complexity $O(K)$ by polynomial parametrization.
-However, the numerical computation of $y$ still requires two full matrix multiplications by $U^T,U$ with a high complexity of $O(n^2)$, and these matrix multiplications must be done several times in the backpropagation process.
-This is a computational bottleneck, identified in \cite{art:HenaffBrunaLeCun15DLgraphs}, with an additional cost of $O(n^3)$ for computing a full eigenvalue decomposition of $L$ to get $U$.
-We can overcome this computational issue using a Chebyshev technique as proposed in SPG \cite{art:HammondVandergheynstGribonval11GraphWav}.
-We will also explore in the future the Lanczos technique \cite{art:SusnjaraPerraudinKressnerVandergheynst15Lanczos}, which seems attractive because of the independence of the coefficients.
-\\
-We first recall the fundamental recurrence property of Chebyshev polynomials.
-Let the Chebyshev polynomial \(T_k(x)\) of order \(k\) be generated by the recurrence relation \(T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)\) with \(T_0 = 1\) and
-\(T_1 = x\).
-These polynomials form an orthogonal basis.
-Then, the K-compact spectral filter $g_{\theta^K}(\Lambda)$ can be defined as $g_{\theta^K}(\Lambda)=\sum_{k=0}^{K-1}\theta_k T_k(\tilde{\Lambda})$ of order \(k\) evaluated at
-\(\tilde{\Lambda} := 2 \Lambda / \lambda_{max} - I_n \in \R^{n \times n}\), a diagonal matrix of scaled eigenvalues such that they lie in \([-1,1]\).
-Consequently, the filtering operation is defined as $y=g_{\theta^K}(L)x=Ug_{\theta^K}(\Lambda) U^Tx=\sum_{k=0}^{K-1} U \theta_k T_k(\tilde{\Lambda}) U^Tx=\sum_{k=0}^{K-1} \theta_k T_k(\tilde{L}) x$, where \(T_k(\tilde{L}) \in \R^{n \times n}\) is the Chebyshev polynomial of order \(k\) evaluated at the scaled Laplacian
-\(\tilde{L} := 2 L / \lambda_{max} - I_n\).
-If we denote $X_k:=T_k(\tilde{\Lambda})x$ and using the recurrence equation, we can generate $X_k=2\tilde{L}X_{k-1}-X_{k-2}$ with \(X_0 = x\) and \(X_1 = \tilde{L} x\).
-As $L$ is sparse, it is straightforward to see that all matrix multiplications are done between sparse matrices, greatly decreasing the computational complexity for graph filtering from $O(n^2)$ to $O(E.K)$.
-Finally, matrix multiplications can be efficiently carried out in a distributive way, allowing an efficient GPU implementation of Chebyshev filtering.
+% Why convolution? Weight sharing, subject to assumptions
+%\paragraph{Low learning complexity.} Learning deep fully connected networks is
+%almost computationally intractable for large data domain.  Fortunately, for
+%data satisfying stationarity and hierarchical compositionality structure, CNNs
+%have proven to be remarkably powerful to identify highly informative features
+%using low learning complexity. For example, learning a single layer with $n$
+%input coordinates and $m<n$ outputs require $\bO(n^2)$ complexity to learn fully
+%connected networks. It also requires the complexity $\bO(n)$ to learn any
+%arbitrary filter, and $\bO(S)$ for localized filters, where $S$ is the size of
+%the filter support, independent of the input size $n$. One of our main
+%contributions is to stay with the same low complexity for learning graph
+%convolutional filters. We will show it is actually possible to keep exactly the
+%same complexity as standard CNNs, i.e. $\bO(S)$ for localized graph filters.
 
-\section{Graph Coarsening and Fast Pooling}
-\vspace{-0.2cm}
-Local stationarity property of data is extracted via localized convolutional kernels.
-In this section, we are interested to extract the multi-scale hierarchical composition property of data.
-This is efficiently achieved via grid subsampling or downsampling and pooling, which trades spatial resolution with feature resolution reducing the learning complexity without compromising the system performances.
-In the case of standard CNNs for data lying on regular grids, the subsampling operation is mathematically sound.
-The translation operator commutes with the downsampling operator, hence not affecting the translation invariance property of the system at different scales of grid resolutions.\\
-How to reduce grid size in the case of graphs?
-We need to construct meaningful neighbourhoods on graphs where similar features may be clustered together.
-This naturally boils down to compute a multi-scale clustering of the graph that preserves local geometric structures.
-Unfortunately, it is well-known that graph clustering is a NP-hard problem \cite{art:BuiJonesGraphPartNPhard}, and approximations must be explored.
-Several clustering techniques exist like spectral clustering \cite{art:VonLuxburg07Tutorial}.
-However, we are more interested in clustering techniques that can reduce the size of the graph by a factor 2, which offers a precise control on the coarsening process.
-This is essential to keep control on the reduction factor to be able to extract fine hierarchical compositional structures.
-Algebraic multigrid techniques on graphs \cite{art:RonSafroBrandt11MultigridGraph} can be used in this context.\\
-In this work, we make use of the downsampling or coarsening technique introduced in METIS \cite{art:DhillonGuanKulis07Graclus,art:KarypisKumar98Metis} which has shown to be extremely efficient at clustering a large variety of graphs.
-METIS uses a greedy algorithm to compute successive coarser versions of a given graph.
-Its greedy rule consists, at any given coarsening level, in matching two nodes that locally maximizes a balanced cut energy like normalized cut \cite{art:ShiMalik00NCut}, which are excellent clustering energies.
-Precisely, we randomly pick an unmarked vertex $i$ and match it with one of its unmarked neighbors $j$ that maximizes the local normalized cut $\frac{W_{ij}}{d_i}+\frac{W_{ij}}{d_j}$.
-The two matched vertices are then marked.
-The matching is repeated until all nodes have been explored.
-This is an extremely fast coarsening technique, which divides by almost a factor 2 the size of the nodes from one level to the next coarser level.
-The coarsening factor is not exactly as there exist a few singletons, non-matched nodes.
-We can however control the reduction factor and set it to exactly 2 by simply adding artificial disconnected nodes to the singletons.
-We then order the nodes such that two successive indexed nodes are the parent nodes of the child node in the next coarser level.
-The major reason to do this is to benefit from parallel hardware architectures like GPUs to considerably speed up the pooling operations, which must be done several times during backprogation.\\
-In a future work, we will explore the use of SPG by using a Kron reduction technique \cite{art:ShumanFarajiVandergheynst16PyramTrans}.
-The Kron graph coarsening technique has the main advantage to preserve the ordering of the Laplacian spectrum, and may be able to commute with the graph convolution operator.
-If proven, this would be a very essential mathematical property of CNNs on graphs, exactly as regular grid downsampling in standard CNNs.
+\todo{spatial approaches provide localization by definition}
+\todo{why spectral approach?}
+While a spatial approach is conceivable, it faces the challenge of matching
+local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.
 
-%\newpage
+\todo{Another challenge is the definition of a pooling layer.}
+Local stationarity property of data is extracted via localized convolutional
+kernels. In this section, we are interested to extract the multi-scale
+hierarchical composition property of data. This is efficiently achieved via grid
+subsampling or downsampling and pooling, which trades spatial resolution with
+feature resolution reducing the learning complexity without compromising the
+system performances. In the case of standard CNNs for data lying on regular
+grids, the subsampling operation is mathematically sound. The translation
+operator commutes with the downsampling operator, hence not affecting the
+translation invariance property of the system at different scales of grid
+resolutions.
+
+\paragraph{Graph Signal Processing.} GSP is an emerging field that aims at
+bridging the gap between signal processing techniques like wavelet analysis and
+graph theory such as spectral graph analysis
+\cite{art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}. A goal is to
+generalize fundamental analysis operations for signals from regular grids to
+irregular structures embodied by graphs. We refer the reader to
+\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
+\todo{others for politics} for an
+introduction of the field. Standard operations on grids such as convolution,
+translation, filtering, dilatation, modulation or downsampling do not extend
+directly to graphs and thus require new mathematical definitions while keeping
+the original intuitive concepts. In this context, the authors of
+\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
+pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
+operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,
+art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform
+mutli-scale pyramid transforms on graphs. The works of
+\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
+art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles
+on graphs, and showed that intuitive concepts may be lost, but can also produce
+enhanced localization principles for signals on graphs. In
+\cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to
+carry out lasso-based signal regularization on graphs, and studied the
+intertwined relationships between smoothness and sparsity on graphs. In
+\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
+investigated compressed sensing recovery conditions for graph spectral signal
+analysis.
+
+\todo{Model statistical assumptions: locality, stationarity, compositionality.}
+These assumptions can be experimentally verified for any dataset. While we know
+for them to be true for audio, images and videos, we show experimentally that
+they might be satisfied for text documents as long as the graph is properly
+constructed.
+We also observe that
+local stationarity is an assumption that is made about the data, and the CNNs
+are fundamentally designed on this data hypothesis. However, we believe this
+stationarity assumption to be present in many real-world data, as structured
+data are usually composed of repetitive small patterns at different scales of
+observation.
+
+\section{Related Work}
+
+%Extending the success of CNNs to
+%non-Euclidean domains is essential to boost the power analysis of data lying on
+%complex networks like biological, social or telecommunication networks.
+
+Extending CNNs to non-Euclidean domains is a recent line of work. A first
+attempt is the so-called local reception fields \cite{pro:GregorLeCun10LRF,
+pro:CoatesNg11LRF}, successfully applied to image recognition. The main idea is
+to group together features based upon a measure of similarity such as to select
+a limited number of connections between two successive layers. While this model
+reduces the number of parameters by exploiting the locality assumption, it does
+not attempt to exploit any weight-sharing strategy.
+%stationarity This approach is however not able to extract similar groups over
+%the data domain, as it is not clear how to define convolutional filters on
+%these groups.
+\todo{read papers}
+
+ShapeNet \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
+art:MasciBoscainiBronsteinVandergheynst15ShapeNet} is a generalization of CNNs
+to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. The
+authors used geodesic polar coordinates to define mesh patches and formulated a
+deep learning architecture which allows comparison across different manifolds.
+They obtained state-of-the-art results for shape recognition tasks.
+
+In \cite{pro:ChenChengMallat14deepHaar} and \cite{pro:RustamovGuibas14deepHaar},
+the authors investigated the construction of Haar wavelet transforms on graphs
+using a deep hierarchical architecture. They applied the method to object
+recognition on sphere, and to sparse reconstruction of faces.
+\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
+they impose a shape to the filters.}
+\todo{Unknown graph geometry?}
+
+Finally, \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} and
+\cite{art:HenaffBrunaLeCun15DLgraphs} introduced a generalization of CNNs on
+graphs using graph counterparts of grid convolution and downsampling. They were
+able to learn convolutional filters on graphs using the convolution theorem
+\cite{book:Mallat99wavelets} that defines convolutions as linear operators that
+diagonalize in the Fourier basis (represented by the eigenvectors of the
+Laplacian operator). They were thus the first to introduce a spectral
+formulation to extend CNNs on graphs. They also introduced a strategy to learn
+the graph structure from the data, and applied their model on image recognition,
+text categorization and bioinformatics. This approach does however not scale due
+to the necessary multiplications by the graph Fourier basis, a matrix of $n^2$
+coefficients, where $n$ is the data dimensionality. Despite the cost of
+computing this matrix, which requires an EVD on the graph Laplacian, the
+dominant cost is the need to multiply the data by this matrix twice (forward and
+inverse Fourier transforms) at a cost of $\bO(n^2)$ operations per forward and
+backward pass, a computational bottleneck already identified by the authors.
+Besides, as they rely on smoothness (via a spline parametrization) in the
+Fourier domain to bring localization in the vertex domain, their model does not
+provide a precise control over the local support of their kernels, which is
+essential to learn localized filters. This work directly builds on theirs, and
+we will show how we overcame these limitations and what we further gained.
+
+\subsection{Contributions}
+
+Our contributions can be summed up as follows:
+\begin{enumerate}
+%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
+\item \textbf{Spectral formulation.} A clean spectral formulation built on
+	established tools in GSP.
+\item \textbf{Strictly localized filters.} Unlike \cite{\bruna}, the proposed spectral
+	filters are strictly localized in a ball of radius $K$, i.e. $K$ hops from
+	the central vertex.
+\item \textbf{Low computational complexity.} The evaluation of our filters is
+	linear w.r.t. the filters support's size $K$ and the number of edges $n \leq
+	|\E| \leq n^2$. As most real-world graphs are highly sparse, we have $|\E|
+	\ll n^2$ and $|\E| = kn$ for the widespread $k$-neighbors graphs. Moreover,
+	this method avoids the Fourier basis altogether, thus the expensive
+	eigenvalue decomposition (EVD) necessary to compute it as well as the need
+	to store the basis, a matrix of size $n^2$. That is especially relevant when
+	working with limited GPU memory. Besides the data, our method only requires
+	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
+\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
+	graphs which, after a rearrangement of the vertices, is analog to pooling of
+	1D signals.
+\item \textbf{Experimental results.} We present many experiments that ultimately
+	show that our formulation is (1) a useful model, (2) computationally
+	efficient and (3) superior both in accuracy and complexity to the previous
+	generation of spectral graph CNNs introduced in \cite{\bruna}. We also show that
+	our formulation performs similarly as a classical CNN on MNIST and study the
+	impact of various graph constructions on performance.
+\end{enumerate}
+
+Beyond this work, our long-term goal is to make use of GSP tools as much as
+possible to formulate mathematically well-designed techniques for CNNs on
+graphs. This paper introduces the mathematical and computational foundations of
+such models, while future developments may consider to enhance the building
+blocks, namely graph filtering and coarsening. In the future, we expect to
+further benefit from newly developed tools in GSP to enhance our formulation of
+CNNs on graphs.
+
+\section{Learning Fast Localized Spectral Filters} \label{sec:filters}
+
+As there is no notion of spatial translation on graphs, convolutions are
+implemented as multiplications in the spectral domain
+\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. While a spatial
+convolution on an Euclidean domain is localized by definition (via the finite
+size of the kernel) and has a linear computational complexity $\bO(n)$, a filter
+defined in the spectral domain is not localized by default and costs $\bO(n^2)$
+operations because of the necessary multiplication by the graph Fourier basis.
+Note that while classical CNNs can be speed-up by leveraging the Fourier basis
+(instead of computing the convolutions in the spatial domain)
+\cite{mathieu_fast_2013}, it is only advantageous because of the $\bO(n \log(n))$
+complexity of the FFT, which has no equivalent on graphs (yet). A good choice of
+filter parametrization however allows us to prove that when translated (via a
+convolution with a delta) to a given center vertex, the filter is strictly
+localized in a ball of radius $K$ around that center vertex. Moreover, a
+recursive formulation allows the filtering operation to scale linearly with $n$.
+
+\paragraph{Graph Fourier Transform.} We are interested in processing signals
+defined on undirected and connected graphs $\G=(\V,\E,W)$, where $\V$ is a
+finite set of vertices with $|\V|=n$, $\E$ is a set of edges and $W$ is a
+weighted adjacency matrix encoding the connection weight between two vertices. A
+signal $x: \V \rightarrow \R$ defined on the nodes of the graph may be regarded
+as a vector $x \in \R^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node.
+An essential operator in spectral graph analysis is the graph Laplacian
+\cite{book:Chung97Spectral}, which combinatorial definition is $L = D - W \in
+\R^{n \times n}$ where $D \in \R^{n \times n}$ is the diagonal degree matrix
+with $D_{ii} = \sum_j W_{ij}$, and its normalized definition is $L = I_n -
+D^{-1/2} W D^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real
+symmetric positive semidefinite matrix, it has a complete set of orthonormal
+eigenvectors $\{u_l\}_{l=0}^{n-1} \in \R^n$, known as the graph Fourier modes,
+and their associated ordered real nonnegative eigenvalues
+$\{\lambda_l\}_{l=0}^{n-1}$, identified as the frequencies of the graph. The
+Laplacian is indeed diagonalized by the Fourier basis $U=[u_0, \ldots, u_{n-1}]
+\in \R^{n \times n}$ such that $L = U \Lambda U^T$ where $\Lambda =
+\diag([\lambda_0, \ldots, \lambda_{n-1}]) \in \R^{n \times n}$. We can now
+define the graph Fourier transform (GFT) of a spatial signal $x \in \R^n$ as
+$\hat{x} = U^T x \in \R^n$, and its inverse as $x = U \hat{x}$
+\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. Similarly to the
+Fourier transform on Euclidean spaces, the GFT enables the formulation of
+fundamental operations such as filtering.
+
+\paragraph{Spectral filtering of graph signals.} As we cannot express a
+meaningful translation operator in the vertex domain, the convolution operator
+on graph $\ast_\G$ is defined in the Fourier domain such that $x \,\ast_\G\, y =
+U((U^Tx) \odot (U^Ty))$, where $\odot$ is the element-wise Hadamard product. It
+follows that a signal $x$ is filtered by $g_\theta$ as
+\begin{equation}
+	y = g_\theta(L) x = g_\theta(U \Lambda U^T) x = U g_\theta(\Lambda) U^T x.
+\end{equation}
+A non-parametric filter, i.e. a filter whose parameters
+are all free, would be defined as
+\begin{equation} \label{eq:filt_non-param}
+	g_\theta(\Lambda) = \diag(\theta),
+\end{equation}
+where the parameter $\theta \in \R^n$ is a vector of Fourier coefficients.
+
+\paragraph{Polynomial parametrization for localized filters.} There are however
+two problems with non-parametric filters: (1) they are not localized and (2)
+they are of learning complexity $\bO(n)$, the dimensionality of the data. We are
+going to solve those two problems together by reducing this complexity to
+$\bO(K)$, the support's size of the filter, exactly as it is for standard CNNs.
+While \cite{\bruna} solves it by parameterizing $g_\theta(\Lambda)$ as a cubic
+B-spline defined by $K$ control points, let us parametrize it as a polynomial
+function
+\begin{equation} \label{eq:filt_poly}
+	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k \Lambda^k,
+\end{equation}
+where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
+The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is given
+by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
+(L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
+delta $\delta_i \in \R^n$. By \cite[Lemma
+5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
+$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the distance
+between two vertices is the minimum number of edges in any path connecting them.
+Consequently, spectral filters represented by $K^\text{th}$-order polynomials of
+the Laplacian are exactly $K$-localized.
+
+% While there is many possible choice of recursive polynomials
+\paragraph{Recursive formulation for fast filtering.} While we have shown how to
+learn localized filters with $K$ parameters, the cost to filter a signal $x$ as
+$y = Ug_\theta(\Lambda)U^Tx$ is still of $\bO(n^2)$ operations because of the
+multiplications with the Fourier basis $U$. A way out is to parametrize
+$g_\theta(L)$ as a polynomial function that can be computed recursively from
+$L$, as $K$ multiplications by a sparse $L$ costs $\bO(K|\E|) \ll \bO(n^2)$. One
+such polynomial, traditionally used in GSP to approximate kernels (like
+wavelets), is the Chebyshev expansion
+\cite{art:HammondVandergheynstGribonval11GraphWav}. Another option, the Lanczos
+algorithm \cite{art:SusnjaraPerraudinKressnerVandergheynst15Lanczos}, which
+constructs an orthonormal basis of the Krylov subspace $\mathcal{K}_K(L,x) =
+\spn\{x,Lx,\ldots,L^{K-1}x\}$, seems attractive because of the coefficients'
+independence. It is however more convoluted and thus left as a future work.
+
+Recall that the Chebyshev polynomial $T_k(x)$ of order $k$ may be generated by
+the stable recurrence relation $T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)$ with $T_0 =
+1$ and $T_1 = x$. These polynomials form an orthogonal basis for $L^2([-1,1], dy
+/ \sqrt{1-y^2})$, the Hilbert space of square integrable functions with respect
+to the measure $dy/\sqrt{1-y^2}$. A filter can thus be parametrized as a
+truncated expansion of order $K-1$ such that
+\begin{equation} \label{eq:filt_cheby}
+	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k T_k(\tilde{\Lambda}),
+\end{equation}
+where the parameter $\theta \in \R^K$ is a vector of Chebyshev coefficients and
+$T_k(\tilde{\Lambda}) \in \R^{n \times n}$ is the Chebyshev polynomial of order
+$k$ evaluated at $\tilde{\Lambda} = 2 \Lambda / \lambda_{max} - I_n$, a diagonal
+matrix of scaled eigenvalues that lie in $[-1,1]$.
+
+The filtering operation can then be written as $y = g_\theta(L) x
+= \sum_{k=0}^{K-1} \theta_k T_k(\tilde{L}) x$, where $T_k(\tilde{L}) \in \R^{n
+\times n}$ is the Chebyshev polynomial of order $k$ evaluated at the scaled
+Laplacian $\tilde{L} = 2 L / \lambda_{max} - I_n$. Note that the spectrum of the
+normalized Laplacian is bounded by $2$ \cite{book:Chung97Spectral}, such that
+the scaling can simply be $\tilde{L} = L - I_n$. Denoting $\bar{x}_k =
+T_k(\tilde{L})x \in \R^n$, we can use the recurrence relation to compute
+$\bar{x}_k = 2\tilde{L} \bar{x}_{k-1} - \bar{x}_{k-2}$ with $\bar{x}_0 = x$ and
+$\bar{x}_1 = \tilde{L}x$. The entire filtering operation
+\begin{equation}
+	y = g_\theta(L) x = [\bar{x}_0, \ldots, \bar{x}_{K-1}] \theta
+\end{equation}
+then costs $\bO(K|\E|))$ operations.
+
+\paragraph{Learning filters.} The $j^\text{th}$ output feature map of the sample
+$s$ is given by
+\begin{equation} \label{eq:filterbank}
+	y_{s,j} = \sum_{i=1}^{F_{in}} g_{\theta_{i,j}}(L) x_{s,i} \in \R^n,
+\end{equation}
+where the $x_{s,i}$ are the input feature maps and the $F_{in} \times F_{out}$ vectors of Chebyshev
+coefficients $\theta_{i,j} \in \R^K$ are the layer's trainable parameters. When
+training multiple convolutional layers with the backpropagation algorithm, one
+needs the two gradients
+\begin{align}
+	\frac{\partial E}{\partial \theta_{i,j}} =
+	\sum_{s=1}^S [\bar{x}_{s,i,0}, \ldots, \bar{x}_{s,i,K-1}]^T
+	\frac{\partial E}{\partial y_{s,j}}
+	&& \text{and} &&
+	\frac{\partial E}{\partial x_{s,i}} =
+	\sum_{j=1}^{F_{out}} g_{\theta_{i,j}}(L)
+	\frac{\partial E}{\partial y_{s,j}},
+\end{align}
+where $E$ is the loss energy over a mini-batch of $S$ samples. Each of the
+above three computations boils down to $K$ sparse matrix-vector multiplications
+and one dense matrix-vector multiplication for a cost of $\bO(K |\E| F_{in}
+F_{out} S)$ operations. These can be efficiently computed on parallel
+architectures by leveraging tensor operations. As an optimization,
+$[\bar{x}_{s,i,0}, \ldots, \bar{x}_{s,i,K-1}]$ can be computed once.
+
+\section{Efficient Pooling on Graphs} \label{sec:pooling}
+
+% Why are we introducing this new pooling definition?
+
+\todo{Why? Bruna uses an ad-hoc technique.}
+
+\paragraph{Graph Coarsening.} In contrast with regular domains, to reduce the
+sampling of a graph signal, we need to construct meaningful neighborhoods on
+graphs where similar vertices may be clustered together. Doing this for multiple
+layers is equivalent to construct a multi-scale clustering of the graph that
+preserves local geometric structures. It is however well-known that graph
+clustering is a NP-hard problem \cite{art:BuiJonesGraphPartNPhard} and that
+approximations must be used. While there exist many clustering techniques (e.g.
+the popular spectral clustering \cite{art:VonLuxburg07Tutorial}), we are most
+interested in multilevel clustering algorithms, where each level produces a
+\textit{coarser} graph which corresponds to the data domain seen at a different
+\textit{resolution}. Moreover, clustering techniques that reduce the size of the
+graph by a factor two at each level offers a precise control on the coarsening
+and pooling size. In this work, we make use of the coarsening phase of the
+Graclus multilevel clustering algorithm \cite{art:DhillonGuanKulis07Graclus},
+which has been shown to be extremely efficient at clustering a large variety of
+graphs. We briefly review it below. Algebraic multigrid techniques on graphs
+\cite{art:RonSafroBrandt11MultigridGraph} and the Kron reduction
+\cite{art:ShumanFarajiVandergheynst16PyramTrans} are two methods worth exploring
+in future works. The later has the main advantage to preserve the ordering of
+the Laplacian spectrum, and may be able to commute with the graph convolution
+operator. \todo{If proven, this would be a very essential mathematical property
+of CNNs on graphs, exactly as downsampling for classical CNNs.}
+
+Graclus \cite{art:DhillonGuanKulis07Graclus}, similarly to METIS, uses a greedy
+algorithm to compute successive coarser versions of a given graph and is able to
+minimize several popular spectral clustering objectives. We chose the normalized
+cut \cite{art:ShiMalik00NCut}, which is an excellent clustering energy.
+Graclus' greedy rule consists, at any given coarsening level, in randomly
+pickling an unmarked vertex $i$ and matching it with one of its unmarked
+neighbors $j$ that maximizes the local normalized cut $W_{ij} (1/d_i + 1/d_j)$.
+The two matched vertices are then marked and the coarsened weights are set as
+the sum of their weights. The matching is repeated until all nodes have been
+explored. This is an extremely fast coarsening scheme which divides the number
+of nodes by approximately two (there exist a few singletons, non-matched nodes)
+from one level to the next coarser level.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{images/coarsening_crop}
+\caption{\textbf{Example of Graph Coarsening and Pooling.} Let's say we want to
+perform a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
+\R^8$ living on $\G_0$, the finest graph given as input. Note that it originally
+possesses $n_0 = |\V_0| = 8$ vertices, arbitrarily ordered. For a pooling of
+size 4, two coarsenings of size 2 are needed: Graclus first gives $\G_1$ of size
+$n_1 = |\V_1| = 5$, then $\G_2$ of size $n_2 = |\V_2| = 3$, the coarsest graph.
+Sizes are thus set to $n_2 = 3$, $n_1 = 6$, $n_0 = 12$ and fake nodes (in blue)
+are added to $\V_1$ (1 node) and $\V_0$ (4 nodes) to pair with the singeltons
+(in orange), such that each node has exactly two children.  Nodes in $\V_2$ are
+then arbitrarily ordered and nodes in $\V_1$ and $\V_0$ are ordered
+consequently.  At that point the arrangement of vertices in $\V_0$ permits a
+regular 1D pooling on $x \in \R^{12}$ such that $z = [\max(x_0,x_1),
+\max(x_4,x_5,x_6), \max(x_8,x_9,x_{10})] \in \R^3$, where the signal components
+$x_2,x_3,x_7,x_{11}$ are set to a neutral value.}
+\label{fig:pooling}
+\end{figure}
+
+\paragraph{Pooling of graph signals.} After coarsening, each node has either two
+parents, if it was matched at the finer level, or one, if it was not, i.e the
+node was a singelton. From the coarsest to finest level, fake nodes, i.e.
+disconnected nodes, are added to pair with the singletons such that each node
+has two children. This structure can be seen as a balanced binary tree: (1)
+regular nodes (and singletons) either have two regular nodes (e.g. level 1
+vertex 0 in \figref{pooling}) or (2) one singleton and a fake node as children
+(e.g. level 2 vertex 0 in \figref{pooling}), and (3) fake nodes always have two
+fake nodes as children (e.g. level 1 vertex 1 in \figref{pooling}). Input
+signals are initialized with a neutral value at the fake nodes, e.g. 0 when
+using a ReLU activation with max pooling. Because these nodes are disconnected,
+filtering won't impact the initial neutral value. While those fake nodes do
+artificially increase the dimensionality thus the computational cost, we found
+that, in practice, the number of singletons left by Graclus is quite low
+% < 10%?
+compared to the number of vertices.
+
+Arbitrarily ordering the nodes at the coarsest level, then propagating this
+ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
+children, produces a regular ordering in the finest level. Regular in the sense
+that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
+rearranged graph signal is analog to pooling a regular 1D signal.
+\figref{pooling} shows an example of the whole coarsening and pooling process.
+This regular arrangement makes the pooling operation very efficient and
+satisfies parallel architectures such as GPUs as memory accesses are local, i.e.
+matched nodes don't have to be fetched.
 
 \section{Numerical Experiments}
-\vspace{-0.4cm}
-We performed the numerical experiments with TensorFlow, an open source software library for deep learning developed by the Google Brain Team, which features various backends, notable CUDA for GPU computing.
 
-\subsection{Revisiting Euclidean CNNs on MNIST}
-\vspace{-0.2cm}
-Our first experiment aims at reproducing the performance of standard CNNs on the highly popular MNIST benchmark dataset \cite{pro:LeCunBottouBengioHaffner98MNIST}.
-This is an important sanity check for our proposed generalized model.
-It should be able to recover the original object classification rate as the convolutional layers are also designed to extract local stationary features on any graph, including the 2D grid graph on which images reside.
-\\
-MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28\times 28$, such that data points lie on a space of $784$ dimensions.
-TensorFlow reports an accuracy of 99.20\% on MNIST with a LeNet5-like CNN architecture.
-Our proposed CNNs on graphs with a graph of $8$-NN (nearest neigbors) of a 2D Euclidean grid achieves a precision of 98.65\%, as reported in Table \ref{tab1}.
+All experiments were performed with TensorFlow, an open-source library for
+numerical computation using data flow graphs, especially suited for deep
+learning \cite{abadi_tensorflow_2016}. It features various backends, notably
+CUDA to compute on Nvidia GPUs. All computations are carried on an Nvidia Tesla
+K40c GPU.
+
+In the sequel, we refer to filters defined by \eqnref{filt_non-param} as
+\textit{Non-Param} and filters defined by \eqnref{filt_cheby} as
+\textit{Chebyshev}. Filters referred to as \textit{Spline} are defined by
+\begin{equation} \label{eq:filt_spline}
+	g_\theta(\Lambda) = B \theta,
+\end{equation}
+where $B \in \R^{n \times K}$ is the cubic B-spline basis and the parameter
+$\theta \in \R^K$ is a vector of control points, as proposed by \cite{\bruna}. It is
+important to note that, for all methods, we used the Graclus state-of-the-art
+coarsening algorithm introduced in \secref{pooling} rather than their "naive
+agglomerative method", as they put it. The rational is that we wanted to compare
+the filters, not the coarsening algorithms.
+
+We use the following notation when describing network architectures: FC$k$
+denotes a fully connected layer with $k$ hidden units, P$k$ denotes a (graph or
+classical) pooling layer of size and stride $k$, C$k$ denotes a convolutional
+layer with $k$ feature maps and GC$k$ denotes a graph convolutional layer with
+$F_{out} = k$ feature maps (with Chebyshev filters if not specified otherwise).
+All FC$k$, C$k$ and GC$k$ layers are followed by a ReLU activation $\max(x,0)$.
+The final layer is always a softmax regression and the loss energy $E$ is the
+cross-entropy loss \todo{with an $\ell_2$ regularization on the weights of all
+FC$k$ layers}. Mini-batches are of size $S = 100$.  \todo{dropout}
+
+\todo{$k$-NN similarity graphs are built as follows: }
+
+%\subsection{Revisiting classical CNNs on MNIST}
+\subsection{Comparison with classical CNNs on MNIST}
+% Struct: motivation, dataset, model, results
+
+To confirm the validity of our model, we first revisit the Euclidean
+case, where the graph is a regular grid, on the standard MNIST classification
+problem \cite{pro:LeCunBottouBengioHaffner98MNIST}.
+This is an important sanity check for our model, which should be able to extract
+features on any graph, including the 2D grid graph on which images reside.
+\tabref{mnist} indeed shows that our model is able to learn localized filters
+and achieve a performance very close to that of a classical CNN with the same
+architecture.
 
@@
 \centering
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
- Algorithms & Accuracy  \\
-\hline
-Linear SVM & 91.76  \\
-Softmax & 92.36  \\
-CNNs [LeNet5-like] & 99.20  \\
-graph CNNs: CN32-P4-CN64-P4-FC512-softmax & 98.65 \\
-\hline
+\begin{tabular}{llc}
+\toprule
+Model & Architecture & Accuracy \\
+\midrule
+%Linear SVM & 91.76  \\
+%Softmax & 92.36  \\
+Classical CNN & C32-P4-C64-P4-FC512 & \textbf{99.20} \todo{99.33} \\
+Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 98.65 \todo{99.14} \\
+\bottomrule
 \end{tabular}
-}
-\caption{Classification results on MNIST.} 
-\label{tab1}
+\caption{Classification accuracies for the proposed model and a classical CNN on
+MNIST.} 
+\label{tab:mnist}
 \end{table*}
 
-\noindent
-We report in Table \ref{tab2} the speedup offered by GPUs.
-A GPU implementation of graph CNNs is 8x faster than a CPU implementation, i.e.
-the same order of speedup than standard CNNs, as it only uses matrix multiplications which are efficiently implemented in CUDA BLAS.
+MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28
+\times 28$, such that data points lie on a space of $784$ dimensions. To test
+our model, we constructed a 8-NN (nearest neighbors) graph of the 2D Euclidean
+grid, which produced a graph of $n = |\V| = 976$ nodes (784 pixels and 192 fake
+nodes) and $|\E| = 3198$ edges.
 
-\begin{table*}[h!]
- \centering
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
- Architecture & Time  & Speedup \\
-\hline
-CNNs with CPU & 210ms & -\\
-CNNs with GPU NVIDIA K40c & 31ms & 6.77x\\
-graph CNNs with CPU & 200ms & -\\
-graph CNNs  with GPU NVIDIA K40c & 25ms & 8.00x\\
-\hline
-\end{tabular}
-}
-\caption{Computational time for processing a mini-batch of 100 images.} 
-\label{tab2}
+The LeNet-5-like network architecture and the following hyper-parameters are
+borrowed from the TensorFlow MNIST tutorial\footnote{
+\url{https://www.tensorflow.org/versions/r0.8/tutorials/mnist/pros}}: dropout
+probability of 0.5, regularization weight of $5\times10^{-4}$, initial learning
+rate of 0.03, learning rate decay of 0.95, momentum of 0.9. Convolutional layer
+have filters of size $5 \times 5$ while graph convolution layers have the same
+support of $K = 5^2 = 25$. All models were trained for \todo{20 epochs}.
+
+\subsection{Text Categorization on 20NEWS}
+
+To demonstrate the usefulness of our model and the versatility of modeling with
+graphs, we applied our model to the text categorization problem on 20NEWS.
+\tabref{20news} indeed shows that CNNs on graphs are a good model which offers
+decent performances. While it does not outperform a simpler algorithm like a
+multinomial naive Bayes classifier on this small dataset, it does defeat fully
+connected networks, which require much more parameters to be learned. More
+importantly, these results show that the statistical assumptions made by the
+model are verified.
+
+\begin{table*}[h!] \centering
+\begin{tabular}{lc} \toprule
+Model & Accuracy \\
+\midrule
+Linear SVM & 65.90 \\
+Multinomial Naive Bayes & 68.51 \\
+%Multinomial logistic regression (softmax) & 66.28 \\
+Softmax & 66.28 \\
+\addlinespace
+FC2500 & 64.64 \\
+FC2500-FC500 & 65.76 \\
+\addlinespace
+GC32 & 68.26 \\
+GC32-FC500 & \todo{?} \\
+\bottomrule \end{tabular}
+\caption{Classification accuracies for the proposed model and a other algorithms
+on 20NEWS.} 
+\label{tab:20news}
 \end{table*}
 
-\noindent
-We now discuss the influence of the graph quality over the performance of the CNNs on graphs model.
-Convolutional layers aims at extracting local stationary features, or repetitive feature patches over the image domain.
-This implies 2 requirements to be successful.
-First, the data must satisfy this property, which we know to be true for MNIST.
-Second, the graph must be well constructed to be able to extract local stationary features.
-In the case of images, a simple $k$-NN graph of the grid is good enough, and the value of $k$ does not have a strong influence on the results.
-However, let us destroy the graph structure by using a random graph, then the accuracy drops to 95.39\%, which is the order of a simple fully connected graph with a softmax classifier, that is when the convolutional layers are not used.
-Table \ref{tab3} reports the results.
+20NEWS consists of 18,846 (11,314 for training and 7,532 for testing) text
+documents associated with 20 classes \cite{art:Joachims9620NEWS}. We extracted
+the 10,000 most common words from the 93,953 unique words in this corpus. Each
+document is represented using the bag-of-words model, normalized across words.
+To test our model, we constructed a 16-NN graph of the word2vec
+\cite{pro:MikolovChenCorradoDean13word2vec} embedding of those words, which
+produced a graph of \todo{$n = |\V| = $ nodes (10k words and ? fake nodes) and
+$|\E| = ?$ edges}. The influence of the graph quality is further studied in
+\secref{graph_quality}. All models were trained for \todo{20 epochs} by the Adam
+optimizer \todo{cite} with a learning rate of 0.1, without regularization nor
+dropout. The filter support was set to \todo{$K = 25$}.
 
-\begin{table*}[h!]
- \centering
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
- Architecture & $G_1=$ 2D Euclidean grid  & $G_2=$ Random graph \\
-\hline
-graph CNNs: CN32-softmax & 97.40 & 96.88\\
-graph CNNs: CN32-P4-CN64-P4-FC512-softmax & 98.65 & 95.39 \\
-\hline
-\end{tabular}
-}
-\caption{Influence of the graph quality over CNNs on graphs.} 
-\label{tab3}
+\subsection{Definition of spectral filters}
+
+To compare our spectral filters with those proposed in \cite{\bruna}, we measure
+the performance of the graph CNN model for each filter definition on the MNIST
+and 20NEWS datasets. \tabref{filters} shows that our parametrization outperforms
+\cite{\bruna} as well as non-parametric filters, which are not localized and
+require $\bO(n)$ parameters to learn, in all cases.
+
+\begin{table*}[h!] \centering
+\begin{tabular}{llccc} \toprule
+& & \multicolumn{3}{c}{Accuracy} \\
+\cmidrule{3-5}
+Dataset & Architecture & Non-Param \eqnref{filt_non-param} &
+Spline \eqnref{filt_spline} \cite{\bruna} &
+Chebyshev \eqnref{filt_cheby} \\
+\midrule
+MNIST & GC10 & 95.75 & 97.26 & \textbf{97.48} \\
+MNIST & GC32-P4-GC64-P4-FC512 & 96.28 & 97.15 & \textbf{98.65} \\
+20NEWS & GC32 & \todo{?} & \todo{?} & \todo{?} \\
+\bottomrule \end{tabular}
+\caption{Comparison of accuracy results for different types of spectral filters.
+Spline and Chebyshev filters use \todo{$K=20$}, \todo{$K=5$ for 20NEWS}.} 
+\label{tab:filters}
 \end{table*}
 
-%\newpage
+\subsection{Computational efficiency}
 
-\subsection{Text Categorization with the 20NEWS Dataset}
-\vspace{-0.2cm}
-For this section, we consider the text categorization task with the 20NEWS dataset which consists of 18,846 text documents associated with 20 classes \cite{art:Joachims9620NEWS}.
-The number of unique words in this corpus is 93,953, so the dimensionality of these data points is around 94k.
-We extract 10,000 words that most occur in the corpus.\\
-The original full dataset is split into a training set of 11,314 documents and a test set of 7,532 documents.
-Our goal is to classify the test set.
-We remind that standard CNNs cannot be directly applied to this problem.
-A natural alternative would be to use a deep fully connected networks, but they are computationally expensive.
-To decrease the computational learning complexity of fully connected networks, the authors of \cite{art:HenaffBrunaLeCun15DLgraphs} suggested to carry out dropout regularization \cite{art:SrivastavaHintonKrizhevskySutskeverSalakhutdinov14dropout} on these networks.
-Table \ref{tab4a} reports the classification performances of a set of algorithms on 20NEWS, and the proposed graph CNNs with one convolutional layer of 32 filters and a softmax classifier.
+\figref{runtime} indeed demonstrates that the training time for our model scales
+as $\bO(n)$ while \cite{\bruna} scales as $\bO(n^2)$. Moreover,
+\figref{convergence} gives a sense of how the validation accuracy as well as the
+loss energy converges w.r.t. the three filter definitions. Finally,
+\tabref{speedup} compares training time on CPU and GPU. The fact that we observe
+the same order of speedup as classical CNNs exemplifies the natural
+parallelization opportunity offered by our model. That is possible because our
+method relies solely on matrix multiplications which are efficiently implemented
+by cuBLAS, the linear algebra routines provided by NVIDIA.
 
-\begin{table*}[h!]
- \centering
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
- Algorithms & Word2vec features  \\
-\hline
-Linear SVM & 65.90  \\
-Multinomial NB & 68.51 \\
-Softmax & 66.28 \\
-FC + softmax + dropout & 64.64 \\
-FC + FC + softmax + dropout & 65.76 \\
-graph CNNs: CN32-softmax & 68.26 \\
-\hline
-\end{tabular}
-}
-\caption{Classification accuracies on 20NEWS with features learned by word2vec \cite{pro:MikolovChenCorradoDean13word2vec}.} 
-\label{tab4a}
+\begin{table*}[h!] \centering
+\begin{tabular}{llccc} \toprule
+& & \multicolumn{2}{c}{Time (ms)} & \\
+Model & Architecture & CPU & GPU & Speedup \\
+\midrule
+Classical CNN & C32-P4-C64-P4-FC512 & 210 & 31 & 6.77x \\
+Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 200 & 25 & 8.00x \\
+\bottomrule \end{tabular}
+\caption{Time to process a mini-batch of 100 MNIST images. That is the total
+training time divided by the number of gradient steps.} 
+\label{tab:speedup}
 \end{table*}
-
-\noindent
-As text documents do not naturally lie on a graph, it is here necessary to construct a graph of words' embedding.
-This is a very essential step as the graph quality will have a strong impact on the ability of convolutional layers to identify local stationary features present in the data.
-We also observe that local stationarity is an assumption that is made about the data, and the CNNs are fundamentally designed on this data hypothesis.
-However, we believe this stationarity assumption to be present in many real-world data, as structured data are usually composed of repetitive small patterns at different scales of observation.
-We investigate here 5 graphs of words: a graph of learned embedding of the corpus using word2vec \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of normalized word counts across each document,
-a graph of pre-trained embedding of words \cite{pro:MikolovChenCorradoDean13word2vec}, a graph of approximate nearest neighbors (ANN) using LSHForest on the learned embedding \cite{pro:MikolovChenCorradoDean13word2vec}, and a random graph.
-The classification results are reported in Table \ref{tab4b}.
-It shows that the graph $G_1$ learned from the training set is the best at capturing the local stationarity property of text documents, improving the fully connected networks performance from 65.76\%, reported Table \ref{tab4a}, to 68.26\%.
-It is worth noticing that the approximate $k$-NN graph $G_4$ constructed from the best learned features is almost as worse as a random graph, meaning that it is a better strategy to learn a good low-dimensional words' embedding first and then construct an exact $k$-NN graph from this embedding.
-As the graph quality has an impact on the performance of the classification system, a next step would be to learn a better graph as already pointed in \cite{art:HenaffBrunaLeCun15DLgraphs}.
-A natural and future approach would be to alternate learning of the NN weights and learning of the graph of data, as a virtuous circle.
-
-\begin{table*}[h!]
- \centering
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
- Architecture & $G_1$= learned  & $G_2$= normalized  & $G_3$= pre-trained & $G_4$= ANN  &  $G_5$= random   \\
-  & graph \cite{pro:MikolovChenCorradoDean13word2vec} & word count graph & graph  &  graph  & graph \\
-\hline
-CN32-softmax & 68.26 & 67.50 & 66.98 & 67.86 & 67.75  \\
-\hline
-\end{tabular}
-}
-\caption{Classification accuracies of graph CNNs with different graph constructions on 20NEWS.} 
-\label{tab4b}
-\end{table*}
-
-\subsection{Comparison with Non-Parametric Filters and Henaff-Bruna-LeCun \cite{art:HenaffBrunaLeCun15DLgraphs}}
-\vspace{-0.2cm}
-The advantages of our graph CNNs over \cite{art:HenaffBrunaLeCun15DLgraphs} are listed below:\\
-$\bullet$ Improved learning complexity from quadratic to linear: In \cite{art:HenaffBrunaLeCun15DLgraphs}, graph convolutional operations need $O(n^2)$ operations, while our technique requires $O(E.K)$ operations, where $E$ is the number of edges and $K$ is the polynomial order of the spectral filters.
-As we consider $k$-NN graphs, $E=n.k$ and the complexity becomes linear $O(n)$ w.r.t.
-the input size $n$, while  \cite{art:HenaffBrunaLeCun15DLgraphs} stays quadratic.
-To give some figures, in our experiments, we have for the first layer, $n=10^4, k=8, K=32$, so the total number of operations is $E.K=10^4.8.32 \simeq 3.10^5$ and $n^2=10^8$.
-Figure \ref{fig2} compares the computational times.\\
-$\bullet$  No eigenvalue decomposition required: The spectral graph CNNs \cite{art:HenaffBrunaLeCun15DLgraphs} require to compute a full eigenvalue decomposition (EVD) to compute the Fourier basis.
-This has a complexity of $0(n^3)$ with $n=10^{4}$, which is memory expensive.
-Our technique does not require an EVD, it only requires multiplications between the graph Laplacian sparse matrix and vectors.
-\\
-$\bullet$ Improved memory efficiency: \cite{art:HenaffBrunaLeCun15DLgraphs} needs to store in memory the Fourier basis of size $O(n^2)$, this can greatly limit the performance of GPU computing.
-Our technique requires to store the sparse Laplacian of size $O(E)=O(n)$ in the case of k-NN graphs.
-
-%\begin{table*}[h!]
- %\centering
-%{\small
-%\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-%\hline
-% Nb of words & Non-Param filters  & Splines filters \cite{art:HenaffBrunaLeCun15DLgraphs} & Chebyshev filters [Our model] \\
-%\hline
-%1k words & 25ms & 25ms & 18.75ms \\
-%2k words & 57.5ms & 58.75ms & 28.75ms \\
-%4k words & 170ms & 170ms & 50ms \\
-%6k words & 348.75ms & 348.75ms & 73.75ms \\
-%8k words & 578.75ms & 580.0ms & 95.0ms \\
-%\hline
-%\end{tabular}
-%}
-%\caption{Comparison of computational times between non-parametric filters (CN32+softmax), spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} (CN32+softmax and $K=5$) and Chebyshev-based filters (CN32+softmax and $K=5$) on 20NEWS.
-We consider a mini-batch size of 100 data, and computations are carried out on an NVIDIA GPU K40c.}
-%\label{tab5}
-%\end{table*}
 
@@
 \includegraphics[width=5cm]{images/speed.png}
-\caption{Plots of computational times w.r.t.
-the number of selected words in 20NEWS for non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters.
-The architecture is CN32+softmax with $K=5$ for splines and Chebyshev.
-We consider a mini-batch size of 100 data, and computations are carried out on an NVIDIA GPU K40c.}
-\label{fig2}
+\caption{Training time over a mini-batch of 100 20NEWS documents w.r.t. $n$, the
+number of selected words, for the three considered filters. The architecture is
+GC32 with $K=5$ for Spline and Chebyshev.}
+\label{fig:runtime}
 \end{figure}
 
-\noindent
-Observe that non-parametric filters are most costly as they require $O(n)$ parameters to learn, and they are unable to learn localized filters, unlike \cite{art:HenaffBrunaLeCun15DLgraphs} and our model.
-This is particularly illustrated on Table \ref{tab6}, where non-parametric filters provide less accurate results.
-Besides, our model outperforms the classification accuracy of \cite{art:HenaffBrunaLeCun15DLgraphs} on MNIST.
+\begin{figure}[h!]
+\centering
+\subfigure[]{\includegraphics[width=0.49\textwidth]{images/acc.png}}
+\hfill
+\subfigure[]{\includegraphics[width=0.49\textwidth]{images/loss.png}}
+\caption{Plots of accuracy (a) and energy loss (b) for the first 2000
+iterations.}
+\label{fig:convergence}
+\end{figure}
+
+\subsection{Influence of graph quality} \label{sec:graph_quality}
+
+For our method to be successful, his statistical assumptions shall be fulfilled
+on the graph where the data resides. The performance, and the quality of the
+learned filters, thus critically depends on the quality of the graph. By the
+success of classical CNNs, we know these assumptions to be true for images in
+Euclidean domains. By our experiments, we know also that a $k$-NN graph of the
+grid is good enough (we did not loose much compared to classical CNNs), and the
+value of $k$ does not have a strong influence. The extent to which a particular
+graph satisfies the assumptions is to compare its performance with a random
+graph, which for sure does not. \tabref{mnist_quality} shows that when using a
+random graph, the accuracy drops (\todo{to the order of an architecture without
+convolution}), as the convolutional layers are not useful anymore to extract
+meaningful features. The structure has been lost.
 
@@
 \centering
-{\small
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
- Architecture & Accuracy  \\
-\hline
-CN10 (Non-Param filters) + softmax & 95.75 \\
-CN10 (Spline filters)  + softmax & 97.26 \\
-CN10 (Chebyshev filters)  + softmax & 97.48 \\
-CN32-P4-CN64-P4-FC512-softmax  (Non-Param filters) & 96.28 \\
-CN32-P4-CN64-P4-FC512-softmax  (Spline filters) & 97.15 \\
-CN32-P4-CN64-P4-FC512-softmax  (Chebyshev filters) & 98.65 \\
-\hline
-\end{tabular}
-}
-\caption{Comparison of precisions between non-parametric filters, spline-based filters \cite{art:HenaffBrunaLeCun15DLgraphs} and Chebyshev-based filters [our model] on MNIST.
-For spline and Chebyshev filters, we picked $K=20$.}
-\label{tab6}
+\begin{tabular}{lcc} \toprule
+Architecture & 8-NN on 2D Euclidean grid & random \\
+\midrule
+GC32 & 97.40 & 96.88 \\
+GC32-P4-GC64-P4-FC512 & 98.65 & 95.39 \\
+\bottomrule \end{tabular}
+\caption{Classification accuracies with different graph constructions on MNIST.} 
+\label{tab:mnist_quality}
 \end{table*}
 
-\noindent
-Eventually, we plot in Figure  \ref{fig1} the curves of accuracy and energy loss w.r.t. the number of iterations for the three types of filters.
+While images naturally lie on a graph, the 2D grid, that is not the case for
+text documents. For our method to be applied, a feature graph does have to be
+built.
+We investigate here 5 $16$-NN similarity graphs of words. The simplest option is
+to represent each word as its corresponding column in the bag-of-words matrix.
+Another approach is to learn embeddings on the corpus with word2vec
+\cite{pro:MikolovChenCorradoDean13word2vec} or to use pre-trained embeddings on
+Google News (given by the authors). As the dataset gets larger (in number of
+samples and dimensionality), it is often not an option to compute the distance
+between all features, such that an approximate nearest neighbors algorithm shall
+be used. We used the LSHForest \cite{pro:BawaCondieGanesan05LSHForest} on the
+learned word2vec embedding. \tabref{20news_quality} reports the classification
+results. It shows that the graph built on the learned word2vec embedding is the
+best at capturing the local and stationarity properties of text documents. It is
+worth noticing that the approximate $k$-NN graph constructed from it is almost
+as bad as the random graph, meaning that it may be a better strategy to learn a
+good low-dimensional embedding first and then construct an exact $k$-NN graph
+from this embedding.
 
-\begin{figure}[h!]
-\centering
-\subfigure[]{\includegraphics[width=6cm]{images/acc.png}}
-\hspace{0.5cm}
-\subfigure[]{\includegraphics[width=6cm]{images/loss.png}}
-\caption{Plots of accuracy (a) and energy loss (b) w.r.t. the number of iterations.}
-\label{fig1}
-\end{figure}
+\begin{table*}[h!] \centering
+\begin{tabular}{ccccc} \toprule
+& \multicolumn{2}{c}{word2vec} & & \\
+bag-of-words & pre-trained & learned & approximate & random \\
+\midrule
+67.50 & \todo{66.98} & 68.26 & 67.86 & 67.75 \\
+\bottomrule \end{tabular}
+\caption{Classification accuracies of GC32 with different graph constructions on 20NEWS.} 
+\label{tab4b}
+\label{tab:20news_quality}
+\end{table*}
 
 \section{Conclusion and Future Work}
-\vspace{-0.2cm}
-We have introduced an efficient implementation of CNNs on non-Euclidean manifolds represented by graphs using tools from SPG.
-Experiments have shown the ability of the model to extract the local stationary features through the graph convolutional layers.
-Future works will investigate two directions.
-First, we will explore applications of this generic model to important fields where the data naturally lie on non-artificially constructed graphs such as social networks, biological networks, or telecommunication networks.
-It makes actually more sense to apply this generalized CNN framework to natural network-based data, rather than artificially created networks which quality may vary as seen in the experiments.
-Second, we will also improve this framework with tools developed in SGP, including graph wavelet operators \cite{art:HammondVandergheynstGribonval11GraphWav,art:CoifmanLafon06DifMap,pro:GavishNadlerCoifman10GraphHaar,pro:ChenChengMallat14deepHaar,pro:RustamovGuibas14deepHaar}, spectral graph coarsening techniques \cite{art:ShumanFarajiVandergheynst16PyramTrans}, and improved localization properties \cite{pro:TsitsveroBarbarossa15Uncert,pro:PasdeloupAlamiGriponRabbat15Uncert,art:PerraudinRicaudShumanVandergheynst16Uncert}.
+
+We have introduced an efficient implementation of CNNs on non-Euclidean
+manifolds represented by graphs using tools from GSP. Experiments have shown the
+ability of the model to extract local and stationary features through the graph
+convolutional layers. Compared with the first generation of spectral graph CNNs
+introduced in \cite{\bruna}, our model provides a strict control over the local
+support of filters, is computationally more efficient (\figref{runtime}) by
+avoiding an explicit use of the Graph Fourier basis and experimentally shows a
+better test accuracy (\tabref{filters}).
+
+In the end we addressed the three concerns raised by
+\cite{art:HenaffBrunaLeCun15DLgraphs}:
+\begin{itemize}
+\item We introduced a model whose computational complexity is linear with the
+	dimensionality of the data (\figref{runtime}).
+\item As they showed already, we confirm that the quality of the input graph is
+	of paramount importance (\tabref{20news_quality}).  Their idea to learn it
+	from data makes a lot of sense. A natural and future approach would be to
+	alternate the learning of the NN parameters and the graph, as a virtuous
+	circle.
+\item We showed that the statistical assumptions of stationarity and
+	compositionality made by the model are verified for text documents
+	(\tabref{20news}), as long as the graph is well constructed.
+\end{itemize}
+
+Future works will investigate two directions. First, we will explore
+applications of this generic model to important fields where the data naturally
+lie on non-artificially constructed graphs such as social networks, biological
+networks, or telecommunication networks. It makes actually more sense to apply
+this generalized CNN framework to natural network-based data, rather than
+artificially created networks which quality may vary as seen in the experiments.
+Second, we will also improve this framework with tools developed in GSP,
+including graph wavelet operators
+\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
+	pro:GavishNadlerCoifman10GraphHaar, pro:ChenChengMallat14deepHaar,
+	pro:RustamovGuibas14deepHaar}, spectral graph coarsening techniques
+	\cite{art:ShumanFarajiVandergheynst16PyramTrans}, and improved localization
+	properties \cite{pro:TsitsveroBarbarossa15Uncert,
+	pro:PasdeloupAlamiGriponRabbat15Uncert,
+art:PerraudinRicaudShumanVandergheynst16Uncert}.
+
+% more data, natural graphs (no need to worry about construction, external info)
+% data graph vs feature graph
 
 \newpage
-
-%\vspace{-0.15cm} in .bbl
 \bibliographystyle{plain}
-{\setstretch{0}
-{\small
-%\bibliography{bib_nips16}
-\bibliography{bib_nips16_short}
-}
+{\setstretch{0} \small
+\bibliography{refs}
+%\bibliography{bib_nips16_short}
 }

### 2016-10-29 nips2016/nips2016.tex
nips v6: edition by Xavier (2016-06-29)

@@
 %\usepackage{nips_2016}
-\usepackage[final]{nips2016}
+%\usepackage[final]{nips_2016}
+\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry} % arxiv
 
@@
 %\usepackage{parskip} \setlength{\parskip}{5pt}
-\setlength{\bibsep}{0.2pt}
+%\setlength{\bibsep}{0.2pt} %comment for arxiv
 
 \title{Convolutional Neural Networks on Graphs\\
-with Fast Localized Spectral Filtering}
+with Fast Localized Spectral Filtering\\}
 
+%\author{
+%  Michaël Defferrard \\ %\thanks{further info} \\
+%  %Michael Defferrard \\ %\thanks{further info} \\
+%  %Department of Electrical Engineering \\
+%  %École Polytechnique Fédérale de Lausanne (EPFL) \\
+%  EPFL, Lausanne, Switzerland \\
+%  \texttt{michael.defferrard@epfl.ch} \\
+%  \And % \AND
+%  Xavier Bresson \\
+%  EPFL, Lausanne, Switzerland \\
+%  \texttt{xavier.bresson@epfl.ch} \\
+%  \And % \AND
+%  Pierre Vandergheynst \\
+%  EPFL, Lausanne, Switzerland \\
+%  \texttt{pierre.vandergheynst@epfl.ch} \\
+%}
+
+%%Arxiv
 \author{
-  Michaël Defferrard \\ %\thanks{further info} \\
-  %Department of Electrical Engineering \\
-  %École Polytechnique Fédérale de Lausanne (EPFL) \\
+  Micha\"{e}l Defferrard\hspace{1cm}Xavier Bresson\hspace{1cm} Pierre Vandergheynst\\ \\
+  Department of Electrical Engineering \\
   EPFL, Lausanne, Switzerland \\
-  \texttt{michael.defferrard@epfl.ch} \\
-  \And % \AND
-  Xavier Bresson \\
-  EPFL, Lausanne, Switzerland \\
-  \texttt{xavier.bresson@epfl.ch} \\
-  \And % \AND
-  Pierre Vandergheynst \\
-  EPFL, Lausanne, Switzerland \\
-  \texttt{pierre.vandergheynst@epfl.ch} \\
+  {\small \texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} } \\\\
 }
@@
 	formulation of CNNs from low-dimensional regular Euclidean domains, where
-	image (2D), video (3D) and audio (1D) are represented, to high-dimensional
-	irregular domains represented by graphs. This paper introduces a formulation
+	images (2D), videos (3D) and audios (1D) are represented, to high-dimensional
+	irregular domains such as social networks or biological networks represented by graphs. This paper introduces a formulation
 	of CNNs on graphs in the context of spectral graph theory. We borrow the
@@
 	which provides the necessary mathematical background and efficient numerical
-	schemes to design localized graph filters which are fast to evaluate.
-	Similarly to classical CNNs, the computational complexity as well as the
-	number of parameters to learn are independent of the data dimensionality.
-	Numerical experiments on MNIST and 20NEWS demonstrate the ability of the
-	model to learn local and stationary features on graphs, as long as the graph
-	is well-constructed. This work can be thought as the foundation of a second
-	generation of CNNs on graphs built on spectral graph theory, being
-	mathematically well defined, bringing better test performance and being
-	faster to evaluate and train than the first generation introduced by Bruna,
-	LeCun et al.
+	schemes to design localized graph filters efficient to learn and evaluate. 
+	As a matter of fact, we introduce the first technique that offers the same computational complexity than standard CNNs, while being universal to any graph structure. 
+	%Similarly to classical CNNs, the computational complexity as well as the
+	%number of parameters to learn are independent of the data dimensionality.
+	Numerical experiments on MNIST and 20NEWS demonstrate the ability of this novel deep learning system to learn local, stationary, and compositional features on graphs, as long as the graph
+	is well-constructed. %This work can be thought as the foundation of a second
+	%generation of CNNs on graphs built on spectral graph theory, being
+	%mathematically well defined, bringing better test performance and being
+	%faster to evaluate and train than the first generation introduced by Bruna,
+	%LeCun et al.
 
@@
 coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
-\label{fig:}
+\label{fig_illustration}
 \end{figure}
@@
 
-Convolutional neural networks (CNNs) \cite{pro:LeCunBottouBengioHaffner98MNIST}
+Convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST}
 offer an efficient architecture to extract highly meaningful statistical
@@
 
-A generalization to irregular grids is not direct as the building blocks of
-CNNs, the convolution and downsampling operators, are not mathematically clear
-on these domains. This makes this extension challenging, both theoretically and
+A generalization to irregular grids is not straightforward as the building blocks of
+CNNs, the convolution, the downsampling, and the pooling operators, cannot be directly applied to these domains. This makes this extension challenging, both theoretically and
 implementation-wise. This work leverages the recent works of
@@
 geometric structures, and can be studied with strong mathematical tools such as
-spectral graph theory, a blend between graph theory and harmonic analysis.
-
-The major bottleneck of generalizing CNNs to graphs, and the primary goal of
-this work, is the definition of localized filters which are efficient to
-evaluate and learn.
-%that are the core elements to extract local stationarity structures.
-We will make use of recent tools developed in the context of graph signal
-processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
-to precisely control the spatial support of graph filters as well as to allow
-fast evaluations. Eventually, a generic and efficient implementation of CNNs for
-graph signals may pave the way to the development of a potentially large class
-of deep learning techniques for graph-based data.
+spectral graph theory \cite{book:Chung97Spectral}, a blend between graph theory and harmonic analysis.
 
@@
 
-\todo{spatial approaches provide localization by definition}
-\todo{why spectral approach?}
-While a spatial approach is conceivable, it faces the challenge of matching
-local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.
+The major bottleneck of generalizing CNNs to graphs, and one of the primary goals of
+this work, is the definition of graph localized filters which are efficient to
+evaluate and learn.
+%that are the core elements to extract local stationarity structures.
+We will make use of recent tools developed in the context of graph signal
+processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to achieve such goals. Precisely, the main contributions of this work are summarized below:
+\begin{enumerate}
+%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
+\item \textbf{Spectral formulation.} A graph spectral theoretical formulation of CNNs on graphs built on
+	established tools in GSP.
+\item \textbf{Strictly localized filters.} Enhancing \cite{\bruna}, the proposed spectral
+	filters are provable to be strictly localized in a ball of radius $K$, i.e. $K$ hops from
+	the central vertex.  
+\item \textbf{Low computational complexity.} The evaluation complexity of our filters is
+	linear w.r.t.
+the filters support's size $K$ and the number of edges $|\E|$.
+Importantly, as most real-world graphs are highly sparse, we have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor (NN) graphs, leading to a linear complexity w.r.t the input data size.
+Moreover,
+	this method avoids the Fourier basis altogether, thus the expensive
+	eigenvalue decomposition (EVD) necessary to compute it as well as the need
+	to store the basis, a matrix of size $n^2$. That is especially relevant when
+	working with limited GPU memory. Besides the data, our method only requires
+	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
+\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
+	graphs which, after a rearrangement of the vertices as a binary tree structure, is analog to pooling of
+	1D signals.
+\item \textbf{Experimental results.} We present many experiments that ultimately
+	show that our formulation is (1) a useful model, (2) computationally
+	efficient and (3) superior both in accuracy and complexity to the pioneer spectral graph CNNs introduced in \cite{\bruna}. We also show that
+	our graph formulation performs similarly as a classical CNNs on MNIST and study the
+	impact of various graph constructions on classification performance.
+\end{enumerate}
 
-\todo{Another challenge is the definition of a pooling layer.}
-Local stationarity property of data is extracted via localized convolutional
-kernels. In this section, we are interested to extract the multi-scale
-hierarchical composition property of data. This is efficiently achieved via grid
-subsampling or downsampling and pooling, which trades spatial resolution with
-feature resolution reducing the learning complexity without compromising the
-system performances. In the case of standard CNNs for data lying on regular
-grids, the subsampling operation is mathematically sound. The translation
-operator commutes with the downsampling operator, hence not affecting the
-translation invariance property of the system at different scales of grid
-resolutions.
+\noindent
+Eventually, generic and efficient implementations of CNNs on
+graphs, as proposed here and related works \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
+art:MasciBoscainiBronsteinVandergheynst15ShapeNet,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, is essential to pave the way to the development of a new class of deep learning techniques for graph-based data.
 
-\paragraph{Graph Signal Processing.} GSP is an emerging field that aims at
-bridging the gap between signal processing techniques like wavelet analysis and
+\section{Related Works}
+\subsection{Graph Signal Processing} GSP is an emerging field that aims at
+bridging the gap between signal processing techniques like wavelet analysis \cite{book:Mallat99wavelets} and
 graph theory such as spectral graph analysis
@@
 \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
-\todo{others for politics} for an
+ for an
 introduction of the field. Standard operations on grids such as convolution,
@@
 
-\todo{Model statistical assumptions: locality, stationarity, compositionality.}
-These assumptions can be experimentally verified for any dataset. While we know
-for them to be true for audio, images and videos, we show experimentally that
-they might be satisfied for text documents as long as the graph is properly
-constructed.
-We also observe that
-local stationarity is an assumption that is made about the data, and the CNNs
-are fundamentally designed on this data hypothesis. However, we believe this
-stationarity assumption to be present in many real-world data, as structured
-data are usually composed of repetitive small patterns at different scales of
-observation.
+In the future, we expect to
+further benefit from newly developed tools in GSP to enhance our mathematical formulation of
+CNNs on graphs. This paper introduces the mathematical and computational foundations of
+such models, while future developments may consider to enhance the building
+blocks, namely graph filtering and coarsening.
 
-\section{Related Work}
+\subsection{CNNs on Non-Euclidean Domains}
 
@@
 
-Extending CNNs to non-Euclidean domains is a recent line of work. A first
-attempt is the so-called local reception fields \cite{pro:GregorLeCun10LRF,
+Extending CNNs to non-Euclidean domains has been a recent line of work. A preliminary work was proposed with 
+ the so-called local reception fields \cite{pro:GregorLeCun10LRF,
 pro:CoatesNg11LRF}, successfully applied to image recognition. The main idea is
@@
 a limited number of connections between two successive layers. While this model
-reduces the number of parameters by exploiting the locality assumption, it does
-not attempt to exploit any weight-sharing strategy.
-%stationarity This approach is however not able to extract similar groups over
+reduced the number of parameters by exploiting the locality assumption, it did
+not attempt to exploit any stationarity property of data, i.e. no weight-sharing strategy. %stationarity This approach is however not able to extract similar groups over
 %the data domain, as it is not clear how to define convolutional filters on
 %these groups.
-\todo{read papers}
+%\todo{read papers}
 
-ShapeNet \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
+A first attempt to design CNNs on graphs was introduced in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} by adopting a spatial approach.
+The idea was to construct a multiresolution graph by exploiting graph neighbourhood structures, and learned a deep neural network.
+However, such approach did not consider convolution operations, and thus no stationarity property.
+%By construction, spatial approaches provide filter localization.
+However, although graph convolution directly in the spatial domain is conceivable, it also faces the challenge of matching local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.
+
+ShapeNet in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
 art:MasciBoscainiBronsteinVandergheynst15ShapeNet} is a generalization of CNNs
 to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. The
-authors used geodesic polar coordinates to define mesh patches and formulated a
+authors used geodesic polar coordinates to define convolution operations on mesh patches, and formulated a
 deep learning architecture which allows comparison across different manifolds.
-They obtained state-of-the-art results for shape recognition tasks.
+They obtained state-of-the-art results for 3D shape recognition tasks.
 
@@
 recognition on sphere, and to sparse reconstruction of faces.
-\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
-they impose a shape to the filters.}
-\todo{Unknown graph geometry?}
+%\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
+%they impose a shape to the filters.}
+%\todo{Unknown graph geometry?}
 
@@
 the graph structure from the data, and applied their model on image recognition,
-text categorization and bioinformatics. This approach does however not scale due
+text categorization and bioinformatics. This approach does however not scale up due
 to the necessary multiplications by the graph Fourier basis, a matrix of $n^2$
@@
 provide a precise control over the local support of their kernels, which is
-essential to learn localized filters. This work directly builds on theirs, and
-we will show how we overcame these limitations and what we further gained.
+essential to learn localized filters. Our technique leverages on this work, and
+we will show how to overcome these limitations and beyond.
 
-\subsection{Contributions}
+\section{Proposed Technique}
+Generalizing CNNs to graphs requires three fundamental steps.
+First step is the design of localized convolutional filters on graphs.
+Second step is a graph coarsening procedure that groups together similar features on graphs.
+And the last step is the graph pooling operation that trades spatial resolution for higher filter resolution.
 
-Our contributions can be summed up as follows:
-\begin{enumerate}
-%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
-\item \textbf{Spectral formulation.} A clean spectral formulation built on
-	established tools in GSP.
-\item \textbf{Strictly localized filters.} Unlike \cite{\bruna}, the proposed spectral
-	filters are strictly localized in a ball of radius $K$, i.e. $K$ hops from
-	the central vertex.
-\item \textbf{Low computational complexity.} The evaluation of our filters is
-	linear w.r.t. the filters support's size $K$ and the number of edges $n \leq
-	|\E| \leq n^2$. As most real-world graphs are highly sparse, we have $|\E|
-	\ll n^2$ and $|\E| = kn$ for the widespread $k$-neighbors graphs. Moreover,
-	this method avoids the Fourier basis altogether, thus the expensive
-	eigenvalue decomposition (EVD) necessary to compute it as well as the need
-	to store the basis, a matrix of size $n^2$. That is especially relevant when
-	working with limited GPU memory. Besides the data, our method only requires
-	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
-\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
-	graphs which, after a rearrangement of the vertices, is analog to pooling of
-	1D signals.
-\item \textbf{Experimental results.} We present many experiments that ultimately
-	show that our formulation is (1) a useful model, (2) computationally
-	efficient and (3) superior both in accuracy and complexity to the previous
-	generation of spectral graph CNNs introduced in \cite{\bruna}. We also show that
-	our formulation performs similarly as a classical CNN on MNIST and study the
-	impact of various graph constructions on performance.
-\end{enumerate}
+\subsection{Learning Fast Localized Spectral Filters} \label{sec:filters}
+The goal of this section is to define fast convolutional localized filters on graphs.
+There are two strategies to define convolutional localized filters; either from a spatial approach or from a spectral approach.
+By construction, spatial approaches provide filter localization via the finite size of the kernel.
+However, although graph convolution directly in the spatial domain is conceivable, it also faces the challenge of matching local neighbourhoods, as pointed out in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.
+Consequently, there is no unique mathematical definition of spatial translations on graphs from a spatial perspective.
+On the other side, a spectral approach provides a well-defined translation operator on graphs via convolutions with a Kronecker delta function implemented in the spectral domain
+\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
+However, a filter defined in the spectral domain is not naturally localized and translations are costly with $\bO(n^2)$ operations due to multiplications with the graph Fourier basis.
+Both limitations can be overcome with a special choice of filter parametrization, as explained below.
 
-Beyond this work, our long-term goal is to make use of GSP tools as much as
-possible to formulate mathematically well-designed techniques for CNNs on
-graphs. This paper introduces the mathematical and computational foundations of
-such models, while future developments may consider to enhance the building
-blocks, namely graph filtering and coarsening. In the future, we expect to
-further benefit from newly developed tools in GSP to enhance our formulation of
-CNNs on graphs.
-
-\section{Learning Fast Localized Spectral Filters} \label{sec:filters}
-
-As there is no notion of spatial translation on graphs, convolutions are
-implemented as multiplications in the spectral domain
-\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. While a spatial
-convolution on an Euclidean domain is localized by definition (via the finite
-size of the kernel) and has a linear computational complexity $\bO(n)$, a filter
-defined in the spectral domain is not localized by default and costs $\bO(n^2)$
-operations because of the necessary multiplication by the graph Fourier basis.
-Note that while classical CNNs can be speed-up by leveraging the Fourier basis
-(instead of computing the convolutions in the spatial domain)
-\cite{mathieu_fast_2013}, it is only advantageous because of the $\bO(n \log(n))$
-complexity of the FFT, which has no equivalent on graphs (yet). A good choice of
-filter parametrization however allows us to prove that when translated (via a
-convolution with a delta) to a given center vertex, the filter is strictly
-localized in a ball of radius $K$ around that center vertex. Moreover, a
-recursive formulation allows the filtering operation to scale linearly with $n$.
+%Note that while classical CNNs can be speed-up by leveraging the Fourier basis
+%(instead of computing the convolutions in the spatial domain)
+%\cite{mathieu_fast_2013}, it is only advantageous because of the $\bO(n \log(n))$
+%complexity of the FFT, which has no equivalent on graphs (yet).
 
@@
 \paragraph{Polynomial parametrization for localized filters.} There are however
-two problems with non-parametric filters: (1) they are not localized and (2)
-they are of learning complexity $\bO(n)$, the dimensionality of the data. We are
-going to solve those two problems together by reducing this complexity to
-$\bO(K)$, the support's size of the filter, exactly as it is for standard CNNs.
-While \cite{\bruna} solves it by parameterizing $g_\theta(\Lambda)$ as a cubic
-B-spline defined by $K$ control points, let us parametrize it as a polynomial
-function
+two limitations with non-parametric filters: (1) they are not localized in space and (2)
+their learning complexity is in $\bO(n)$, the dimensionality of the data. These issues can be overcome 
+with the use of Laplacian-based polynomial spectral filters: 
 \begin{equation} \label{eq:filt_poly}
@@
 (L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
-delta $\delta_i \in \R^n$. By \cite[Lemma
+delta function $\delta_i \in \R^n$. By \cite[Lemma
 5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
-$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the distance
-between two vertices is the minimum number of edges in any path connecting them.
+$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum number of edges connecting two vertices on the graph.
 Consequently, spectral filters represented by $K^\text{th}$-order polynomials of
-the Laplacian are exactly $K$-localized.
+the Laplacian are exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the support size of the filter, and thus the same complexity as in standard CNNs.
 
@@
 learn localized filters with $K$ parameters, the cost to filter a signal $x$ as
-$y = Ug_\theta(\Lambda)U^Tx$ is still of $\bO(n^2)$ operations because of the
-multiplications with the Fourier basis $U$. A way out is to parametrize
+$y = Ug_\theta(\Lambda)U^Tx$ is still high with $\bO(n^2)$ operations because of the
+multiplications with the Fourier basis $U$. A solution of this problem is to parametrize
 $g_\theta(L)$ as a polynomial function that can be computed recursively from
@@
 \times n}$ is the Chebyshev polynomial of order $k$ evaluated at the scaled
-Laplacian $\tilde{L} = 2 L / \lambda_{max} - I_n$. Note that the spectrum of the
-normalized Laplacian is bounded by $2$ \cite{book:Chung97Spectral}, such that
-the scaling can simply be $\tilde{L} = L - I_n$. Denoting $\bar{x}_k =
+Laplacian $\tilde{L} = 2 L / \lambda_{max} - I_n$. %Note that the spectrum of the
+%normalized Laplacian is bounded by $2$ \cite{book:Chung97Spectral}, such that
+%the scaling can simply be $\tilde{L} = L - I_n$. 
+Denoting $\bar{x}_k =
 T_k(\tilde{L})x \in \R^n$, we can use the recurrence relation to compute
@@
 \end{equation}
-then costs $\bO(K|\E|))$ operations.
+then costs $\bO(K|\E|)$ operations.
 
@@
 training multiple convolutional layers with the backpropagation algorithm, one
-needs the two gradients
+needs the two gradients:
 \begin{align}
@@
 F_{out} S)$ operations. These can be efficiently computed on parallel
-architectures by leveraging tensor operations. As an optimization,
-$[\bar{x}_{s,i,0}, \ldots, \bar{x}_{s,i,K-1}]$ can be computed once.
+architectures by leveraging tensor operations. Eventually, $[\bar{x}_{s,i,0}, \ldots, \bar{x}_{s,i,K-1}]$ only need to be computed once.
 
-\section{Efficient Pooling on Graphs} \label{sec:pooling}
-
-% Why are we introducing this new pooling definition?
-
-\todo{Why? Bruna uses an ad-hoc technique.}
-
-\paragraph{Graph Coarsening.} In contrast with regular domains, to reduce the
-sampling of a graph signal, we need to construct meaningful neighborhoods on
-graphs where similar vertices may be clustered together. Doing this for multiple
+\subsection{Graph Coarsening} \label{sec:coarsening}
+Local stationarity property of data is extracted via localized convolutional kernels.
+We are now interested to extract the multi-scale hierarchical composition property of data.
+In standard CNNs, this is efficiently achieved via grid subsampling and pooling, which trades spatial resolution with feature resolution reducing the learning complexity without compromising the system performances.
+In contrast with regular domains, the subsampling operation on graphs or graph coarsening is not mathematically sound.
+It requires to construct meaningful neighborhoods on
+graphs where similar vertices are clustered together. Doing this for multiple
 layers is equivalent to construct a multi-scale clustering of the graph that
@@
 interested in multilevel clustering algorithms, where each level produces a
-\textit{coarser} graph which corresponds to the data domain seen at a different
-\textit{resolution}. Moreover, clustering techniques that reduce the size of the
+{coarser} graph which corresponds to the data domain seen at a different
+{resolution}. Moreover, clustering techniques that reduce the size of the
 graph by a factor two at each level offers a precise control on the coarsening
@@
 \cite{art:ShumanFarajiVandergheynst16PyramTrans} are two methods worth exploring
-in future works. The later has the main advantage to preserve the ordering of
-the Laplacian spectrum, and may be able to commute with the graph convolution
-operator. \todo{If proven, this would be a very essential mathematical property
-of CNNs on graphs, exactly as downsampling for classical CNNs.}
+in future works. %The later has the main advantage to preserve the ordering of
+%the Laplacian spectrum, and may be able to commute with the graph convolution
+%operator. \todo{If proven, this would be a very essential mathematical property
+%of CNNs on graphs, exactly as downsampling for classical CNNs.}
 
-Graclus \cite{art:DhillonGuanKulis07Graclus}, similarly to METIS, uses a greedy
+Graclus \cite{art:DhillonGuanKulis07Graclus}, built on Metis \cite{art:KarypisKumar98Metis}, uses a greedy
 algorithm to compute successive coarser versions of a given graph and is able to
@@
 cut \cite{art:ShiMalik00NCut}, which is an excellent clustering energy.
-Graclus' greedy rule consists, at any given coarsening level, in randomly
-pickling an unmarked vertex $i$ and matching it with one of its unmarked
+Graclus' greedy rule consists, at any given coarsening level, in
+picking an unmarked vertex $i$ and matching it with one of its unmarked
 neighbors $j$ that maximizes the local normalized cut $W_{ij} (1/d_i + 1/d_j)$.
@@
 explored. This is an extremely fast coarsening scheme which divides the number
-of nodes by approximately two (there exist a few singletons, non-matched nodes)
+of nodes by approximately two (there may exist a few singletons, non-matched nodes)
 from one level to the next coarser level.
 
+\subsection{Fast Pooling of Graph Signals}  \label{sec:pooling}
+As standard CNNs, pooling operations reduce the spatial resolution allowing higher filter resolution.
+Pooling operations, such as max pooling or average pooling, are carried out many times during the optimization, and thus must be as efficient as possible.
+After the graph coarsening operation in the previous section, the graph and its coarsened versions have an unstructured arrangement of the vertices.
+Hence, if the pooling is directly applied then it would need a table to store all matched vertices, which would result in memory consuming, slow, and not (easily) parallelizable pooling operations.
+
+It is however possible to structure the arrangement of the vertices, and achieve a graph pooling operation as efficient as a 1D grid pooling.
+We proceed as follows.
+After coarsening, each node has either two
+parents, if it was matched at the finer level, or one, if it was not, i.e the
+node was a singleton. From the coarsest to finest level, fake nodes, i.e.
+disconnected nodes, are added to pair with the singletons such that each node
+has two children. This structure is a balanced binary tree: (1)
+regular nodes (and singletons) either have two regular nodes (e.g. level 1
+vertex 0 in \figref{pooling}) or (2) one singleton and a fake node as children
+(e.g. level 2 vertex 0 in \figref{pooling}), and (3) fake nodes always have two
+fake nodes as children (e.g. level 1 vertex 1 in \figref{pooling}). Input
+signals are initialized with a neutral value at the fake nodes, e.g. 0 when
+using a ReLU activation with max pooling. Because these nodes are disconnected,
+filtering does not impact the initial neutral value. While those fake nodes do
+artificially increase the dimensionality thus the computational cost, we found
+that, in practice, the number of singletons left by Graclus is quite low
+% < 10%?
+compared to the number of vertices.
+
+Arbitrarily ordering the nodes at the coarsest level, then propagating this
+ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
+children, produces a regular ordering in the finest level. Regular in the sense
+that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
+rearranged graph signal is analog to pooling a regular 1D signal.
+\figref{pooling} shows an example of the whole coarsening and pooling process.
+This regular arrangement makes the pooling operation very efficient and
+satisfies parallel architectures such as GPUs as memory accesses are local, i.e.
+matched nodes do not have to be fetched.
+
 \begin{figure}[ht]
+\vspace{0.5cm}
 \centering
 \includegraphics[width=\textwidth]{images/coarsening_crop}
-\caption{\textbf{Example of Graph Coarsening and Pooling.} Let's say we want to
-perform a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
+\caption{\textbf{Example of Graph Coarsening and Pooling.} Let us carry out a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
 \R^8$ living on $\G_0$, the finest graph given as input. Note that it originally
@@
 
-\paragraph{Pooling of graph signals.} After coarsening, each node has either two
-parents, if it was matched at the finer level, or one, if it was not, i.e the
-node was a singelton. From the coarsest to finest level, fake nodes, i.e.
-disconnected nodes, are added to pair with the singletons such that each node
-has two children. This structure can be seen as a balanced binary tree: (1)
-regular nodes (and singletons) either have two regular nodes (e.g. level 1
-vertex 0 in \figref{pooling}) or (2) one singleton and a fake node as children
-(e.g. level 2 vertex 0 in \figref{pooling}), and (3) fake nodes always have two
-fake nodes as children (e.g. level 1 vertex 1 in \figref{pooling}). Input
-signals are initialized with a neutral value at the fake nodes, e.g. 0 when
-using a ReLU activation with max pooling. Because these nodes are disconnected,
-filtering won't impact the initial neutral value. While those fake nodes do
-artificially increase the dimensionality thus the computational cost, we found
-that, in practice, the number of singletons left by Graclus is quite low
-% < 10%?
-compared to the number of vertices.
-
-Arbitrarily ordering the nodes at the coarsest level, then propagating this
-ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
-children, produces a regular ordering in the finest level. Regular in the sense
-that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
-rearranged graph signal is analog to pooling a regular 1D signal.
-\figref{pooling} shows an example of the whole coarsening and pooling process.
-This regular arrangement makes the pooling operation very efficient and
-satisfies parallel architectures such as GPUs as memory accesses are local, i.e.
-matched nodes don't have to be fetched.
-
 \section{Numerical Experiments}
-
 All experiments were performed with TensorFlow, an open-source library for
@@
 In the sequel, we refer to filters defined by \eqnref{filt_non-param} as
-\textit{Non-Param} and filters defined by \eqnref{filt_cheby} as
+\textit{Non-Param}, i.e. the spatially non-localized filters, and the proposed filters defined by \eqnref{filt_cheby} as
 \textit{Chebyshev}. Filters referred to as \textit{Spline} are defined by
@@
 where $B \in \R^{n \times K}$ is the cubic B-spline basis and the parameter
-$\theta \in \R^K$ is a vector of control points, as proposed by \cite{\bruna}. It is
-important to note that, for all methods, we used the Graclus state-of-the-art
-coarsening algorithm introduced in \secref{pooling} rather than their "naive
-agglomerative method", as they put it. The rational is that we wanted to compare
-the filters, not the coarsening algorithms.
+$\theta \in \R^K$ is a vector of control points, as proposed in \cite{\bruna}. We use for all techniques the advanced Graclus
+coarsening algorithm introduced in \secref{coarsening} rather than the simple
+agglomerative method of \cite{\bruna}. The motivation is to compare
+the quality of learned filters, not the coarsening algorithms.
 
@@
 The final layer is always a softmax regression and the loss energy $E$ is the
-cross-entropy loss \todo{with an $\ell_2$ regularization on the weights of all
-FC$k$ layers}. Mini-batches are of size $S = 100$.  \todo{dropout}
+cross-entropy loss with an $\ell_2$ regularization on the weights of all
+FC$k$ layers. Mini-batches are of size $S = 100$.  %\todo{dropout}
 
-\todo{$k$-NN similarity graphs are built as follows: }
+%\todo{$k$-NN similarity graphs are built as follows: }
 
-%\subsection{Revisiting classical CNNs on MNIST}
-\subsection{Comparison with classical CNNs on MNIST}
+\subsection{Revisiting Standard CNNs on MNIST} \label{sec:MNIST}
+%\subsection{Comparison with classical CNNs on MNIST}
 % Struct: motivation, dataset, model, results
 
-To confirm the validity of our model, we first revisit the Euclidean
-case, where the graph is a regular grid, on the standard MNIST classification
+To validate our model, we first apply it to the Euclidean case with the benchmark MNIST classification
 problem \cite{pro:LeCunBottouBengioHaffner98MNIST}.
-This is an important sanity check for our model, which should be able to extract
-features on any graph, including the 2D grid graph on which images reside.
-\tabref{mnist} indeed shows that our model is able to learn localized filters
-and achieve a performance very close to that of a classical CNN with the same
-architecture.
+In this situation, the graph is simply a $k$-NN graph of the Euclidean 2D grid.
+This is an important sanity check for our model, which must be able to extract
+features on any graph, including the regular 2D grid on which images reside.
 
@@
 
-MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28
-\times 28$, such that data points lie on a space of $784$ dimensions. To test
-our model, we constructed a 8-NN (nearest neighbors) graph of the 2D Euclidean
-grid, which produced a graph of $n = |\V| = 976$ nodes (784 pixels and 192 fake
-nodes) and $|\E| = 3198$ edges.
+We recall that MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28
+\times 28$, such that data points lie on a space of $784$ dimensions. For
+our graph model, we construct a $8$-NN graph of the 2D Euclidean
+grid, which produces a graph of $n = |\V| = 976$ nodes (784 pixels and 192 fake
+nodes as explained in \secref{pooling}) and $|\E| = 3198$ edges.
 
-The LeNet-5-like network architecture and the following hyper-parameters are
+\tabref{mnist} confirms the ability of our model to learn localized filters
+as it achieves a performance very close to the classical CNNs with the same LeNet-5-like
+architecture. The LeNet-5-like network architecture and the following hyper-parameters are
 borrowed from the TensorFlow MNIST tutorial\footnote{
@@
 have filters of size $5 \times 5$ while graph convolution layers have the same
-support of $K = 5^2 = 25$. All models were trained for \todo{20 epochs}.
+support of $K = 5^2 = 25$. All models were trained for 20 epochs.
 
-\subsection{Text Categorization on 20NEWS}
-
-To demonstrate the usefulness of our model and the versatility of modeling with
-graphs, we applied our model to the text categorization problem on 20NEWS.
-\tabref{20news} indeed shows that CNNs on graphs are a good model which offers
-decent performances. While it does not outperform a simpler algorithm like a
-multinomial naive Bayes classifier on this small dataset, it does defeat fully
-connected networks, which require much more parameters to be learned. More
-importantly, these results show that the statistical assumptions made by the
-model are verified.
+\subsection{Text Categorization with 20NEWS}
+%usefulness of our model and its 
+To demonstrate the versatility of our model to work with graphs generated from unstructured data, we applied our technique to the text categorization problem with the 20NEWS dataset.
 
 \begin{table*}[h!] \centering
+\vspace{0.5cm}
 \begin{tabular}{lc} \toprule
@@
 produced a graph of \todo{$n = |\V| = $ nodes (10k words and ? fake nodes) and
-$|\E| = ?$ edges}. The influence of the graph quality is further studied in
-\secref{graph_quality}. All models were trained for \todo{20 epochs} by the Adam
-optimizer \todo{cite} with a learning rate of 0.1, without regularization nor
-dropout. The filter support was set to \todo{$K = 25$}.
+$|\E| = ?$ edges}.  All CNN models were trained for 20 epochs by the Adam
+optimizer \cite{art:KingmaBa14AdamOpt} with a learning rate of 0.1, without regularization nor
+dropout. The filter support was set to $K = 5$.
 
-\subsection{Definition of spectral filters}
+\tabref{20news} shows that CNNs on graphs provide decent performances.
+While it does not outperform the kernel-based method of multinomial naive Bayes classifier on this small dataset, it does defeat fully connected networks, which require much more parameters to be learned.
+More importantly, these results verify the validity of the statistical assumptions made for the data, that are locality and stationarity, and which are at the core of the design of any CNN technique.
+While we know that these data properties are true for low-dimensional Euclidean data like audios, images and videos, we also show experimentally that they are also satisfied for text documents as long as the graph is properly constructed.
+\secref{graph_quality} will study the influence of the graph quality.
 
-To compare our spectral filters with those proposed in \cite{\bruna}, we measure
-the performance of the graph CNN model for each filter definition on the MNIST
-and 20NEWS datasets. \tabref{filters} shows that our parametrization outperforms
-\cite{\bruna} as well as non-parametric filters, which are not localized and
-require $\bO(n)$ parameters to learn, in all cases.
+\subsection{Numerical Comparison between Spectral Filters}
+
+In this section, we compare the proposed spectral filters to the non-parametric filters, and those proposed in \cite{\bruna} on the MNIST and 20NEWS datasets.
+\tabref{filters} reports that the proposed kernel parametrization outperforms
+\cite{\bruna} as well as the non-parametric filters, which are not localized and require $\bO(n)$ parameters to learn.
 
 \begin{table*}[h!] \centering
+\vspace{0.5cm}
 \begin{tabular}{llccc} \toprule
@@
 
-\subsection{Computational efficiency}
+\subsection{Computational Efficiency}
 
-\figref{runtime} indeed demonstrates that the training time for our model scales
-as $\bO(n)$ while \cite{\bruna} scales as $\bO(n^2)$. Moreover,
+\figref{runtime} validates the low computational complexity of the proposed CNN technique on graphs.
+The training time of our model scales as $\bO(n)$, while \cite{\bruna} scales as $\bO(n^2)$.
+Moreover,
 \figref{convergence} gives a sense of how the validation accuracy as well as the
@@
 \begin{table*}[h!] \centering
+\vspace{0.5cm}
 \begin{tabular}{llccc} \toprule
@@
 \begin{figure}[h!]
+\vspace{0.5cm}
 \centering
-\includegraphics[width=5cm]{images/speed.png}
+\includegraphics[width=7cm]{images/speed.png}
 \caption{Training time over a mini-batch of 100 20NEWS documents w.r.t. $n$, the
@@
 
-\subsection{Influence of graph quality} \label{sec:graph_quality}
+\subsection{Influence of Graph Quality} \label{sec:graph_quality}
 
-For our method to be successful, his statistical assumptions shall be fulfilled
-on the graph where the data resides. The performance, and the quality of the
-learned filters, thus critically depends on the quality of the graph. By the
-success of classical CNNs, we know these assumptions to be true for images in
-Euclidean domains. By our experiments, we know also that a $k$-NN graph of the
-grid is good enough (we did not loose much compared to classical CNNs), and the
-value of $k$ does not have a strong influence. The extent to which a particular
-graph satisfies the assumptions is to compare its performance with a random
-graph, which for sure does not. \tabref{mnist_quality} shows that when using a
-random graph, the accuracy drops (\todo{to the order of an architecture without
-convolution}), as the convolutional layers are not useful anymore to extract
-meaningful features. The structure has been lost.
+For the proposed method to be successful, the statistical assumptions of  locality, stationarity, and compositionality regarding the data must be fulfilled on the graph where the data resides.
+Therefore, the classification performance, the quality of learned filters, all critically depends on the quality of the graph.
+For data lying on Euclidean space, experiments in
+ \secref{MNIST} show that a simple $k$-NN graph of the grid is good enough to recover almost exactly the performance of standard CNNs.
+We also noticed that the value of $k$ does not have a strong influence on the results.
+We can witness the importance of a graph satisfying the data assumptions by comparing its performance with a random
+graph.
+\tabref{mnist_quality} reports a large drop of accuracy when using a random graph, that is when the data structure is lost and the convolutional layers are not useful anymore to extract meaningful features.
+
+%the accuracy drops (\todo{to the order of an architecture without
+%convolution}), as the convolutional layers are not useful anymore to extract
+%meaningful features. The structure has been lost.
 
 \begin{table*}[h!]
+\vspace{0.5cm}
 \centering
@@
 GC32 & 97.40 & 96.88 \\
-GC32-P4-GC64-P4-FC512 & 98.65 & 95.39 \\
+GC32-P4-GC64-P4-FC512 & 98.65\todo{99.14} & 95.39 \\
 \bottomrule \end{tabular}
@@
 
-While images naturally lie on a graph, the 2D grid, that is not the case for
-text documents. For our method to be applied, a feature graph does have to be
-built.
+While images naturally lie on Euclidean 2D spaces, represented by graphs of regular grids, this is not the case for
+text documents. For our method to be applied to this class of data, a feature graph does have to be built.
 We investigate here 5 $16$-NN similarity graphs of words. The simplest option is
@@
 \begin{table*}[h!] \centering
+\vspace{0.5cm}
 \begin{tabular}{ccccc} \toprule
@@
 ability of the model to extract local and stationary features through the graph
-convolutional layers. Compared with the first generation of spectral graph CNNs
+convolutional layers.
+
+Compared with the first works of spectral graph CNNs
 introduced in \cite{\bruna}, our model provides a strict control over the local
-support of filters, is computationally more efficient (\figref{runtime}) by
-avoiding an explicit use of the Graph Fourier basis and experimentally shows a
-better test accuracy (\tabref{filters}).
+support of filters, is computationally more efficient by
+avoiding an explicit use of the Graph Fourier basis, and experimentally shows a
+better test accuracy. Besides, we addressed the three concerns raised by
+\cite{art:HenaffBrunaLeCun15DLgraphs}. First, we introduced a model whose computational complexity is linear with the
+dimensionality of the data.
+Second, we confirmed that the quality of the input graph is of paramount importance.
+Along this line, the idea of \cite{art:HenaffBrunaLeCun15DLgraphs} to learn the input graph from data is highly meaningful.
+And a natural and future approach would be to alternate the learning of the CNN parameters and the graph, as a virtuous circle.
+Third, we showed that the statistical assumptions of local stationarity and compositionality made by the model are verified for text documents as long as the graph is well constructed.
 
-In the end we addressed the three concerns raised by
-\cite{art:HenaffBrunaLeCun15DLgraphs}:
-\begin{itemize}
-\item We introduced a model whose computational complexity is linear with the
-	dimensionality of the data (\figref{runtime}).
-\item As they showed already, we confirm that the quality of the input graph is
-	of paramount importance (\tabref{20news_quality}).  Their idea to learn it
-	from data makes a lot of sense. A natural and future approach would be to
-	alternate the learning of the NN parameters and the graph, as a virtuous
-	circle.
-\item We showed that the statistical assumptions of stationarity and
-	compositionality made by the model are verified for text documents
-	(\tabref{20news}), as long as the graph is well constructed.
-\end{itemize}
-
-Future works will investigate two directions. First, we will explore
+Future works will investigate two directions. On one hand, we will explore
 applications of this generic model to important fields where the data naturally
@@
 artificially created networks which quality may vary as seen in the experiments.
-Second, we will also improve this framework with tools developed in GSP,
+On the other hand, we will also improve this framework with tools developed in GSP,
 including graph wavelet operators
@@
 
+%\todo{Model statistical assumptions: locality, stationarity, compositionality.}
+%These assumptions can be experimentally verified for any dataset. While we know
+%for them to be true for audio, images and videos, we show experimentally that
+%they might be satisfied for text documents as long as the graph is properly
+%constructed.
+%We also observe that
+%local stationarity is an assumption that is made about the data, and the CNNs
+%are fundamentally designed on this data hypothesis. However, we believe this
+%stationarity assumption to be present in many real-world data, as structured
+%data are usually composed of repetitive small patterns at different scales of
+%observation.
+
 % more data, natural graphs (no need to worry about construction, external info)
@@
 \bibliographystyle{plain}
-{\setstretch{0} \small
+%{\setstretch{0} \small
+%\bibliography{bib_nips16}
+%%\bibliography{bib_nips16_short}
+%}
 \bibliography{refs}
-%\bibliography{bib_nips16_short}
-}
 

### 2016-10-29 nips2016/nips2016.tex
nips v7: spacing and runtime pdf image

@@
 %\setlength{\bibsep}{0.2pt} %comment for arxiv
+
+\setlength{\textfloatsep}{20.0pt plus 2.0pt minus 2.0pt}
+\setlength{\floatsep}{20.0pt plus 2.0pt minus 2.0pt}
+\setlength{\intextsep}{20.0pt plus 2.0pt minus 2.0pt}
 
@@
 \begin{figure}[ht]
-\vspace{0.5cm}
 \centering
@@
 \begin{table*}[h!] \centering
-\vspace{0.5cm}
-\begin{tabular}{lc} \toprule
-Model & Accuracy \\
+\begin{tabular}{lcc} \toprule
+Model & Accuracy & \todo{Parameters} \\
 \midrule
@@
 %Multinomial logistic regression (softmax) & 66.28 \\
-Softmax & 66.28 \\
+Softmax & 66.28 & $200\times10^3$ \\
 \addlinespace
-FC2500 & 64.64 \\
-FC2500-FC500 & 65.76 \\
+FC2500 & 64.64 & $\sim25\times10^6$ \\
+FC2500-FC500 & 65.76 & $\sim26\times10^6$ \\
 \addlinespace
-GC32 & 68.26 \\
+GC32 & 68.26 5\\
 GC32-FC500 & \todo{?} \\
@@
 \cite{pro:MikolovChenCorradoDean13word2vec} embedding of those words, which
-produced a graph of \todo{$n = |\V| = $ nodes (10k words and ? fake nodes) and
-$|\E| = ?$ edges}.  All CNN models were trained for 20 epochs by the Adam
+produced a graph of $n = |\V| = 10,000$ nodes (no fake nodes because we do not
+coarsen the graph) and
+$|\E| = 132,834$ edges. All CNN models were trained for 20 epochs by the Adam
 optimizer \cite{art:KingmaBa14AdamOpt} with a learning rate of 0.1, without regularization nor
@@
 \begin{table*}[h!] \centering
-\vspace{0.5cm}
 \begin{tabular}{llccc} \toprule
@@
 \caption{Comparison of accuracy results for different types of spectral filters.
-Spline and Chebyshev filters use \todo{$K=20$}, \todo{$K=5$ for 20NEWS}.} 
+Spline and Chebyshev filters use \todo{$K=20$}, \todo{$K=5$ for 20NEWS}
+parameters.} 
 \label{tab:filters}
@@
 
-\begin{table*}[h!] \centering
-\vspace{0.5cm}
+\begin{table*}[ht] \centering
 \begin{tabular}{llccc} \toprule
@@
 \bottomrule \end{tabular}
-\caption{Time to process a mini-batch of 100 MNIST images. That is the total
+\caption{Time to process a mini-batch of $S=100$ MNIST images. That is the total
 training time divided by the number of gradient steps.} 
@@
 
-\begin{figure}[h!]
-\vspace{0.5cm}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=7cm]{images/speed.png}
-\caption{Training time over a mini-batch of 100 20NEWS documents w.r.t. $n$, the
+\vspace{-20pt}
+\includegraphics[width=\textwidth]{images/runtime_crop}
+\caption{Training time over a mini-batch of $S=100$ 20NEWS documents w.r.t. $n$, the
 number of selected words, for the three considered filters. The architecture is
@@
 \begin{table*}[h!]
-\vspace{0.5cm}
 \centering
@@
 \begin{table*}[h!] \centering
-\vspace{0.5cm}
 \begin{tabular}{ccccc} \toprule

### 2016-10-29 nips2016/images/runtime.py

@@ -1,0 +1,19 @@
+#!/bin/env python
+
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+matplotlib.rcParams.update({'font.size': 18})
+
+fourier = np.array([10,23,68,139.5,231.5,351,495])/400*1000
+cheby = np.array([7.5,11.5,20,29.5,38,48.5,59])/400*1000
+x = np.array([1,2,4,6,8,10,12]) * 1000
+
+plt.figure(figsize=(15,5))
+plt.plot(x, cheby, '.-', markersize=15, label='Chebyshev');
+plt.plot(x, fourier, '.--', markersize=15, label='Non-Param / Spline');
+plt.legend(loc='upper left')
+plt.xlim(1000,12000)
+plt.xlabel('number of features (words)')
+plt.ylabel('time (ms)')
+plt.savefig('runtime.pdf')

### 2016-10-29 nips2016/nips2016.tex
nips v8: small edition (Xavier) (2016-06-30)

--- 2016-10-29 nips2016/nips2016.tex
@@
 
+%\newpage
 The compositional local stationarity property of CNNs can be efficiently
-implemented on regular grids, like image and sound domains, as convolutional and
-downsampling operators are mathematically well-defined on such discretized
+implemented on regular grids, like image and sound domains, as convolutional, 
+downsampling and pooling operators are mathematically well-defined on such discretized
 Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
@@
 important examples of data lying on irregular or non-Euclidean domains.
-Unfortunately, CNNs cannot be directly applied to such complex domains.
+Unfortunately, CNNs cannot be directly applied to such complex domains.% as the building blocks of CNNs, convolution, downsampling, and pooling, are only defined for regular grids.
 
-A generalization to irregular grids is not straightforward as the building blocks of
-CNNs, the convolution, the downsampling, and the pooling operators, cannot be directly applied to these domains. This makes this extension challenging, both theoretically and
-implementation-wise. This work leverages the recent works of
+%A generalization to irregular grids is not straightforward as the building blocks of CNNs, convolution, downsampling, and pooling, are only defined for regular grids.
+
+A generalization to irregular grids is not straightforward as the standard operators of convolution, downsampling, and pooling, are only defined for regular grids.
+This makes this extension challenging, both theoretically and implementation-wise.
+This work leverages the recent works of
 \cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
@@
 %Softmax & 92.36  \\
-Classical CNN & C32-P4-C64-P4-FC512 & \textbf{99.20} \todo{99.33} \\
-Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 98.65 \todo{99.14} \\
+Classical CNN & C32-P4-C64-P4-FC512 & \textbf{99.33}  \\
+Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 99.14  \\
 \bottomrule
@@
 \begin{table*}[h!] \centering
-\vspace{0.5cm}
 \begin{tabular}{lc} \toprule
@@
 GC32 & 68.26 \\
-GC32-FC500 & \todo{?} \\
+%GC32-FC500 & \todo{?} \\
 \bottomrule \end{tabular}
@@
 \cite{pro:MikolovChenCorradoDean13word2vec} embedding of those words, which
-produced a graph of \todo{$n = |\V| = $ nodes (10k words and ? fake nodes) and
-$|\E| = ?$ edges}.  All CNN models were trained for 20 epochs by the Adam
+produced a graph of $n = |\V| = 10,000$ nodes and
+$|\E| = 132,834$ edges.  All CNN models were trained for 20 epochs by the Adam
 optimizer \cite{art:KingmaBa14AdamOpt} with a learning rate of 0.1, without regularization nor
@@
 MNIST & GC10 & 95.75 & 97.26 & \textbf{97.48} \\
-MNIST & GC32-P4-GC64-P4-FC512 & 96.28 & 97.15 & \textbf{98.65} \\
-20NEWS & GC32 & \todo{?} & \todo{?} & \todo{?} \\
+MNIST & GC32-P4-GC64-P4-FC512 & 96.28 & 97.15 & \textbf{99.14} \\
+%20NEWS & GC32 & \todo{?} & \todo{?} & \todo{?} \\
 \bottomrule \end{tabular}
 \caption{Comparison of accuracy results for different types of spectral filters.
-Spline and Chebyshev filters use \todo{$K=20$}, \todo{$K=5$ for 20NEWS}.} 
+Spline and Chebyshev filters use $K=25$.} % \todo{$K=5$ for 20NEWS}
 \label{tab:filters}
@@
 
-\begin{figure}[h!]
-\vspace{0.5cm}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=7cm]{images/speed.png}
+\vspace{-20pt}
+\includegraphics[width=\textwidth]{images/runtime_crop}
 \caption{Training time over a mini-batch of 100 20NEWS documents w.r.t. $n$, the
@@
 GC32 & 97.40 & 96.88 \\
-GC32-P4-GC64-P4-FC512 & 98.65\todo{99.14} & 95.39 \\
+GC32-P4-GC64-P4-FC512 & 99.14 & 95.39 \\
 \bottomrule \end{tabular}
@@
 \midrule
-67.50 & \todo{66.98} & 68.26 & 67.86 & 67.75 \\
+67.50 & 66.98 & 68.26 & 67.86 & 67.75 \\
 \bottomrule \end{tabular}

### 2016-10-29 nips2016/nips2016.tex
arxiv v1 (2016-06-30)

@@
+\pdfoutput=1
+
 \documentclass{article}
@@
 \usepackage{color}
+
+\usepackage{amsmath}
 
@@
 
-	Convolutional neural networks (CNNs) have greatly improved state-of-the-art
-	performances in a number of fields, notably computer vision and natural
-	language processing. In this work, we are interested in generalizing the
-	formulation of CNNs from low-dimensional regular Euclidean domains, where
-	images (2D), videos (3D) and audios (1D) are represented, to high-dimensional
-	irregular domains such as social networks or biological networks represented by graphs. This paper introduces a formulation
-	of CNNs on graphs in the context of spectral graph theory. We borrow the
-	fundamental tools from the emerging field of signal processing on graphs,
-	which provides the necessary mathematical background and efficient numerical
-	schemes to design localized graph filters efficient to learn and evaluate. 
+Convolutional neural networks (CNNs) have greatly improved state-of-the-art performances in a number of fields, notably computer vision and natural language processing.
+In this work, we are interested in generalizing the formulation of CNNs from low-dimensional regular Euclidean domains, where images (2D), videos (3D) and audios (1D) are represented, to high-dimensional irregular domains such as social networks or biological networks represented by graphs.
+This paper introduces a formulation of CNNs on graphs in the context of spectral graph theory.
+We borrow the fundamental tools from the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design localized graph filters efficient to learn and evaluate.
 As a matter of fact, we introduce the first technique that offers the same computational complexity than standard CNNs, while being universal to any graph structure.
-	%Similarly to classical CNNs, the computational complexity as well as the
-	%number of parameters to learn are independent of the data dimensionality.
-	Numerical experiments on MNIST and 20NEWS demonstrate the ability of this novel deep learning system to learn local, stationary, and compositional features on graphs, as long as the graph
-	is well-constructed. %This work can be thought as the foundation of a second
-	%generation of CNNs on graphs built on spectral graph theory, being
-	%mathematically well defined, bringing better test performance and being
-	%faster to evaluate and train than the first generation introduced by Bruna,
-	%LeCun et al.
+Numerical experiments on MNIST and 20NEWS demonstrate the ability of this novel deep learning system to learn local, stationary, and compositional features on graphs, as long as the graph is well-constructed.
 
@@
 \centering
-\includegraphics[width=\textwidth]{images/illustrationCNNgraphs.eps}
-%\includegraphics[width=13.5cm]{images/illustrationCNNgraphs.png}
+\includegraphics[width=\textwidth]{images/illustrationCNNgraphs}
 \caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
@@
 %Softmax & 92.36  \\
-Classical CNN & C32-P4-C64-P4-FC512 & \textbf{99.33}  \\
+Classical CNN & C32-P4-C64-P4-FC512 & 99.33  \\
 Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 99.14  \\
@@
 \midrule
-MNIST & GC10 & 95.75 & 97.26 & \textbf{97.48} \\
-MNIST & GC32-P4-GC64-P4-FC512 & 96.28 & 97.15 & \textbf{99.14} \\
+MNIST & GC10 & 95.75 & 97.26 & {97.48} \\
+MNIST & GC32-P4-GC64-P4-FC512 & 96.28 & 97.15 & {99.14} \\
 %20NEWS & GC32 & \todo{?} & \todo{?} & \todo{?} \\

### 2016-10-29 nips2016/review/rebuttal_michael.txt
rebuttal michael (2016-07-21)

Todo
* compare with GNN
* read recent spatial approaches
* constructed graph on MNIST
* run on RCV1
* review "?" in text

Acceptance
* oral: ~2%
* spotlight: ~5%
* poster: ~22%

Private

It appears to us that we got good reviews from reviewers who consider themselves
experts or are confident. Less confident reviewers seem to have not understood
the paper, thus its potential. E.g. they ask questions whose answers are in the
paper already.

General

We thank all the reviewers for reading the paper and providing valuable
feedback. The specific comments of each reviewer are addressed below.

We already cleaned up and edited the paper in the hope to make it more
accessible. We'll take into account the feedbacks for the final revision.

We plan to publicly release the code once the paper is published so that people
can reproduce our results, compare with our model or apply it to their problems.

Note that very recent works [post-submission refs?] (posted on arXiv after the
submission deadline) shows the growing interest for a generalization of CNNs to
arbitrary graphs. While both these approaches are spatial, we look forward to
compare with them.

Reviewer 1: comparison with GNN.

GNN does embed the graph into an high-dimensional space.
Limited to combinatorial Laplacian, no normalized?.
Classify graphs (graph-focused) or nodes (node-focused).
Thanks for pointing this up. We'll add this discussion to the final version of
the paper.

First, the goal of GNNs is to classify graphs or nodes, while the proposed
method classifies signals on graphs. Then, each approach could be twisted and we
could treat a graph signal as labels, but then there'll be no pooling.
Our method may be used to classify nodes, by treating a graph signal as a set of
labels, but we don't see how GNNs could be used to classify graph signals.

Note suited for multiple signals on graph. But for features who leave on nodes.

GNNs is designed as a spatial method.

Second, GNNs need to perform as many backpropagation operations as there are
edges and nodes in the graph per learning iteration. Complexity of

Graph neural networks (GNNs) (Scarselli et al., 2009) are a recurrent neural
network architecture defined on graphs. GNNs apply recurrent neural networks for
walks on the graph structure, propagating node representations until a fixed
point is reached. The resulting node representations are then used as features
in classification and regression problems. GNNs support only discrete labels and
perform as many backpropagation operations as there are edges and nodes in the
graph per learning iteration. Gated Graph Sequence Neural Networks modify GNNs
to use gated recurrent units and to output sequences (Li et al., 2015).

One back-propagation per output: each node embedding is an output.

Reviewer 1: rotational invariance.

Your analysis is correct, the proposed filters are isotropic, i.e. they are
oblivious to the orientation of an edge. The learned filters therefore provide
rotational invariance. Whether this is a limitation or an asset depends on the
problem and should be verified empirically, as for any invariance (translation,
scale). In an experiment (not in the paper) where we did randomly rotate (from
-90° to 90°) each MNIST digit (no data augmentation), we obtained 91.47%
accuracy while classical CNNs got 70.65%. Rotational invariance can therefore be
an asset. Moreover, data augmentation schemes have used rotated versions of
(part of) images to force a network to learn this invariance, and Spatial
Transformer Networks [1], who learn invariance to translation, scale, rotation
and more generic warping, did show state-of-the-art performance on several
benchmarks. We did not include these preliminary results in the paper as we want
to test on larger image datasets the impact of this invariance. As part of this
future work, we'll investigate the possibility to build our filters on
anisotropic Laplacians [2], which should allow us to learn directional /
oriented filters.

Reviewer 3.

The limited results in Table 4 could be due to the small size (~19k documents)
of the 20NEWS dataset. We performed the same task on RCV1 (800k documents), and
obtained...
Unfortunately after the submission deadline. Or say that Bruna did it.

Reviewer 6.

Thanks for your encouragement. Please note that most of the recent works are
based on a spatial formulation. What are the limitations? We look forward to
compare all these approaches.

Reviewer 7: single graph.

As you rightfully pointed out, our spectral construction does not transfer
directly to other graphs. That is an advantage of spatial constructions?. We do
however believe that graphs from a certain class (e.g. social graphs) share a
similar spectrum, which would allow to learn and transfer the Chebyshev
coefficients across graphs. That is a planned future work.?

Reviewer 7: pooling on graphs.

A pooling strategy on arbitrary graphs will always? require some sort of
coarsening, i.e. a way to group similar vertices. As we pointed out, there
exists plenty of methods for multi-level clustering. The proposed CNN on graphs
is agnostic to the choice of a particular algorithm, and we hope that future
research in the area will bring better algorithms. Moreover, the coarsening step
only has to be performed once, before training.

Reviewer 7: MNIST experiment.

You are right, we did experiment with a regular grid on MNIST because it is the
sole way to compare with classical CNNs. The ~3% performance loss you mention
(from Table 3 I guess?) is caused by the use of random graph, i.e. random
connections between pixels. It is clear that in this setting we expect to loose
much performance, which indeed shows that the grid structure is important for
images. We could also build a graph from the pixel values only, which would
approach a regular grid?. Performances are obviously not state-of-the-art, we
only wanted to compare with the classical LeNet-5 and not bother with model
tuning.

Reviewer 7 & 8: 20NEWS experiment, usefulness of the graph.

We did compare with baselines only. The small performance improvement is
probably because the constructed graph is not good enough. There is probably a
better way. Constructing graphs from data is an open research question. This is
however a non-issue when the data naturally lies on a graph structure. Then why
didn't we do it ??

The usefulness of the graph structure obviously depends on the quality of the
graph. That is the point we tried to make in section 6.2: if a given graph does
not improve much on a random graph, then this graph is not well suited for the
problem at hand. If the graph is well constructed, e.g. the grid for MNIST, then
results are good. Moreover, the very recent works in this area [refs] suggest
that this is a research direction of interest.

Reviewer 8 & 9: understanding.

As you pointed out, you do not seem to be familiar with the field of Graph
Signal Processing. Due to space constraints, we had to write densely and
specifically pointed out a review of the field [29]. Moreover, most of the
reviewers seem to have understood the paper well. We did however clean the paper
up and hope it is easier to read.

Reviewer 8: performance.

Our method is both faster and more effective, see Table 6 (Spline filters), than
[13]. The theoretical difference is that our filters are exactly localized in K
hops.

Reviewer 9: natural graphs.

The problem with signals on natural graphs is that, at least for now, not many
datasets are public, and when they are, they are quite small. ?? We do however
plan to make experiments on Wikipedia, using hyperlinks and user activity to
build a graph. This will however require us to build a new dataset.

[1] Spatial Transformer Networks 
[2] anisotropic 
[3] Rotation-invariant convolutional neural networks for galaxy morphology prediction

### 2016-10-29 nips2016/review/rebuttal_xavier.txt
rebuttal xavier (2016-07-21)

2763

We warmly thank the referees for their time and their helpful comments regarding our manuscript.

MNIST experiments (R1,R3,R7): The accuracy gap between standard CNN and graph CNN can be reduced with more epochs. With 20 epochs (instead of 10 in our submission) and LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%. This reveals a slower convergence of graph CNN, and future work will investigate better optimization strategies.

R1: Thanks for pointing out the GNN model (we will add this ref). GNN processes data in the spatial/graph domain, unlike our model that is defined in the spectral domain. It is not direct to formulate GNN in spectral form (and therefore to compare both techniques), that is, how to define spectral filters g s.t. Psi(x,y)=g(L)x (not only identity filter g(z)=z)? 
The spectral filters are indeed rotationally invariant. We think that this property might actually be an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN. We will explore this idea in future work.

R7: We indeed consider a single graph for this work, which is already meaningful for multiple applications s.a. brain connectivity analysis, and social network analysis (one graph and many data signals on this graph). Extension to multiple graphs is obviously of great interest, and can also be explored with a spectral technique like the one proposed here. We reserve this extension for another paper. 
The pooling strategy is inevitable in any CNN architecture as it allows to reduce spatial resolution and to increase filter resolution. In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering, which is NP-hard. We opted here for the Graclus model, which has proved to be a very efficient graph clustering technique.
It is expected that accuracy drops (3%) when the graph structure is destroyed with a random graph, as convolutional layers become meaningless. 
For 20NEWS, we compare against state-of-the-art Multinomial NB technique, and other baseline techniques.

R8: We have recently improved the presentation, accessibility, and Section 4 of the paper. We will be happy to release it.
When random graphs are used, then the convolutional layers become meaningless. However, the last part of the CNN system, the fully connected layers stay active, and absorbs the drop in performance. 
There are plenty of theoretical works in the area of spectral analysis, see e.g. the review [29].
The complexity of our technique is O(EK), which reduces to linear O(n) in case of sparse graphs, vs O(n^2) for [13].

R9: For the presentation of the paper, see our answer to R8.
We did not present any natural graph experiments due to the space limitation.

### 2016-10-29 nips2016/review/rebuttal.txt
rebuttal v1 (2016-07-21)

6500

############
ALL
############

We warmly thank the referees for their time and their valuable comments regarding our manuscript.

We first address a common comment from Referees 1,3,7.

MNIST experiments: The accuracy gap between standard CNN and graph CNN can be reduced with more epochs. With 20 epochs (instead of 10 in our submission) and a LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%, bringing closer the accuracy gap between graph and classic CNN. This reveals a slower convergence of graph CNN, and future work will investigate better optimization strategies. ISOTROPIC?

We now address the specific comments of each referee.

Referee 1.
GNN: Thanks for pointing out the GNN model (we will add this ref). On the high level, GNN processes data in the spatial/graph domain, unlike our model that is defined in the spectral domain. Technically, GNN was originally designed to classify nodes or graphs by first embedding each node in an Euclidean space with RNNs, and then using those as features for classification or regression. Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN. Regarding the referee’s comment, we read very carefully the two 2009 GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee. We would be grateful if the referee could point out more connections between our model and GNN.
Rotational invariance: The spectral filters are indeed rotationally invariant. Although it first appears to be a limitation of this system, we think the rational invariance property might actually be an advantage for large-scale image datasets. Our motivation is inspired by the following experiment (not reported in the paper): We learned the classic and graph CNN from a rotated MNIST dataset, that is each training MNIST image is randomly rotated from -90° to 90° (no data augmentation here, simply rotation of original images). Classic CNN obtained 70.65% of accuracy, and graph CNN got 91.47%. Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features. We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN. 
We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.

Referee 3.
Table 4 XXX: The goal of this table was to show that convolutional layers, model CN32-softmax, are also effective on graphs. This is justified by comparing to fully connected networks, models FC. 
The performance of graph CNN did not overcome the state-of-the-art? text document classification technique, namely Multinomial NB. This may due to the small size of the dataset, 20NEWS only having 19K documents. CNN usually perform less than other classification techniques for small datasets. As a consequence, we will apply our technique to the larger text dataset RCV1 which has 800K documents, as [13].

Referee 7.
Single graph: We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a. brain connectivity analysis, and social network analysis (one graph and many data signals on this graph). Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (for example learn and transfer Chebyshev coefficients across graphs). We reserve this extension for another paper. 
Graph pooling: Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution. In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e. grouping similar vertices together, which is NP-hard. AGNOSTIC We opted here for the Metis/Graclus model [8,15], which has proved to be a very efficient graph clustering technique. Note also that the coarsening step is performed only once, before training.
Random graph: It is clearly expected that classification accuracy drops (3%) when the graph structure is inconsistent, i.e. for random graphs (graphs that hold random connexions between pixels), as convolutional layers become meaningless. It shows that the regular grid structure, exploiting by standard CNN, is obviously of first importance for images. Eventually, note that our technique has the flexibility to exploit additional information like pixel values through the arbitrary graph definition, unlike classic CNN.
20NEWS XXX: For this dataset, we compare against the state-of-the-art text document classification technique, namely Multinomial NB, and the SVM and linear classifier baseline techniques. As illustrated in Table 5, classification performances depend on the quality of the graph construction. However, we did not try to optimize the graph in our experiments. Pls, see also our answer to Referee 3.

Referee 8.
Presentation: We have recently improved the presentation and accessibility of the paper. We will be happy to release this new version in the final version of the paper.
Graph usefulness XXX: The goal of Section 6.2 and Table 5 is to show that the usefulness of the graph structure directly depends on the quality of the graph. When random graphs are used, then the convolutional layers become meaningless and graph CNN are not well suited for the problem at hand. On the contrary, if the graph is well-constructed, like the grid for MNIST, then graph CNN will provide good performances. 
Small performance drop: XXX For random graphs, convolutional layers are not useful. However, the last part of the CNN system, the fully connected layers stay active, and absorbs the drop in performance. 
Theory: XXX PERFORMANCES There are plenty of theoretical works in the area of spectral analysis for graphs and signal processing, see e.g. the review [29].
Comparison to [13]: Our technique is both faster and more accurate. See Table 6 with [13] being referred as Spline Filters. Besides, a theoretical advantage of our model is that our filters are exactly localized in K hops on graphs.

Referee 9.
Presentation: See our answer to Referee 8.
Natural graphs: Indeed, we did not present any natural graph experiments in this work. Our first goal was to introduce our efficient graph CNN technique. However, a data analysis project on Wikipedia using this technique is ongoing.

############
Private to AC
############

The general consensus of the referees lean towards acceptance. Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful. The only concerns come from Referees 8,9 whom acknowledge their low confidence in the understanding of the main contributions of the paper. XXX Besides, questions of Referee 8 about theoretical background and effectiveness of the algorithm are repeatedly provided throughout the paper like in Section 2 and Section 6.3.

### 2016-10-29 nips2016/review/rebuttal.txt
rebuttal v2 (2016-07-21)

@@
-6500
+############
+Private to AC
+############
+
+The general consensus of the referees lean towards acceptance.
+Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful.
+The only concerns come from Referees 8,9 who acknowledge their low confidence in the understanding of the main contributions of the paper.
 
@@
 
+5900
+
 We warmly thank the referees for their time and their valuable comments regarding our manuscript.
 
-We first address a common comment from Referees 1,3,7.
+We first address common comments.
 
-MNIST experiments: The accuracy gap between standard CNN and graph CNN can be reduced with more epochs.
-With 20 epochs (instead of 10 in our submission) and a LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%, bringing closer the accuracy gap between graph and classic CNN.
-This reveals a slower convergence of graph CNN, and future work will investigate better optimization strategies.
-ISOTROPIC?
+Accuracy Gap on MNIST (Referees 1,3,7): There is indeed an accuracy gap between standard CNN and graph CNN.
+It may due to several reasons such as architecture tuning and isotopic filters.
+We also observed that the accuracy gap can be reduced with more epochs.
+With 20 epochs (instead of 10 in our submission) and a LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%.
+This reveals a slower convergence of graph CNN, and we will investigate better optimization strategies.
+
+Graph Usefulness and Experiments (R3,7,8): The goal of Tables 3 and 5 is to show that the usefulness of the graph structure directly depends on the quality of the graph.
+When random graphs are used (graphs that hold random connexions between pixels/vertices), then the convolutional layers become meaningless, and graph CNN are not well suited for the problem at hand.
+On the contrary, if the graph is well constructed, like the regular grid for MNIST, then graph CNN provide good performances.
+For 20NEWS, we did not try to optimize the graph structure in our experiments.
+There may be better ways to construct higher-quality graphs.
+However, constructing graphs from data is an open and challenging question, and was not our objective here.
+The goal of Table 4 was to demonstrate that convolutional layers, model CN32-softmax, are also effective on graphs that are not regular image grids, but a graph of 20NEWS words.
+This is justified by comparing CN32-softmax to fully connected networks, FC models.
+We also compared against the effective and popular text document classification technique, Multinomial NB, and the SVM and linear classifier baseline techniques.
+Graph CNN did not overcome Multinomial NB.
+It may due to the small size of the dataset, 20NEWS having only 19K documents.
+CNN usually perform less than other classification techniques for small datasets.
+As a consequence, we will apply our technique to the larger text dataset RCV1 which has 800K documents, as [13].
 
@@
 GNN: Thanks for pointing out the GNN model (we will add this ref).
-On the high level, GNN processes data in the spatial/graph domain, unlike our model that is defined in the spectral domain.
-Technically, GNN was originally designed to classify nodes or graphs by first embedding each node in an Euclidean space with RNNs, and then using those as features for classification or regression.
+On the high level, GNN is a spatial/graph technique, unlike our model which is spectral.
+Technically, GNN was originally designed to classify nodes or graphs by first embedding each node in an Euclidean space with RNN, and then using those as features for classification or regression.
 Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN.
 Regarding the referee’s comment, we read very carefully the two 2009 GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee.
-We would be grateful if the referee could point out more connections between our model and GNN.
+We would be grateful if the referee could point out more connections between our model and the GNN papers.
 Rotational invariance: The spectral filters are indeed rotationally invariant.
 Although it first appears to be a limitation of this system, we think the rational invariance property might actually be an advantage for large-scale image datasets.
-Our motivation is inspired by the following experiment (not reported in the paper): We learned the classic and graph CNN from a rotated MNIST dataset, that is each training MNIST image is randomly rotated from -90° to 90° (no data augmentation here, simply rotation of original images).
+Our motivation is inspired by the following experiment (not reported in the paper).
+We learned the classic and graph CNN from a rotated MNIST dataset, that is each training MNIST image is randomly rotated from -90° to 90° (no data augmentation here, simply rotation of original images).
 Classic CNN obtained 70.65% of accuracy, and graph CNN got 91.47%.
@@
 We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.
-
-Referee 3.
-Table 4 XXX: The goal of this table was to show that convolutional layers, model CN32-softmax, are also effective on graphs. This is justified by comparing to fully connected networks, models FC. 
-The performance of graph CNN did not overcome the state-of-the-art?
-text document classification technique, namely Multinomial NB.
-This may due to the small size of the dataset, 20NEWS only having 19K documents.
-CNN usually perform less than other classification techniques for small datasets.
-As a consequence, we will apply our technique to the larger text dataset RCV1 which has 800K documents, as [13].
 
@@
 grouping similar vertices together, which is NP-hard.
-AGNOSTIC We opted here for the Metis/Graclus model [8,15], which has proved to be a very efficient graph clustering technique.
+We opted here for the Metis/Graclus model [8,15], which has proved to be a very efficient graph clustering technique.
 Note also that the coarsening step is performed only once, before training.
-Random graph: It is clearly expected that classification accuracy drops (3%) when the graph structure is inconsistent, i.e.
-for random graphs (graphs that hold random connexions between pixels), as convolutional layers become meaningless.
-It shows that the regular grid structure, exploiting by standard CNN, is obviously of first importance for images.
-Eventually, note that our technique has the flexibility to exploit additional information like pixel values through the arbitrary graph definition, unlike classic CNN.
-20NEWS XXX: For this dataset, we compare against the state-of-the-art text document classification technique, namely Multinomial NB, and the SVM and linear classifier baseline techniques.
-As illustrated in Table 5, classification performances depend on the quality of the graph construction.
-However, we did not try to optimize the graph in our experiments.
-Pls, see also our answer to Referee 3.
 
@@
 Presentation: We have recently improved the presentation and accessibility of the paper. We will be happy to release this new version in the final version of the paper.
-Graph usefulness XXX: The goal of Section 6.2 and Table 5 is to show that the usefulness of the graph structure directly depends on the quality of the graph.
-When random graphs are used, then the convolutional layers become meaningless and graph CNN are not well suited for the problem at hand.
-On the contrary, if the graph is well-constructed, like the grid for MNIST, then graph CNN will provide good performances.
-Small performance drop: XXX For random graphs, convolutional layers are not useful.
-However, the last part of the CNN system, the fully connected layers stay active, and absorbs the drop in performance.
-Theory: XXX PERFORMANCES There are plenty of theoretical works in the area of spectral analysis for graphs and signal processing, see e.g. the review [29].
-Comparison to [13]: Our technique is both faster and more accurate.
-See Table 6 with [13] being referred as Spline Filters.
+Theory: The analysis of the performances of standard CNN is currently a central and open mathematical question.
+It is thus challenging to state anything about the theoretical performances of our model vs [13].
+The only performance comparison is experimental and are summarised in Table 6.
 Besides, a theoretical advantage of our model is that our filters are exactly localized in K hops on graphs.
@@
 Referee 9.
-Presentation: See our answer to Referee 8.
+Presentation: See our answer to R8.
 Natural graphs: Indeed, we did not present any natural graph experiments in this work.
 Our first goal was to introduce our efficient graph CNN technique.
-However, a data analysis project on Wikipedia using this technique is ongoing.
-
-############
-Private to AC
-############
-
-The general consensus of the referees lean towards acceptance.
-Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful.
-The only concerns come from Referees 8,9 whom acknowledge their low confidence in the understanding of the main contributions of the paper.
-XXX Besides, questions of Referee 8 about theoretical background and effectiveness of the algorithm are repeatedly provided throughout the paper like in Section 2 and Section 6.3.
+However, a data analysis project on Wikipedia network using this technique is ongoing.

### 2016-10-29 nips2016/review/rebuttal.txt
rebuttal v3 (2016-07-22)

@@
+############
+ALL
+############
+
+4950
+
+We warmly thank the referees for their time and valuable comments.
+
+Referees 1,3,7
+Accuracy Gap for MNIST: The gap between standard CNN and graph CNN may be due to our isotopic filters.
+We also observed that with 20 epochs (instead of 10 in our submission) for LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%.
+This reveals a slower convergence of graph CNN, and the need to investigate better optimization strategies.
+
+Referees 3,7,8
+Graph Usefulness and Experiments: The goal of Tables 3,5 is to show that the usefulness of the graph structure directly depends on the graph quality.
+When random graphs are used, then convolutional layers are useless, and graph CNN are not well suited for the problem.
+If the graph is well constructed, like the grid for MNIST, then graph CNN provide good performances.
+For 20NEWS, we did not try to optimize the graph structure.
+There may be better ways to construct higher-quality graphs, but it was not our objective.
+The goal of Table 4 is to demonstrate that convolutional layers CN32 are also effective on graphs that are not regular image grids.
+This is showed by comparing CN32 to fully connected (FC) networks.
+We also compared against the effective Multinomial NB document classification technique, and SVM and linear classifier baseline techniques.
+Graph CNN did not overcome Multinomial NB.
+It may be due to the small size of the dataset (20NEWS has 19K documents).
+CNN usually perform less than other classification techniques for small datasets.
+We plan to apply our technique to the larger text dataset RCV1 (800K documents).
+
+Referee 1
+Thanks for pointing out GNN (we will add this ref).
+On the high level, GNN is a spatial technique, and our model is spectral.
+GNN was designed to classify nodes/graphs by first embedding each node in an Euclidean space with RNN, and then using those as features for classification.
+Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN.
+Regarding the referee’s comment, we read carefully the two GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee.
+We would be grateful if the referee could point out more connections between our model and the GNN papers.
+The spectral filters are indeed rotationally invariant.
+Although it first appears to be a limitation, we think the rotational invariance property might actually be an advantage.
+Our motivation is inspired by the following experiment (not reported in the paper).
+We learned the classic and graph CNN from a rotated MNIST dataset, i.e.
+each training image is randomly rotated (no data augmentation here, simply rotation of original images).
+Classic CNN obtained 70% of accuracy, and graph CNN 91%.
+Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features.
+We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN.
+We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.
+
+Referee 7
+We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a.
+brain connectivity analysis, and social network analysis.
+Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (e.g.
+learn and transfer Chebyshev coefficients across graphs).
+We reserve this extension for another paper.
+Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution.
+In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e.
+grouping similar vertices together, which is NP-hard.
+We opted here for the Metis model, which has proved to be very efficient.
+Note also that the coarsening step is performed only once.
+
+Referee 8
+We have recently improved the presentation of the paper, and will release the new version in the final version of the paper.
+The performance analysis of standard CNN is a central and open mathematical question.
+It is thus challenging to state anything about the theoretical performances of our model.
+The only performance study is experimental and summarised in Table 6.
+
+Referee 9
+Presentation: See our answer to R8.
+We indeed did not introduce any natural graph experiments in this work.
+Our first goal was to introduce an efficient graph CNN technique.
+However, a data analysis project on Wikipedia network using this technique is ongoing.
+
 ############
@@
 Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful.
-The only concerns come from Referees 8,9 who acknowledge their low confidence in the understanding of the main contributions of the paper.
-
-############
-ALL
-############
-
-5900
-
-We warmly thank the referees for their time and their valuable comments regarding our manuscript.
-
-We first address common comments.
-
-Accuracy Gap on MNIST (Referees 1,3,7): There is indeed an accuracy gap between standard CNN and graph CNN.
-It may due to several reasons such as architecture tuning and isotopic filters.
-We also observed that the accuracy gap can be reduced with more epochs.
-With 20 epochs (instead of 10 in our submission) and a LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%.
-This reveals a slower convergence of graph CNN, and we will investigate better optimization strategies.
-
-Graph Usefulness and Experiments (R3,7,8): The goal of Tables 3 and 5 is to show that the usefulness of the graph structure directly depends on the quality of the graph.
-When random graphs are used (graphs that hold random connexions between pixels/vertices), then the convolutional layers become meaningless, and graph CNN are not well suited for the problem at hand.
-On the contrary, if the graph is well constructed, like the regular grid for MNIST, then graph CNN provide good performances.
-For 20NEWS, we did not try to optimize the graph structure in our experiments.
-There may be better ways to construct higher-quality graphs.
-However, constructing graphs from data is an open and challenging question, and was not our objective here.
-The goal of Table 4 was to demonstrate that convolutional layers, model CN32-softmax, are also effective on graphs that are not regular image grids, but a graph of 20NEWS words.
-This is justified by comparing CN32-softmax to fully connected networks, FC models.
-We also compared against the effective and popular text document classification technique, Multinomial NB, and the SVM and linear classifier baseline techniques.
-Graph CNN did not overcome Multinomial NB.
-It may due to the small size of the dataset, 20NEWS having only 19K documents.
-CNN usually perform less than other classification techniques for small datasets.
-As a consequence, we will apply our technique to the larger text dataset RCV1 which has 800K documents, as [13].
-
-We now address the specific comments of each referee.
-
-Referee 1.
-GNN: Thanks for pointing out the GNN model (we will add this ref).
-On the high level, GNN is a spatial/graph technique, unlike our model which is spectral.
-Technically, GNN was originally designed to classify nodes or graphs by first embedding each node in an Euclidean space with RNN, and then using those as features for classification or regression.
-Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN.
-Regarding the referee’s comment, we read very carefully the two 2009 GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee.
-We would be grateful if the referee could point out more connections between our model and the GNN papers.
-Rotational invariance: The spectral filters are indeed rotationally invariant.
-Although it first appears to be a limitation of this system, we think the rational invariance property might actually be an advantage for large-scale image datasets.
-Our motivation is inspired by the following experiment (not reported in the paper).
-We learned the classic and graph CNN from a rotated MNIST dataset, that is each training MNIST image is randomly rotated from -90° to 90° (no data augmentation here, simply rotation of original images).
-Classic CNN obtained 70.65% of accuracy, and graph CNN got 91.47%.
-Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features.
-We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN.
-We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.
-
-Referee 7.
-Single graph: We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a.
-brain connectivity analysis, and social network analysis (one graph and many data signals on this graph).
-Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (for example learn and transfer Chebyshev coefficients across graphs).
-We reserve this extension for another paper.
-Graph pooling: Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution.
-In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e.
-grouping similar vertices together, which is NP-hard.
-We opted here for the Metis/Graclus model [8,15], which has proved to be a very efficient graph clustering technique.
-Note also that the coarsening step is performed only once, before training.
-
-Referee 8.
-Presentation: We have recently improved the presentation and accessibility of the paper. We will be happy to release this new version in the final version of the paper.
-Theory: The analysis of the performances of standard CNN is currently a central and open mathematical question.
-It is thus challenging to state anything about the theoretical performances of our model vs [13].
-The only performance comparison is experimental and are summarised in Table 6.
-Besides, a theoretical advantage of our model is that our filters are exactly localized in K hops on graphs.
-
-Referee 9.
-Presentation: See our answer to R8.
-Natural graphs: Indeed, we did not present any natural graph experiments in this work.
-Our first goal was to introduce our efficient graph CNN technique.
-However, a data analysis project on Wikipedia network using this technique is ongoing.
+The only concerns come from Referees 8,9 who acknowledge their low confidence in their understanding of the paper main contributions.

### 2016-10-29 nips2016/review/rebuttal.txt
rebuttal v4 (2016-07-22)

@@
 
-4950
+4992
 
@@
 Referees 1,3,7
-Accuracy Gap for MNIST: The gap between standard CNN and graph CNN may be due to our isotopic filters.
-We also observed that with 20 epochs (instead of 10 in our submission) for LeNet5 architecture, standard CNN has 99.33% accuracy and graph CNN has 99.14%.
-This reveals a slower convergence of graph CNN, and the need to investigate better optimization strategies.
+While the isotropic nature of our filters (see answer to R1) and the lack of experience on architecture design are certainly explanations for the gap between classical CNNs and graph CNNs on MNIST, we observed that with 20 epochs (instead of 10 in our submission) for the same LeNet5 architectures, classical CNNs obtained 99.33% accuracy and graph CNNs 99.14%.
+This reveals a slower convergence of graph CNNs, and the need to investigate better suited optimization or initialization strategies.
 
 Referees 3,7,8
-Graph Usefulness and Experiments: The goal of Tables 3,5 is to show that the usefulness of the graph structure directly depends on the graph quality.
-When random graphs are used, then convolutional layers are useless, and graph CNN are not well suited for the problem.
-If the graph is well constructed, like the grid for MNIST, then graph CNN provide good performances.
-For 20NEWS, we did not try to optimize the graph structure.
-There may be better ways to construct higher-quality graphs, but it was not our objective.
-The goal of Table 4 is to demonstrate that convolutional layers CN32 are also effective on graphs that are not regular image grids.
-This is showed by comparing CN32 to fully connected (FC) networks.
-We also compared against the effective Multinomial NB document classification technique, and SVM and linear classifier baseline techniques.
-Graph CNN did not overcome Multinomial NB.
-It may be due to the small size of the dataset (20NEWS has 19K documents).
-CNN usually perform less than other classification techniques for small datasets.
-We plan to apply our technique to the larger text dataset RCV1 (800K documents).
+The goal of Tables 3 and 5 is to show that the usefulness of the graph structure directly depends on the graph quality.
+Random graphs are used as a baseline to assess graph quality, as convolutional layers are almost useless in this setting.
+A well suited graph, like the grid for MNIST, is useful and allows our model to perform well.
+We did however not try to optimize the graph structure for 20NEWS: there may be better ways to construct higher-quality graphs.
+It was not our objective, which was to show how graph quality impacts performance.
+The goal of Table 4 is to demonstrate that a convolutional layer is also effective on graphs that are not regular grids, by comparing it to fully connected networks.
+We also compared against the effective Multinomial Naive Bayes and two baselines.
+Graph CNNs did not overcome Multinomial NB, likely because of the small size of the 20NEWS dataset (~19k documents).
+We plan to apply our technique to the larger RCV1 dataset (800k documents).
 
@@
 Thanks for pointing out GNN (we will add this ref).
-On the high level, GNN is a spatial technique, and our model is spectral.
-GNN was designed to classify nodes/graphs by first embedding each node in an Euclidean space with RNN, and then using those as features for classification.
-Our technique is developed to classify graph signals by successively extracting higher level features, as done for classical CNN.
-Regarding the referee’s comment, we read carefully the two GNN papers, and were not able to express the graph CNN in terms of GNN, as suggested by the referee.
+At a high level, GNN was designed to classify nodes or graphs by first embedding each node in an Euclidean space with an RNN, then using those as features for classification/regression.
+Our technique was developed to classify graph signals by successively extracting higher level features, similarly to classical CNNs.
+After reading the paper carefully, we were not able to express our graph CNN in terms of a GNN, as suggested by the referee.
 We would be grateful if the referee could point out more connections between our model and the GNN papers.
 The spectral filters are indeed rotationally invariant.
-Although it first appears to be a limitation, we think the rotational invariance property might actually be an advantage.
-Our motivation is inspired by the following experiment (not reported in the paper).
-We learned the classic and graph CNN from a rotated MNIST dataset, i.e.
-each training image is randomly rotated (no data augmentation here, simply rotation of original images).
-Classic CNN obtained 70% of accuracy, and graph CNN 91%.
-Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications where the training data are not enough to capture all meaningful features.
-We will investigate such idea in a future work, that is if such invariance property is an advantage for large-scale image datasets, where same objects may rotate, and thus requiring standard CNN to learn more filters than graph CNN.
-We will also study the possibility to learn anisotropic Laplacians, and thus anisotropic filters.
+Whether this is a limitation or an advantage depends on the problem and should be verified empirically, as for any other invariance.
+Our claim is motivated by the following experiment (not reported in the paper): on a rotated MNIST dataset, i.e.
+each image is randomly rotated (no data augmentation here), our model obtained an accuracy of 91% while classical CNNs obtained 70%.
+Although such experiment is naive, it reveals that spatial invariance properties like rotation can be an asset in some applications.
+Moreover, rotational invariance has been sought in the past.
+Many data augmentation schemes have used rotated versions of (part of) images to force a network to learn this invariance.
+Spatial Transformer Networks have been developed to learn invariance to translation, scale, rotation and more generic warping.
+In a future work, we will test our model on large-scale image datasets and investigate if such an invariance is an advantage.
+We will also study the possibility to learn anisotropic filters with the help of anisotropic Laplacians.
 
 Referee 7
-We indeed consider a single graph in this work, which is already meaningful for multiple applications s.a.
-brain connectivity analysis, and social network analysis.
-Transfer graph features/filters from one graph to another graph is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (e.g.
+We indeed consider a single graph in this work, which is already meaningful for multiple applications such as brain connectivity analysis, documents network analysis and social network analysis.
+Transferring graph features/filters from one graph to another is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (e.g.
 learn and transfer Chebyshev coefficients across graphs).
 We reserve this extension for another paper.
-Pooling is a pillar operation in any CNN architecture as it allows to reduce spatial resolution and increase feature resolution.
-In the case of grids, it is mathematically well-defined, but in case of graphs it is equivalent to graph clustering or coarsening, i.e.
-grouping similar vertices together, which is NP-hard.
-We opted here for the Metis model, which has proved to be very efficient.
-Note also that the coarsening step is performed only once.
+A pooling strategy will always require some sort of coarsening, i.e.
+a way to group nodes to form a receptive field.
+While grouping is straightforward on grids, clustering or coarsening on arbitrary graphs is NP-hard.
+We opted here for the greedy Graclus algorithm, which has proved to be very efficient.
+Note that our method is however agnostic to the chosen coarsening algorithm and that coarsening is performed only once.
 
 Referee 8
-We have recently improved the presentation of the paper, and will release the new version in the final version of the paper.
-The performance analysis of standard CNN is a central and open mathematical question.
+We did recently improve the presentation of the paper. Expect the final version .
+The performance analysis of classical CNNs is a central and open mathematical question.
 It is thus challenging to state anything about the theoretical performances of our model.
-The only performance study is experimental and summarised in Table 6.
+The only performance study is experimental and summarized in Table 6, which shows that our model is indeed superior to [13].
 
 Referee 9
-Presentation: See our answer to R8.
-We indeed did not introduce any natural graph experiments in this work.
-Our first goal was to introduce an efficient graph CNN technique.
-However, a data analysis project on Wikipedia network using this technique is ongoing.
+Presentation: see our answer to R8.
+We did not introduce any natural graph experiments in this work.
+Our first goal was to introduce an efficient extension of CNNs to graphs and demonstrate its usefulness.
+Constructed graphs are a great way of demonstrating this.
+However, a data analysis project on the network of Wikipedia articles using this technique is ongoing.
 
@@
 The only concerns come from Referees 8,9 who acknowledge their low confidence in their understanding of the paper main contributions.
+
+Please note that, to foster innovation, we are going to publish the code used to produce our results once the paper is published.

### 2016-10-29 nips2016/nips2016.tex
large edition for nips camera-ready, first submission (2016-10-27)

@@
-\pdfoutput=1
-
 \documentclass{article}
@@
 %\usepackage{nips_2016}
-%\usepackage[final]{nips_2016}
-\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry} % arxiv
+\usepackage[final]{nips2016}
+%\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry} % arxiv
 
@@
 \usepackage{amsfonts}       % blackboard math symbols
-%\usepackage{amssymb}
+\usepackage{amssymb}
+\usepackage{amsmath}
 \usepackage{xfrac}       % compact symbols for 1/2, etc.
@@
 \usepackage{color}
-
-\usepackage{amsmath}
+\usepackage{capt-of}
 
 % Use less space
-%\usepackage{titlesec}		% space before / after headings
-%\titlespacing\section{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
-%\titlespacing\subsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
+\usepackage{titlesec}		% space before / after headings
+\titlespacing\section{0pt}{6pt plus 1pt minus 2pt}{1pt plus 1pt minus 2pt}
+\titlespacing\subsection{0pt}{6pt plus 1pt minus 2pt}{1pt plus 1pt minus 2pt}
 %\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
-%\usepackage{parskip} \setlength{\parskip}{5pt}
-%\setlength{\bibsep}{0.2pt} %comment for arxiv
+\setlength{\bibsep}{1.5pt} %comment for arxiv
 
-\setlength{\textfloatsep}{20.0pt plus 2.0pt minus 2.0pt}
-\setlength{\floatsep}{20.0pt plus 2.0pt minus 2.0pt}
-\setlength{\intextsep}{20.0pt plus 2.0pt minus 2.0pt}
+\usepackage{parskip} \setlength{\parskip}{0pt} \setlength{\parindent}{1em}
+
+\setlength{\textfloatsep}{7.2pt plus 1.0pt minus 2.0pt}
+\setlength{\floatsep}{7.2pt plus 1.0pt minus 2.0pt}
+\setlength{\intextsep}{7.2pt plus 1.0pt minus 2.0pt}
+
+\setcounter{topnumber}{3}
+%\setcounter{bottomnumber}{8}
+%\setcounter{totalnumber}{8}
 
@@
 
+\author{
+  Michaël Defferrard \\ %\thanks{further info} \\
+  %Department of Electrical Engineering \\
+  %École Polytechnique Fédérale de Lausanne (EPFL) \\
+  EPFL, Lausanne, Switzerland \\
+  \texttt{michael.defferrard@epfl.ch} \\
+  \And % \AND
+  Xavier Bresson \\
+  %School of Computer Science and Engineering \\
+  Nanyang Technological University, Singapore \\
+  \texttt{xavier.bresson@ntu.edu.sg} \\
+  \And % \AND
+  Pierre Vandergheynst \\
+  %Department of Electrical Engineering \\
+  EPFL, Lausanne, Switzerland \\
+  \texttt{pierre.vandergheynst@epfl.ch} \\
+}
+
+%%Arxiv
 %\author{
-%  Michaël Defferrard \\ %\thanks{further info} \\
-%  %Michael Defferrard \\ %\thanks{further info} \\
-%  %Department of Electrical Engineering \\
-%  %École Polytechnique Fédérale de Lausanne (EPFL) \\
+%  Micha\"{e}l Defferrard\hspace{1cm}Xavier Bresson\hspace{1cm} Pierre Vandergheynst\\ \\
+%  Department of Electrical Engineering \\
 %  EPFL, Lausanne, Switzerland \\
-%  \texttt{michael.defferrard@epfl.ch} \\
-%  \And % \AND
-%  Xavier Bresson \\
-%  EPFL, Lausanne, Switzerland \\
-%  \texttt{xavier.bresson@epfl.ch} \\
-%  \And % \AND
-%  Pierre Vandergheynst \\
-%  EPFL, Lausanne, Switzerland \\
-%  \texttt{pierre.vandergheynst@epfl.ch} \\
+%  {\small \texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} } \\\\
 %}
 
-%%Arxiv
-\author{
-  Micha\"{e}l Defferrard\hspace{1cm}Xavier Bresson\hspace{1cm} Pierre Vandergheynst\\ \\
-  Department of Electrical Engineering \\
-  EPFL, Lausanne, Switzerland \\
-  {\small \texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} } \\\\
-}
-
-\usepackage[acronym]{glossaries}
-\newacronym{CNN}{CNN}{Convolutional Neural Network}
-\newacronym{SVD}{SVD}{Singular Value Decomposition}
-\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
-\newacronym{PSD}{PSD}{positive semidefinite}
+%\usepackage[acronym]{glossaries}
+%\newacronym{CNN}{CNN}{Convolutional Neural Network}
+%\newacronym{SVD}{SVD}{Singular Value Decomposition}
+%\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
+%\newacronym{PSD}{PSD}{positive semidefinite}
 
@@
 \newcommand{\eqnref}[1]{(\ref{eq:#1})}
-\newcommand{\bruna}{art:BrunaZarembaSzlamLeCun13DLgraphs,
-art:HenaffBrunaLeCun15DLgraphs}
 
@@
 
-Convolutional neural networks (CNNs) have greatly improved state-of-the-art performances in a number of fields, notably computer vision and natural language processing.
-In this work, we are interested in generalizing the formulation of CNNs from low-dimensional regular Euclidean domains, where images (2D), videos (3D) and audios (1D) are represented, to high-dimensional irregular domains such as social networks or biological networks represented by graphs.
-This paper introduces a formulation of CNNs on graphs in the context of spectral graph theory.
-We borrow the fundamental tools from the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design localized graph filters efficient to learn and evaluate.
-As a matter of fact, we introduce the first technique that offers the same computational complexity than standard CNNs, while being universal to any graph structure.
-Numerical experiments on MNIST and 20NEWS demonstrate the ability of this novel deep learning system to learn local, stationary, and compositional features on graphs, as long as the graph is well-constructed.
+In this work, we are interested in generalizing convolutional neural networks
+(CNNs) from low-dimensional regular grids, where image, video and speech are
+represented, to high-dimensional irregular domains, such as social networks,
+brain connectomes or words’ embedding, represented by graphs.
+We present a formulation of CNNs in the context of spectral graph theory, which
+provides the necessary mathematical background and efficient numerical schemes
+to design fast localized convolutional filters on graphs. Importantly, the
+proposed technique offers the same linear computational complexity and constant
+learning complexity as classical CNNs, while being universal to any graph
+structure. Experiments on MNIST and 20NEWS demonstrate the ability of this
+novel deep learning system to learn local, stationary, and compositional
+features on graphs.
 
@@
 	%except the rotation invariance
-
-	%Graphs can either be the natural domain of the data, such as social networks,
-	%biological graphs like gene regulatory and brain connectivity networks,
-	%telecommunication networks, or constructed from the data, such as similarity
-	%graphs.
 
@@
 
-%Previous works have been developed along this today essential line of research, but they lack of mathematical foundations and efficient learning of filters.
+%Previous works have been developed along this today essential line of
+%research, but they lack of mathematical foundations and efficient learning of
+%filters.
 
-\begin{figure}[h!]
+\section{Introduction}
+
+%Convolutional neural networks (CNNs) have greatly improved state-of-the-art performances in a number of fields, notably computer vision.
+
+	%Graphs can either be the natural domain of the data, such as social networks,
+	%biological graphs like gene regulatory and brain connectivity networks,
+	%telecommunication networks, or constructed from the data, such as similarity
+	%graphs.
+% Constructed graphs make sense because of comp complexity, as evidenced by LRF.
+
+Convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST} offer
+an efficient architecture to extract highly meaningful statistical patterns in
+large-scale and high-dimensional datasets. The ability of CNNs to learn local
+stationary structures and compose them to form multi-scale hierarchical
+patterns has led to breakthroughs in image, video, and sound recognition tasks
+\cite{art:LeCunBengioHinton15DL}. Precisely, CNNs extract the local
+stationarity property of the input data or signals by revealing local features
+that are shared across the data domain. These similar features are identified
+with localized convolutional filters or kernels, which are learned from the
+data. Convolutional filters are shift- or translation-invariant filters,
+meaning they are able to recognize identical features independently of their
+spatial locations. Localized kernels or compactly supported filters refer to
+filters that extract local features independently of the input data size, with
+a support size that can be much smaller than the input size.
+
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{images/illustrationCNNgraphs}
-\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
-coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
+\includegraphics[width=0.8\textwidth]{images/illustrationCNNgraphs}
+\caption{Architecture of the proposed CNN on graphs.
+	%Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at
+	%layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral
+	%filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the
+	%coarsening exponent, $n_c$ is the number of classes, $y$ is the output
+	%signal, and $\theta^l$ is the number of parameters to learn at $l$.
+}
 \label{fig_illustration}
@@
 
-\section{Introduction}
+% Eventually, generic and efficient implementations of CNNs on graphs, as
+% proposed here and related works
+% \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,masci2015geodesic,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs},
+% is essential to pave the way to the development of a new class of deep learning
+% techniques for graph-based data.
 
-Convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST}
-offer an efficient architecture to extract highly meaningful statistical
-patterns in large-scale and high-dimensional datasets. The key success of CNNs
-is its advanced ability to learn local stationary structures and compose them to
-form multi-scale hierarchical patterns. Precisely, CNNs extract the local
-stationarity property of the input data or signals by revealing local features
-that are shared across the data domain. These similar features are identified
-with localized convolutional filters or kernels, which are learned from the
-data. Convolutional filters are shift- or translation-invariant filters, meaning
-they are able to recognize identical features independently of their spatial
-locations. Localized kernels or compactly supported filters refer to filters
-that extract local features independently of the input data size, with a support
-size that can be much smaller than the input size.
-
-%\newpage
-The compositional local stationarity property of CNNs can be efficiently
-implemented on regular grids, like image and sound domains, as convolutional, 
-downsampling and pooling operators are mathematically well-defined on such discretized
-Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
-sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on
-low-dimensional regular lattices. But not all data lie on regular grids. User
-data on social networks, gene data on biological regulatory networks, log data
-on telecommunication networks, or text documents on words' embedding are
-important examples of data lying on irregular or non-Euclidean domains.
-Unfortunately, CNNs cannot be directly applied to such complex domains.% as the building blocks of CNNs, convolution, downsampling, and pooling, are only defined for regular grids.
-
-%A generalization to irregular grids is not straightforward as the building blocks of CNNs, convolution, downsampling, and pooling, are only defined for regular grids.
-
-A generalization to irregular grids is not straightforward as the standard operators of convolution, downsampling, and pooling, are only defined for regular grids.
-This makes this extension challenging, both theoretically and implementation-wise.
-This work leverages the recent works of
-\cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
-art:BrunaZarembaSzlamLeCun13DLgraphs, art:HenaffBrunaLeCun15DLgraphs,
-art:HammondVandergheynstGribonval11GraphWav,
-art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to introduce an
-efficient implementation of CNNs on irregular domains represented here as
-graphs. Graphs are universal representations of heterogeneous pairwise
-relationships between possibly high-dimensional data. Graphs can encode complex
-geometric structures, and can be studied with strong mathematical tools such as
-spectral graph theory \cite{book:Chung97Spectral}, a blend between graph theory and harmonic analysis.
+%The compositional local stationarity property of CNNs can be efficiently
+%implemented on regular grids, like image and sound domains, as convolutional, 
+%downsampling and pooling operators are mathematically well-defined on such discretized
+%Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
+%sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on
+%low-dimensional regular lattices. But not all data lie on regular grids. User
+%data on social networks, gene data on biological regulatory networks, log data
+%on telecommunication networks, or text documents on words' embedding are
+%important examples of data lying on irregular or non-Euclidean domains.
+%Unfortunately, CNNs cannot be directly applied to such complex domains.
 
@@
 
-The major bottleneck of generalizing CNNs to graphs, and one of the primary goals of
-this work, is the definition of graph localized filters which are efficient to
+%Unfortunately, CNNs cannot be directly applied to such complex domains.
+
+User data on social networks, gene data on biological regulatory networks, log
+data on telecommunication networks, or text documents on word embeddings are
+important examples of data lying on irregular or non-Euclidean domains which
+can be structured with graphs, which are universal representations of
+heterogeneous pairwise relationships. Graphs can encode complex geometric
+structures and can be studied with strong mathematical tools such as spectral
+graph theory \cite{book:Chung97Spectral}.
+
+A generalization of CNNs to graphs is not straightforward as the convolution
+and pooling operators are only defined for regular grids. This makes this
+extension challenging, both theoretically and implementation-wise. The major
+bottleneck of generalizing CNNs to graphs, and one of the primary goals of this
+work, is the definition of localized graph filters which are efficient to
 evaluate and learn.
-%that are the core elements to extract local stationarity structures.
-We will make use of recent tools developed in the context of graph signal
-processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to achieve such goals. Precisely, the main contributions of this work are summarized below:
+Precisely, the main contributions of this work are summarized below.
 \begin{enumerate}
 %\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
-\item \textbf{Spectral formulation.} A graph spectral theoretical formulation of CNNs on graphs built on
-	established tools in GSP.
-\item \textbf{Strictly localized filters.} Enhancing \cite{\bruna}, the proposed spectral
-	filters are provable to be strictly localized in a ball of radius $K$, i.e. $K$ hops from
-	the central vertex.  
-\item \textbf{Low computational complexity.} The evaluation complexity of our filters is
-	linear w.r.t.
-the filters support's size $K$ and the number of edges $|\E|$.
-Importantly, as most real-world graphs are highly sparse, we have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor (NN) graphs, leading to a linear complexity w.r.t the input data size.
-Moreover,
-	this method avoids the Fourier basis altogether, thus the expensive
-	eigenvalue decomposition (EVD) necessary to compute it as well as the need
-	to store the basis, a matrix of size $n^2$. That is especially relevant when
-	working with limited GPU memory. Besides the data, our method only requires
-	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
+\item \textbf{Spectral formulation.} A spectral graph theoretical formulation
+	of CNNs on graphs built on established tools in GSP
+	\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
+\item \textbf{Strictly localized filters.} Enhancing
+	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, the proposed spectral filters
+	are provable to be strictly localized in a ball of radius $K$, i.e. $K$
+	hops from the central vertex.  
+\item \textbf{Low computational complexity.} The evaluation complexity of our
+	filters is linear w.r.t. the filters support's size $K$ and the number of
+	edges $|\E|$. Importantly, as most real-world graphs are highly sparse, we
+	have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor
+	(NN) graphs, leading to a linear complexity w.r.t the input data size.
+	Moreover, this method avoids the Fourier basis altogether, thus the
+	expensive eigenvalue decomposition (EVD) necessary to compute it as well as
+	the need to store the basis, a matrix of size $n^2$. That is especially
+	relevant when working with limited GPU memory. Besides the data, our method
+	only requires to store the Laplacian, a sparse matrix of $|\E|$ non-zero
+	values.
 \item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
-	graphs which, after a rearrangement of the vertices as a binary tree structure, is analog to pooling of
-	1D signals.
-\item \textbf{Experimental results.} We present many experiments that ultimately
-	show that our formulation is (1) a useful model, (2) computationally
-	efficient and (3) superior both in accuracy and complexity to the pioneer spectral graph CNNs introduced in \cite{\bruna}. We also show that
-	our graph formulation performs similarly as a classical CNNs on MNIST and study the
-	impact of various graph constructions on classification performance.
+	graphs which, after a rearrangement of the vertices as a binary tree
+	structure, is analog to pooling of 1D signals.
+\item \textbf{Experimental results.} We present multiple experiments that
+	ultimately show that our formulation is (i) a useful model, (ii)
+	computationally efficient and (iii) superior both in accuracy and
+	complexity to the pioneer spectral graph CNN introduced in
+	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. We also show that our graph
+	formulation performs similarly as a classical CNNs on MNIST and study the
+	impact of various graph constructions on classification performance. The
+	TensorFlow \cite{abadi_tensorflow_2016} code to reproduce our results and
+	apply the model to other data is available as an open-source
+	software.\footnote{\url{https://github.com/mdeff/cnn_graph}}
 \end{enumerate}
 
-\noindent
-Eventually, generic and efficient implementations of CNNs on
-graphs, as proposed here and related works \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
-art:MasciBoscainiBronsteinVandergheynst15ShapeNet,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, is essential to pave the way to the development of a new class of deep learning techniques for graph-based data.
+\section{Proposed Technique}
 
-\section{Related Works}
-\subsection{Graph Signal Processing} GSP is an emerging field that aims at
-bridging the gap between signal processing techniques like wavelet analysis \cite{book:Mallat99wavelets} and
-graph theory such as spectral graph analysis
-\cite{art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}. A goal is to
-generalize fundamental analysis operations for signals from regular grids to
-irregular structures embodied by graphs. We refer the reader to
-\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
- for an
-introduction of the field. Standard operations on grids such as convolution,
-translation, filtering, dilatation, modulation or downsampling do not extend
-directly to graphs and thus require new mathematical definitions while keeping
-the original intuitive concepts. In this context, the authors of
-\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
-pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
-operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,
-art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform
-mutli-scale pyramid transforms on graphs. The works of
-\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
-art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles
-on graphs, and showed that intuitive concepts may be lost, but can also produce
-enhanced localization principles for signals on graphs. In
-\cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to
-carry out lasso-based signal regularization on graphs, and studied the
-intertwined relationships between smoothness and sparsity on graphs. In
-\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
-investigated compressed sensing recovery conditions for graph spectral signal
-analysis.
-
-In the future, we expect to
-further benefit from newly developed tools in GSP to enhance our mathematical formulation of
-CNNs on graphs. This paper introduces the mathematical and computational foundations of
-such models, while future developments may consider to enhance the building
-blocks, namely graph filtering and coarsening.
-
-\subsection{CNNs on Non-Euclidean Domains}
-
-%Extending the success of CNNs to
-%non-Euclidean domains is essential to boost the power analysis of data lying on
-%complex networks like biological, social or telecommunication networks.
-
-Extending CNNs to non-Euclidean domains has been a recent line of work. A preliminary work was proposed with 
- the so-called local reception fields \cite{pro:GregorLeCun10LRF,
-pro:CoatesNg11LRF}, successfully applied to image recognition. The main idea is
-to group together features based upon a measure of similarity such as to select
-a limited number of connections between two successive layers. While this model
-reduced the number of parameters by exploiting the locality assumption, it did
-not attempt to exploit any stationarity property of data, i.e. no weight-sharing strategy. %stationarity This approach is however not able to extract similar groups over
-%the data domain, as it is not clear how to define convolutional filters on
-%these groups.
-%\todo{read papers}
-
-A first attempt to design CNNs on graphs was introduced in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} by adopting a spatial approach.
-The idea was to construct a multiresolution graph by exploiting graph neighbourhood structures, and learned a deep neural network.
-However, such approach did not consider convolution operations, and thus no stationarity property.
-%By construction, spatial approaches provide filter localization.
-However, although graph convolution directly in the spatial domain is conceivable, it also faces the challenge of matching local neighborhoods, as shown by \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.
-
-ShapeNet in \cite{pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
-art:MasciBoscainiBronsteinVandergheynst15ShapeNet} is a generalization of CNNs
-to 3D-meshes, a class of smooth low-dimensional non-Euclidean spaces. The
-authors used geodesic polar coordinates to define convolution operations on mesh patches, and formulated a
-deep learning architecture which allows comparison across different manifolds.
-They obtained state-of-the-art results for 3D shape recognition tasks.
-
-In \cite{pro:ChenChengMallat14deepHaar} and \cite{pro:RustamovGuibas14deepHaar},
-the authors investigated the construction of Haar wavelet transforms on graphs
-using a deep hierarchical architecture. They applied the method to object
-recognition on sphere, and to sparse reconstruction of faces.
-%\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
-%they impose a shape to the filters.}
-%\todo{Unknown graph geometry?}
-
-Finally, \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} and
-\cite{art:HenaffBrunaLeCun15DLgraphs} introduced a generalization of CNNs on
-graphs using graph counterparts of grid convolution and downsampling. They were
-able to learn convolutional filters on graphs using the convolution theorem
-\cite{book:Mallat99wavelets} that defines convolutions as linear operators that
-diagonalize in the Fourier basis (represented by the eigenvectors of the
-Laplacian operator). They were thus the first to introduce a spectral
-formulation to extend CNNs on graphs. They also introduced a strategy to learn
-the graph structure from the data, and applied their model on image recognition,
-text categorization and bioinformatics. This approach does however not scale up due
-to the necessary multiplications by the graph Fourier basis, a matrix of $n^2$
-coefficients, where $n$ is the data dimensionality. Despite the cost of
-computing this matrix, which requires an EVD on the graph Laplacian, the
-dominant cost is the need to multiply the data by this matrix twice (forward and
-inverse Fourier transforms) at a cost of $\bO(n^2)$ operations per forward and
-backward pass, a computational bottleneck already identified by the authors.
-Besides, as they rely on smoothness (via a spline parametrization) in the
-Fourier domain to bring localization in the vertex domain, their model does not
-provide a precise control over the local support of their kernels, which is
-essential to learn localized filters. Our technique leverages on this work, and
-we will show how to overcome these limitations and beyond.
-
-\section{Proposed Technique}
-Generalizing CNNs to graphs requires three fundamental steps.
-First step is the design of localized convolutional filters on graphs.
-Second step is a graph coarsening procedure that groups together similar features on graphs.
-And the last step is the graph pooling operation that trades spatial resolution for higher filter resolution.
+Generalizing CNNs to graphs requires three fundamental steps: (i) the design of
+localized convolutional filters on graphs, (ii) a graph coarsening procedure
+that groups together similar vertices and (iii) a graph pooling operation that
+trades spatial resolution for higher filter resolution.
 
 \subsection{Learning Fast Localized Spectral Filters} \label{sec:filters}
-The goal of this section is to define fast convolutional localized filters on graphs.
-There are two strategies to define convolutional localized filters; either from a spatial approach or from a spectral approach.
-By construction, spatial approaches provide filter localization via the finite size of the kernel.
-However, although graph convolution directly in the spatial domain is conceivable, it also faces the challenge of matching local neighbourhoods, as pointed out in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}.
-Consequently, there is no unique mathematical definition of spatial translations on graphs from a spatial perspective.
-On the other side, a spectral approach provides a well-defined translation operator on graphs via convolutions with a Kronecker delta function implemented in the spectral domain
-\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
-However, a filter defined in the spectral domain is not naturally localized and translations are costly with $\bO(n^2)$ operations due to multiplications with the graph Fourier basis.
-Both limitations can be overcome with a special choice of filter parametrization, as explained below.
+
+There are two strategies to define convolutional filters; either from a spatial
+approach or from a spectral approach. By construction, spatial approaches
+provide filter localization via the finite size of the kernel. However,
+although graph convolution directly in the spatial domain is conceivable, it
+also faces the challenge of matching local neighborhoods, as pointed out in
+\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Consequently, there is no unique
+mathematical definition of translation on graphs from a spatial perspective. On
+the other side, a spectral approach provides a well-defined translation
+operator on graphs via convolutions with a Kronecker delta implemented in the
+spectral domain \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
+The convolution theorem \cite{book:Mallat99wavelets} defines convolutions as
+linear operators that diagonalize in the Fourier basis (represented by the
+eigenvectors of the Laplacian operator). However, a filter defined in the
+spectral domain is not naturally localized and translations are costly due to
+the $\bO(n^2)$ multiplication with the graph Fourier basis. Both limitations
+can however be overcome with a special choice of filter parametrization.
 
@@
 \R^{n \times n}$ where $D \in \R^{n \times n}$ is the diagonal degree matrix
-with $D_{ii} = \sum_j W_{ij}$, and its normalized definition is $L = I_n -
+with $D_{ii} = \sum_j W_{ij}$, and normalized definition is $L = I_n -
 D^{-1/2} W D^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real
@@
 \paragraph{Polynomial parametrization for localized filters.} There are however
-two limitations with non-parametric filters: (1) they are not localized in space and (2)
-their learning complexity is in $\bO(n)$, the dimensionality of the data. These issues can be overcome 
-with the use of Laplacian-based polynomial spectral filters: 
+two limitations with non-parametric filters: (i) they are not localized in
+space and (ii) their learning complexity is in $\bO(n)$, the dimensionality of
+the data. These issues can be overcome with the use of a polynomial filter
+% a Laplacian-based polynomial spectral filter
 \begin{equation} \label{eq:filt_poly}
@@
 where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
-The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is given
-by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
+The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is
+given by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
 (L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
@@
 5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
-$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum number of edges connecting two vertices on the graph.
-Consequently, spectral filters represented by $K^\text{th}$-order polynomials of
-the Laplacian are exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the support size of the filter, and thus the same complexity as in standard CNNs.
+$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum
+number of edges connecting two vertices on the graph.  Consequently, spectral
+filters represented by $K^\text{th}$-order polynomials of the Laplacian are
+exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the
+support size of the filter, and thus the same complexity as classical CNNs.
 
 % While there is many possible choice of recursive polynomials
-\paragraph{Recursive formulation for fast filtering.} While we have shown how to
-learn localized filters with $K$ parameters, the cost to filter a signal $x$ as
-$y = Ug_\theta(\Lambda)U^Tx$ is still high with $\bO(n^2)$ operations because of the
-multiplications with the Fourier basis $U$. A solution of this problem is to parametrize
-$g_\theta(L)$ as a polynomial function that can be computed recursively from
-$L$, as $K$ multiplications by a sparse $L$ costs $\bO(K|\E|) \ll \bO(n^2)$. One
-such polynomial, traditionally used in GSP to approximate kernels (like
-wavelets), is the Chebyshev expansion
+\paragraph{Recursive formulation for fast filtering.} While we have shown how
+to learn localized filters with $K$ parameters, the cost to filter a signal $x$
+as $y = Ug_\theta(\Lambda)U^Tx$ is still high with $\bO(n^2)$ operations
+because of the multiplication with the Fourier basis $U$. A solution to this
+problem is to parametrize $g_\theta(L)$ as a polynomial function that can be
+computed recursively from $L$, as $K$ multiplications by a sparse $L$ costs
+$\bO(K|\E|) \ll \bO(n^2)$. One such polynomial, traditionally used in GSP to
+approximate kernels (like wavelets), is the Chebyshev expansion
 \cite{art:HammondVandergheynstGribonval11GraphWav}. Another option, the Lanczos
@@
 \end{equation}
-where the $x_{s,i}$ are the input feature maps and the $F_{in} \times F_{out}$ vectors of Chebyshev
-coefficients $\theta_{i,j} \in \R^K$ are the layer's trainable parameters. When
-training multiple convolutional layers with the backpropagation algorithm, one
-needs the two gradients:
+where the $x_{s,i}$ are the input feature maps and the $F_{in} \times F_{out}$
+vectors of Chebyshev coefficients $\theta_{i,j} \in \R^K$ are the layer's
+trainable parameters. When training multiple convolutional layers with the
+backpropagation algorithm, one needs the two gradients
 \begin{align}
@@
 F_{out} S)$ operations. These can be efficiently computed on parallel
-architectures by leveraging tensor operations. Eventually, $[\bar{x}_{s,i,0}, \ldots, \bar{x}_{s,i,K-1}]$ only need to be computed once.
+architectures by leveraging tensor operations. Eventually, $[\bar{x}_{s,i,0},
+\ldots, \bar{x}_{s,i,K-1}]$ only needs to be computed once.
 
 \subsection{Graph Coarsening} \label{sec:coarsening}
-Local stationarity property of data is extracted via localized convolutional kernels.
-We are now interested to extract the multi-scale hierarchical composition property of data.
-In standard CNNs, this is efficiently achieved via grid subsampling and pooling, which trades spatial resolution with feature resolution reducing the learning complexity without compromising the system performances.
-In contrast with regular domains, the subsampling operation on graphs or graph coarsening is not mathematically sound.
-It requires to construct meaningful neighborhoods on
-graphs where similar vertices are clustered together. Doing this for multiple
-layers is equivalent to construct a multi-scale clustering of the graph that
-preserves local geometric structures. It is however well-known that graph
-clustering is a NP-hard problem \cite{art:BuiJonesGraphPartNPhard} and that
-approximations must be used. While there exist many clustering techniques (e.g.
-the popular spectral clustering \cite{art:VonLuxburg07Tutorial}), we are most
-interested in multilevel clustering algorithms, where each level produces a
-{coarser} graph which corresponds to the data domain seen at a different
-{resolution}. Moreover, clustering techniques that reduce the size of the
-graph by a factor two at each level offers a precise control on the coarsening
-and pooling size. In this work, we make use of the coarsening phase of the
-Graclus multilevel clustering algorithm \cite{art:DhillonGuanKulis07Graclus},
-which has been shown to be extremely efficient at clustering a large variety of
-graphs. We briefly review it below. Algebraic multigrid techniques on graphs
-\cite{art:RonSafroBrandt11MultigridGraph} and the Kron reduction
-\cite{art:ShumanFarajiVandergheynst16PyramTrans} are two methods worth exploring
-in future works. %The later has the main advantage to preserve the ordering of
-%the Laplacian spectrum, and may be able to commute with the graph convolution
-%operator. \todo{If proven, this would be a very essential mathematical property
-%of CNNs on graphs, exactly as downsampling for classical CNNs.}
 
-Graclus \cite{art:DhillonGuanKulis07Graclus}, built on Metis \cite{art:KarypisKumar98Metis}, uses a greedy
-algorithm to compute successive coarser versions of a given graph and is able to
-minimize several popular spectral clustering objectives. We chose the normalized
-cut \cite{art:ShiMalik00NCut}, which is an excellent clustering energy.
-Graclus' greedy rule consists, at any given coarsening level, in
-picking an unmarked vertex $i$ and matching it with one of its unmarked
-neighbors $j$ that maximizes the local normalized cut $W_{ij} (1/d_i + 1/d_j)$.
-The two matched vertices are then marked and the coarsened weights are set as
-the sum of their weights. The matching is repeated until all nodes have been
-explored. This is an extremely fast coarsening scheme which divides the number
-of nodes by approximately two (there may exist a few singletons, non-matched nodes)
-from one level to the next coarser level.
+% Local stationarity property of data is extracted via localized convolutional
+% kernels. We are now interested to extract the multi-scale hierarchical
+% composition property of data. In standard CNNs, this is efficiently achieved
+% via grid subsampling and pooling, which trades spatial resolution with
+% feature resolution reducing the learning complexity without compromising the
+% system performances. In contrast with regular domains, the subsampling
+% operation on graphs or graph coarsening is not mathematically sound.
+
+The pooling operation requires meaningful neighborhoods on graphs, where
+similar vertices are clustered together. Doing this for multiple layers is
+equivalent to a multi-scale clustering of the graph that preserves local
+geometric structures. It is however known that graph clustering is
+\mbox{NP-hard} \cite{art:BuiJonesGraphPartNPhard} and that approximations must
+be used. While there exist many clustering techniques, e.g. the popular
+spectral clustering \cite{art:VonLuxburg07Tutorial}, we are most interested in
+multilevel clustering algorithms where each level produces a coarser graph
+which corresponds to the data domain seen at a different resolution.  Moreover,
+clustering techniques that reduce the size of the graph by a factor two at each
+level offers a precise control on the coarsening and pooling size.  In this
+work, we make use of the coarsening phase of the Graclus multilevel clustering
+algorithm \cite{art:DhillonGuanKulis07Graclus}, which has been shown to be
+extremely efficient at clustering a large variety of graphs. Algebraic
+multigrid techniques on graphs \cite{art:RonSafroBrandt11MultigridGraph} and
+the Kron reduction \cite{art:ShumanFarajiVandergheynst16PyramTrans} are two
+methods worth exploring in future works.
+%The later has the main advantage to preserve the ordering of the Laplacian
+%spectrum, and may be able to commute with the graph convolution operator.
+%\todo{If proven, this would be a very essential mathematical property of CNNs
+%on graphs, exactly as downsampling for classical CNNs.}
+
+Graclus \cite{art:DhillonGuanKulis07Graclus}, built on Metis
+\cite{art:KarypisKumar98Metis}, uses a greedy algorithm to compute successive
+coarser versions of a given graph and is able to minimize several popular
+spectral clustering objectives. We chose the normalized cut
+\cite{art:ShiMalik00NCut}, which is an excellent clustering energy. Graclus'
+greedy rule consists, at any given coarsening level, in picking an unmarked
+vertex $i$ and matching it with one of its unmarked neighbors $j$ that
+maximizes the local normalized cut $W_{ij} (1/d_i + 1/d_j)$.  The two matched
+vertices are then marked and the coarsened weights are set as the sum of their
+weights. The matching is repeated until all nodes have been explored. This is
+an extremely fast coarsening scheme which divides the number of nodes by
+approximately two (there may exist a few singletons, non-matched nodes) from
+one level to the next coarser level.
 
 \subsection{Fast Pooling of Graph Signals}  \label{sec:pooling}
-As standard CNNs, pooling operations reduce the spatial resolution allowing higher filter resolution.
-Pooling operations, such as max pooling or average pooling, are carried out many times during the optimization, and thus must be as efficient as possible.
-After the graph coarsening operation in the previous section, the graph and its coarsened versions have an unstructured arrangement of the vertices.
-Hence, if the pooling is directly applied then it would need a table to store all matched vertices, which would result in memory consuming, slow, and not (easily) parallelizable pooling operations.
 
-It is however possible to structure the arrangement of the vertices, and achieve a graph pooling operation as efficient as a 1D grid pooling.
-We proceed as follows.
-After coarsening, each node has either two
-parents, if it was matched at the finer level, or one, if it was not, i.e the
-node was a singleton. From the coarsest to finest level, fake nodes, i.e.
-disconnected nodes, are added to pair with the singletons such that each node
-has two children. This structure is a balanced binary tree: (1)
-regular nodes (and singletons) either have two regular nodes (e.g. level 1
-vertex 0 in \figref{pooling}) or (2) one singleton and a fake node as children
-(e.g. level 2 vertex 0 in \figref{pooling}), and (3) fake nodes always have two
-fake nodes as children (e.g. level 1 vertex 1 in \figref{pooling}). Input
-signals are initialized with a neutral value at the fake nodes, e.g. 0 when
-using a ReLU activation with max pooling. Because these nodes are disconnected,
-filtering does not impact the initial neutral value. While those fake nodes do
-artificially increase the dimensionality thus the computational cost, we found
-that, in practice, the number of singletons left by Graclus is quite low
-% < 10%?
-compared to the number of vertices.
-
-Arbitrarily ordering the nodes at the coarsest level, then propagating this
-ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
-children, produces a regular ordering in the finest level. Regular in the sense
-that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
-rearranged graph signal is analog to pooling a regular 1D signal.
-\figref{pooling} shows an example of the whole coarsening and pooling process.
-This regular arrangement makes the pooling operation very efficient and
-satisfies parallel architectures such as GPUs as memory accesses are local, i.e.
-matched nodes do not have to be fetched.
-
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
@@
 
-\section{Numerical Experiments}
-All experiments were performed with TensorFlow, an open-source library for
-numerical computation using data flow graphs, especially suited for deep
-learning \cite{abadi_tensorflow_2016}. It features various backends, notably
-CUDA to compute on Nvidia GPUs. All computations are carried on an Nvidia Tesla
-K40c GPU.
+Pooling operations are carried out many times and must be efficient. After
+coarsening, the vertices of the input graph and its coarsened versions are not
+arranged in any meaningful way. Hence, a direct application of the pooling
+operation would need a table to store all matched vertices. That would result
+in a memory inefficient, slow, and hardly parallelizable implementation. It is
+however possible to structure the arrangement of the vertices such that a graph
+pooling operation becomes as efficient as a 1D pooling. We proceed in two
+steps: (i) create a balanced binary tree and (ii) rearrange the vertices.
+After coarsening, each node has either two parents, if it was matched at the
+finer level, or one, if it was not, i.e the node was a singleton. From the
+coarsest to finest level, fake nodes, i.e.  disconnected nodes, are added to
+pair with the singletons such that each node has two children. This structure
+is a balanced binary tree: (i) regular nodes (and singletons) either have two
+regular nodes (e.g. level 1 vertex 0 in \figref{pooling}) or (ii) one singleton
+and a fake node as children (e.g. level 2 vertex 0 in \figref{pooling}), and
+(iii) fake nodes always have two fake nodes as children (e.g. level 1 vertex 1
+in \figref{pooling}). Input signals are initialized with a neutral value at the
+fake nodes, e.g. 0 when using a ReLU activation with max pooling. Because these
+nodes are disconnected, filtering does not impact the initial neutral value.
+While those fake nodes do artificially increase the dimensionality thus the
+computational cost, we found that, in practice, the number of singletons left
+by Graclus is quite low.
+% < 10%? compared to the number of vertices.
+Arbitrarily ordering the nodes at the coarsest level, then propagating this
+ordering to the finest levels, i.e. node $k$ has nodes $2k$ and $2k+1$ as
+children, produces a regular ordering in the finest level. Regular in the sense
+that adjacent nodes are hierarchically merged at coarser levels. Pooling such a
+rearranged graph signal is analog to pooling a regular 1D signal.
+\figref{pooling} shows an example of the whole coarsening and pooling process.
+This regular arrangement makes the pooling operation very efficient and
+satisfies parallel architectures such as GPUs as memory accesses are local,
+i.e.  matched nodes do not have to be fetched.
 
-In the sequel, we refer to filters defined by \eqnref{filt_non-param} as
-\textit{Non-Param}, i.e. the spatially non-localized filters, and the proposed filters defined by \eqnref{filt_cheby} as
-\textit{Chebyshev}. Filters referred to as \textit{Spline} are defined by
+\section{Related Works}
+
+\subsection{Graph Signal Processing}
+
+The emerging field of GSP aims at bridging the gap between signal processing
+and spectral graph theory \cite{book:Chung97Spectral,
+art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}, a blend between
+graph theory and harmonic analysis. A goal is to generalize fundamental
+analysis operations for signals from regular grids to irregular structures
+embodied by graphs. We refer the reader to
+\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} for an
+introduction of the field.  Standard operations on grids such as convolution,
+translation, filtering, dilatation, modulation or downsampling do not extend
+directly to graphs and thus require new mathematical definitions while keeping
+the original intuitive concepts. In this context, the authors of
+\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
+pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
+operators on graphs and techniques to perform mutli-scale pyramid transforms on
+graphs were proposed in \cite{art:ShumanFarajiVandergheynst16PyramTrans,
+art:RamEladCohen11TreeWavelets}. The works of
+\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
+art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty
+principles on graphs and showed that while intuitive concepts may be lost,
+enhanced localization principles can be derived.
+
+%In \cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown
+%how to carry out lasso-based signal regularization on graphs, and studied the
+%intertwined relationships between smoothness and sparsity on graphs. In
+%\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
+%investigated compressed sensing recovery conditions for graph spectral signal
+%analysis.
+
+\subsection{CNNs on Non-Euclidean Domains}
+
+%Extending the success of CNNs to
+%non-Euclidean domains is essential to boost the power analysis of data lying on
+%complex networks like biological, social or telecommunication networks.
+
+The Graph Neural Network framework \cite{scarselli_gnn_2009}, simplified in
+\cite{li_ggsnn_2015}, was designed to embed each node in an Euclidean space
+with a RNN and use those embeddings as features for classification or
+regression of nodes or graphs. By setting their \textit{transition function}
+$f$ as a simple diffusion instead of a neural net with a recursive relation,
+their \textit{state} vector becomes $s = f(x) = Wx$.  Their point-wise
+\textit{output function} $g_\theta$ can further be set as $\hat{x} =
+g_\theta(s, x) = \theta (s - Dx) + x = \theta Lx + x$ instead of another neural
+net. The Chebyshev polynomials of degree $K$ can then be obtained with a
+$K$-layer GNN, to be followed by a non-linear layer and a graph pooling
+operation. Our model can thus be interpreted as multiple layers of diffusions
+and node-local operations.
+
+The works of \cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF} introduced the
+concept of constructing a local receptive field to reduce the number of learned
+parameters. The idea is to group together features based upon a measure of
+similarity such as to select a limited number of connections between two
+successive layers. While this model reduces the number of parameters by
+exploiting the locality assumption, it did not attempt to exploit any
+stationarity property, i.e. no weight-sharing strategy.
+%This approach is however not able to extract similar groups over the data
+%domain, as it is not clear how to define convolutional filters on these
+%groups.
+The authors of \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} used this idea for
+their spatial formulation of graph CNNs. They use the weighted graph to define
+the local neighborhood and compute a multiscale clustering of the graph for the
+pooling operation. Inducing weight sharing in a spatial construction is however
+challenging, as it requires to select and order the neighborhoods when a
+problem-specific ordering (spatial, temporal, or otherwise) is missing.
+
+A spatial generalization of CNNs to 3D-meshes, a class of smooth
+low-dimensional non-Euclidean spaces, was proposed in \cite{masci2015geodesic}.
+The authors used geodesic polar coordinates to define convolution operations on
+mesh patches, and formulated a deep learning architecture which allows
+comparison across different manifolds. They obtained state-of-the-art results
+for 3D shape recognition tasks.
+
+% For learning similarity graph
+%In \cite{pro:ChenChengMallat14deepHaar} and \cite{pro:RustamovGuibas14deepHaar},
+%the authors investigated the construction of Haar wavelet transforms on graphs
+%using a deep hierarchical architecture. They applied the method to object
+%recognition on sphere, and to sparse reconstruction of faces.
+%\todo{Learn parameters of a wavelet kernel? Then it is a parametric method,
+%they impose a shape to the filters.}
+%\todo{Unknown graph geometry?}
+
+The first spectral formulation of graph CNN, proposed in
+\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, defines a filter as
 \begin{equation} \label{eq:filt_spline}
@@
 where $B \in \R^{n \times K}$ is the cubic B-spline basis and the parameter
-$\theta \in \R^K$ is a vector of control points, as proposed in \cite{\bruna}. We use for all techniques the advanced Graclus
-coarsening algorithm introduced in \secref{coarsening} rather than the simple
-agglomerative method of \cite{\bruna}. The motivation is to compare
-the quality of learned filters, not the coarsening algorithms.
+$\theta \in \R^K$ is a vector of control points.
+They later proposed a strategy to learn the graph structure from the data and
+applied the model to image recognition, text categorization and bioinformatics
+\cite{art:HenaffBrunaLeCun15DLgraphs}.
+This approach does however not scale up due to the necessary multiplications by
+the graph Fourier basis $U$. Despite the cost of computing this matrix, which
+requires an EVD on the graph Laplacian, the dominant cost is the need to
+multiply the data by this matrix twice (forward and inverse Fourier transforms)
+at a cost of $\bO(n^2)$ operations per forward and backward pass, a
+computational bottleneck already identified by the authors.  Besides, as they
+rely on smoothness, via the spline parametrization, in the Fourier domain to
+bring localization in the vertex domain, their model does not provide a precise
+control over the local support of their kernels, which is essential to learn
+localized filters. Our technique leverages on this work, and we showed how to
+overcome these limitations and beyond.
+
+\section{Numerical Experiments}
+
+%All experiments were performed with TensorFlow, an open-source library for
+%numerical computation using data flow graphs, especially suited for deep
+%learning \cite{abadi_tensorflow_2016}. It features various backends, notably
+%CUDA to compute on Nvidia GPUs.
+%All computations are carried on an Nvidia Tesla K40c GPU.
+
+In the sequel, we refer to the non-parametric and non-localized filters
+\eqnref{filt_non-param} as \textit{Non-Param}, the filters \eqnref{filt_spline}
+proposed in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} as \textit{Spline} and the proposed filters
+\eqnref{filt_cheby} as \textit{Chebyshev}. We always use the Graclus coarsening
+algorithm introduced in \secref{coarsening} rather than the simple
+agglomerative method of \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Our motivation is to compare the learned
+filters, not the coarsening algorithms.
 
@@
 denotes a fully connected layer with $k$ hidden units, P$k$ denotes a (graph or
-classical) pooling layer of size and stride $k$, C$k$ denotes a convolutional
-layer with $k$ feature maps and GC$k$ denotes a graph convolutional layer with
-$F_{out} = k$ feature maps (with Chebyshev filters if not specified otherwise).
+classical) pooling layer of size and stride $k$, GC$k$ and C$k$ denote a
+(graph) convolutional layer with $k$ feature maps.
 All FC$k$, C$k$ and GC$k$ layers are followed by a ReLU activation $\max(x,0)$.
 The final layer is always a softmax regression and the loss energy $E$ is the
-cross-entropy loss with an $\ell_2$ regularization on the weights of all
-FC$k$ layers. Mini-batches are of size $S = 100$.  %\todo{dropout}
+cross-entropy with an $\ell_2$ regularization on the weights of all FC$k$
+layers. Mini-batches are of size $S = 100$.
 
-%\todo{$k$-NN similarity graphs are built as follows: }
-
-\subsection{Revisiting Standard CNNs on MNIST} \label{sec:MNIST}
+\subsection{Revisiting Classical CNNs on MNIST} \label{sec:MNIST}
 %\subsection{Comparison with classical CNNs on MNIST}
@@
 
-To validate our model, we first apply it to the Euclidean case with the benchmark MNIST classification
-problem \cite{pro:LeCunBottouBengioHaffner98MNIST}.
-In this situation, the graph is simply a $k$-NN graph of the Euclidean 2D grid.
-This is an important sanity check for our model, which must be able to extract
-features on any graph, including the regular 2D grid on which images reside.
-
-\begin{table*}[h!]
+\begin{table*}[t]
 \centering
@@
 \end{tabular}
-\caption{Classification accuracies for the proposed model and a classical CNN on
-MNIST.} 
+\caption{Classification accuracies of the proposed graph CNN and a classical
+CNN on MNIST.} 
 \label{tab:mnist}
@@
 
-We recall that MNIST is a dataset of 70,000 digit numbers represented on a 2D grid of size $28
-\times 28$, such that data points lie on a space of $784$ dimensions. For
-our graph model, we construct a $8$-NN graph of the 2D Euclidean
-grid, which produces a graph of $n = |\V| = 976$ nodes (784 pixels and 192 fake
-nodes as explained in \secref{pooling}) and $|\E| = 3198$ edges.
+To validate our model, we applied it to the Euclidean case on the benchmark
+MNIST classification problem \cite{pro:LeCunBottouBengioHaffner98MNIST}, a
+dataset of 70,000 digits represented on a 2D grid of size $28 \times 28$. For
+our graph model, we construct a $8$-NN graph of the 2D grid which produces a
+graph of $n = |\V| = 976$ nodes ($28^2 = 784$ pixels and 192 fake nodes as
+explained in \secref{pooling}) and $|\E| = 3198$ edges. Following standard
+practice, the weights of a $k$-NN similarity graph (between features) are
+computed as
+%$W_{ij} = \exp(-\|z_i - z_j\|_2^2 / \sigma^2)$
+\begin{equation} \label{eq:knngraph}
+	W_{ij} = \exp \left( -\frac{\|z_i - z_j\|_2^2}{\sigma^2} \right)
+\end{equation}
+where $z_i$ is the 2D coordinate of pixel $i$.
+% and \sigma is set using the scheme proposed in
 
-\tabref{mnist} confirms the ability of our model to learn localized filters
-as it achieves a performance very close to the classical CNNs with the same LeNet-5-like
-architecture. The LeNet-5-like network architecture and the following hyper-parameters are
+This is an important sanity check for our model, which must be able to extract
+features on any graph, including the regular 2D grid. \tabref{mnist} shows the
+ability of our model to achieve a performance very close to a classical CNN
+with the same architecture. The gap in performance may be explained by the
+isotropic nature of the spectral filters, i.e. the fact that edges in a general
+graph do not possess an orientation (like up, down, right and left for pixels
+on a 2D grid). Whether this is a limitation or an advantage depends on the
+problem and should be verified, as for any invariance. Moreover, rotational
+invariance has been sought: (i) many data augmentation schemes have used
+rotated versions of images and (ii) models have been developed to learn this
+invariance, like the Spatial Transformer Networks \cite{jaderberg2015spatial}.
+Other explanations are the lack of experience on architecture design and the
+need to investigate better suited optimization or initialization strategies.
+
+The LeNet-5-like network architecture and the following hyper-parameters are
 borrowed from the TensorFlow MNIST tutorial\footnote{
@@
 probability of 0.5, regularization weight of $5\times10^{-4}$, initial learning
-rate of 0.03, learning rate decay of 0.95, momentum of 0.9. Convolutional layer
-have filters of size $5 \times 5$ while graph convolution layers have the same
+rate of 0.03, learning rate decay of 0.95, momentum of 0.9. Convolutional layers
+have filters of size $5 \times 5$ and graph convolutional layers have the same
 support of $K = 5^2 = 25$. All models were trained for 20 epochs.
 
-\subsection{Text Categorization with 20NEWS}
-%usefulness of our model and its 
-To demonstrate the versatility of our model to work with graphs generated from unstructured data, we applied our technique to the text categorization problem with the 20NEWS dataset.
+\subsection{Text Categorization on 20NEWS}
 
-\begin{table*}[h!] \centering
-\begin{tabular}{lc} \toprule
+\begin{table*}[t]
+\begin{minipage}[b]{0.42\linewidth}
+	\centering
+	\begin{tabular}[b]{lc}
+		\toprule
 		Model & Accuracy \\
@@
 		%GC32-FC500 & \todo{?} \\
-\bottomrule \end{tabular}
-\caption{Classification accuracies for the proposed model and a other algorithms
-on 20NEWS.} 
+		\bottomrule
+	\end{tabular}
+	\caption{Accuracies of the proposed graph CNN and other
+	methods on 20NEWS.} 
 	\label{tab:20news}
+\end{minipage}
+\hfill
+\begin{minipage}[b]{0.55\linewidth}
+	\centering
+	\includegraphics[width=\textwidth]{images/runtime_crop}
+	\captionof{figure}{Time to process a mini-batch of $S=100$ 20NEWS documents
+	w.r.t. the number of words $n$.}
+	\label{fig:runtime}
+\end{minipage}
 \end{table*}
 
-20NEWS consists of 18,846 (11,314 for training and 7,532 for testing) text
-documents associated with 20 classes \cite{art:Joachims9620NEWS}. We extracted
-the 10,000 most common words from the 93,953 unique words in this corpus. Each
-document is represented using the bag-of-words model, normalized across words.
-To test our model, we constructed a 16-NN graph of the word2vec
-\cite{pro:MikolovChenCorradoDean13word2vec} embedding of those words, which
-produced a graph of $n = |\V| = 10,000$ nodes and
-$|\E| = 132,834$ edges.  All CNN models were trained for 20 epochs by the Adam
-optimizer \cite{art:KingmaBa14AdamOpt} with a learning rate of 0.1, without regularization nor
-dropout. The filter support was set to $K = 5$.
+To demonstrate the versatility of our model to work with graphs generated from
+unstructured data, we applied our technique to the text categorization problem
+on the 20NEWS dataset which consists of 18,846 (11,314 for training and 7,532
+for testing) text documents associated with 20 classes
+\cite{art:Joachims9620NEWS}. We extracted the 10,000 most common words from the
+93,953 unique words in this corpus. Each document $x$ is represented using the
+bag-of-words model, normalized across words. To test our model, we constructed
+a 16-NN graph with \eqnref{knngraph} where $z_i$ is the word2vec embedding
+\cite{pro:MikolovChenCorradoDean13word2vec} of word $i$, which produced a graph
+of $n = |\V| = 10,000$ nodes and $|\E| = 132,834$ edges. All models were
+trained for 20 epochs by the Adam optimizer \cite{art:KingmaBa14AdamOpt} with
+an initial learning rate of 0.001. The architecture is GC32 with support $K =
+5$. \tabref{20news} shows decent performances: while the proposed model does
+not outperform the multinomial naive Bayes classifier on this small dataset, it
+does defeat fully connected networks, which require much more parameters.
 
-\tabref{20news} shows that CNNs on graphs provide decent performances.
-While it does not outperform the kernel-based method of multinomial naive Bayes classifier on this small dataset, it does defeat fully connected networks, which require much more parameters to be learned.
-More importantly, these results verify the validity of the statistical assumptions made for the data, that are locality and stationarity, and which are at the core of the design of any CNN technique.
-While we know that these data properties are true for low-dimensional Euclidean data like audios, images and videos, we also show experimentally that they are also satisfied for text documents as long as the graph is properly constructed.
-\secref{graph_quality} will study the influence of the graph quality.
+%todo hyper-parameters
 
-\subsection{Numerical Comparison between Spectral Filters}
+\subsection{Comparison between Spectral Filters and Computational Efficiency}
 
-In this section, we compare the proposed spectral filters to the non-parametric filters, and those proposed in \cite{\bruna} on the MNIST and 20NEWS datasets.
-\tabref{filters} reports that the proposed kernel parametrization outperforms
-\cite{\bruna} as well as the non-parametric filters, which are not localized and require $\bO(n)$ parameters to learn.
-
-\begin{table*}[h!] \centering
+\begin{table*}[t!] \centering
 \begin{tabular}{llccc} \toprule
@@
 Dataset & Architecture & Non-Param \eqnref{filt_non-param} &
-Spline \eqnref{filt_spline} \cite{\bruna} &
+Spline \eqnref{filt_spline} \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} &
 Chebyshev \eqnref{filt_cheby} \\
@@
 \bottomrule \end{tabular}
-\caption{Comparison of accuracy results for different types of spectral filters.
-Spline and Chebyshev filters use $K=25$.} % \todo{$K=5$ for 20NEWS}
+\caption{Classification accuracies for different types of spectral filters
+($K=25$).} % \todo{on MNIST and $K=5$ on 20NEWS.}
 \label{tab:filters}
@@
 
-\subsection{Computational Efficiency}
-
-\figref{runtime} validates the low computational complexity of the proposed CNN technique on graphs.
-The training time of our model scales as $\bO(n)$, while \cite{\bruna} scales as $\bO(n^2)$.
-Moreover,
-\figref{convergence} gives a sense of how the validation accuracy as well as the
-loss energy converges w.r.t. the three filter definitions. Finally,
-\tabref{speedup} compares training time on CPU and GPU. The fact that we observe
-the same order of speedup as classical CNNs exemplifies the natural
-parallelization opportunity offered by our model. That is possible because our
-method relies solely on matrix multiplications which are efficiently implemented
-by cuBLAS, the linear algebra routines provided by NVIDIA.
-
-\begin{table*}[ht] \centering
+\begin{table*}[t]
+\centering
+%\vspace{-20pt}
 \begin{tabular}{llccc} \toprule
 & & \multicolumn{2}{c}{Time (ms)} & \\
-Model & Architecture & CPU & GPU & Speedup \\
+Model & Architecture & CPU & GPU & Speedup \\ % {\small(Tesla K40c)} 
 \midrule
 Classical CNN & C32-P4-C64-P4-FC512 & 210 & 31 & 6.77x \\
-Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 200 & 25 & 8.00x \\
+Proposed graph CNN & GC32-P4-GC64-P4-FC512 & 1600 & 200 & 8.00x \\
 \bottomrule \end{tabular}
-\caption{Time to process a mini-batch of 100 MNIST images. That is the total
-training time divided by the number of gradient steps.} 
+\caption{Time to process a mini-batch of $S=100$ MNIST images.} 
 \label{tab:speedup}
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
-\vspace{-20pt}
-\includegraphics[width=\textwidth]{images/runtime_crop}
-\caption{Training time over a mini-batch of 100 20NEWS documents w.r.t. $n$, the
-number of selected words, for the three considered filters. The architecture is
-GC32 with $K=5$ for Spline and Chebyshev.}
-\label{fig:runtime}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\subfigure[]{\includegraphics[width=0.49\textwidth]{images/acc.png}}
-\hfill
-\subfigure[]{\includegraphics[width=0.49\textwidth]{images/loss.png}}
-\caption{Plots of accuracy (a) and energy loss (b) for the first 2000
-iterations.}
+\includegraphics[width=0.8\textwidth]{images/training}
+\caption{Plots of validation accuracy and training loss for the first 2000
+iterations on MNIST.}
 \label{fig:convergence}
@@
 
+\tabref{filters} reports that the proposed parametrization \eqnref{filt_cheby}
+outperforms \eqnref{filt_spline} from \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} as well as non-parametric
+filters \eqnref{filt_non-param} which are not localized and require $\bO(n)$
+parameters. Moreover, \figref{convergence} gives a sense of how the validation
+accuracy and the loss $E$ converges w.r.t. the filter definitions.
+
+\figref{runtime} validates the low computational complexity of our model which
+scales as $\bO(n)$ while \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} scales as $\bO(n^2)$. The measured
+runtime is the total training time divided by the number of gradient steps.
+\tabref{speedup} shows a similar speedup as classical CNNs when moving to GPUs.
+This exemplifies the parallelization opportunity offered by our model, who
+relies solely on matrix multiplications. Those are efficiently implemented by
+cuBLAS, the linear algebra routines provided by NVIDIA.
+
 \subsection{Influence of Graph Quality} \label{sec:graph_quality}
 
-For the proposed method to be successful, the statistical assumptions of  locality, stationarity, and compositionality regarding the data must be fulfilled on the graph where the data resides.
-Therefore, the classification performance, the quality of learned filters, all critically depends on the quality of the graph.
-For data lying on Euclidean space, experiments in
- \secref{MNIST} show that a simple $k$-NN graph of the grid is good enough to recover almost exactly the performance of standard CNNs.
-We also noticed that the value of $k$ does not have a strong influence on the results.
-We can witness the importance of a graph satisfying the data assumptions by comparing its performance with a random
-graph.
-\tabref{mnist_quality} reports a large drop of accuracy when using a random graph, that is when the data structure is lost and the convolutional layers are not useful anymore to extract meaningful features.
-
-%the accuracy drops (\todo{to the order of an architecture without
-%convolution}), as the convolutional layers are not useful anymore to extract
-%meaningful features. The structure has been lost.
-
-\begin{table*}[h!]
+\begin{table*}[t]
 \centering
@@
 
-While images naturally lie on Euclidean 2D spaces, represented by graphs of regular grids, this is not the case for
-text documents. For our method to be applied to this class of data, a feature graph does have to be built.
-We investigate here 5 $16$-NN similarity graphs of words. The simplest option is
-to represent each word as its corresponding column in the bag-of-words matrix.
-Another approach is to learn embeddings on the corpus with word2vec
-\cite{pro:MikolovChenCorradoDean13word2vec} or to use pre-trained embeddings on
-Google News (given by the authors). As the dataset gets larger (in number of
-samples and dimensionality), it is often not an option to compute the distance
-between all features, such that an approximate nearest neighbors algorithm shall
-be used. We used the LSHForest \cite{pro:BawaCondieGanesan05LSHForest} on the
-learned word2vec embedding. \tabref{20news_quality} reports the classification
-results. It shows that the graph built on the learned word2vec embedding is the
-best at capturing the local and stationarity properties of text documents. It is
-worth noticing that the approximate $k$-NN graph constructed from it is almost
-as bad as the random graph, meaning that it may be a better strategy to learn a
-good low-dimensional embedding first and then construct an exact $k$-NN graph
-from this embedding.
-
-\begin{table*}[h!] \centering
+\begin{table*}[t] \centering
 \begin{tabular}{ccccc} \toprule
 & \multicolumn{2}{c}{word2vec} & & \\
-bag-of-words & pre-trained & learned & approximate & random \\
+bag-of-words & pre-learned & learned & approximate & random \\
 \midrule
@@
 
+% While we know that these data properties are true for low-dimensional
+% Euclidean data like audios, images and videos, we also show experimentally
+% that they are also satisfied for text documents as long as the graph is
+% properly constructed.
+
+% More importantly, these results verify the validity of the statistical
+% assumptions made on the data, that are locality and stationarity, and which
+% are at the core of the design of any CNN technique.
+
+For any graph CNN to be successful, the statistical assumptions of locality,
+stationarity, and compositionality regarding the data must be fulfilled on the
+graph where the data resides. Therefore, the learned filters' quality and thus
+the classification performance critically depends on the quality of the graph.
+For data lying on Euclidean space, experiments in \secref{MNIST} show that a
+simple $k$-NN graph of the grid is good enough to recover almost exactly the
+performance of standard CNNs. We also noticed that the value of $k$ does not
+have a strong influence on the results. We can witness the importance of a
+graph satisfying the data assumptions by comparing its performance with a
+random graph. \tabref{mnist_quality} reports a large drop of accuracy when
+using a random graph, that is when the data structure is lost and the
+convolutional layers are not useful anymore to extract meaningful features.
+
+While images can be structured by a grid graph, a feature graph has to be built
+for text documents represented as bag-of-words.  We investigate here three ways
+to represent a word $z$ when constructing a $16$-NN similarity graph with
+\eqnref{knngraph}: while the simplest option is to represent each word as its
+corresponding column in the bag-of-words matrix, another approach is to learn
+an embedding for each word with word2vec
+\cite{pro:MikolovChenCorradoDean13word2vec} or to use the pre-learned
+embeddings provided by the authors. For larger datasets, an approximate nearest
+neighbors algorithm may be required, which is the reason we tried LSHForest
+\cite{pro:BawaCondieGanesan05LSHForest} on the learned word2vec embeddings.
+\tabref{20news_quality} reports the classification results which highlight the
+importance of a well constructed graph.
+
+%the graph built on the learned word2vec embedding is the
+%best at capturing the local and stationarity properties of text documents. It is
+%worth noticing that the approximate $k$-NN graph constructed from it is almost
+%as bad as a random graph, meaning that it may be a better strategy to learn a
+%good low-dimensional embedding first and then construct an exact $k$-NN graph
+%from this embedding.
+
 \section{Conclusion and Future Work}
 
-We have introduced an efficient implementation of CNNs on non-Euclidean
-manifolds represented by graphs using tools from GSP. Experiments have shown the
-ability of the model to extract local and stationary features through the graph
-convolutional layers.
+In this paper, we have introduced the mathematical and computational
+foundations of an efficient generalization of CNNs on graphs using tools from
+GSP. Experiments have shown the ability of the model to extract local and
+stationary features through graph convolutional layers. Compared with the first
+work on spectral graph CNNs introduced in \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, our model provides a
+strict control over the local support of filters, is computationally more
+efficient by avoiding an explicit use of the Graph Fourier basis, and
+experimentally shows a better test accuracy. Besides, we addressed the three
+concerns raised by \cite{art:HenaffBrunaLeCun15DLgraphs}: (i) we introduced a
+model whose computational complexity is linear with the dimensionality of the
+data, (ii) we confirmed that the quality of the input graph is of paramount
+importance, (iii) we showed that the statistical assumptions of local
+stationarity and compositionality made by the model are verified for text
+documents as long as the graph is well constructed.
 
-Compared with the first works of spectral graph CNNs
-introduced in \cite{\bruna}, our model provides a strict control over the local
-support of filters, is computationally more efficient by
-avoiding an explicit use of the Graph Fourier basis, and experimentally shows a
-better test accuracy. Besides, we addressed the three concerns raised by
-\cite{art:HenaffBrunaLeCun15DLgraphs}. First, we introduced a model whose computational complexity is linear with the
-dimensionality of the data.
-Second, we confirmed that the quality of the input graph is of paramount importance.
-Along this line, the idea of \cite{art:HenaffBrunaLeCun15DLgraphs} to learn the input graph from data is highly meaningful.
-And a natural and future approach would be to alternate the learning of the CNN parameters and the graph, as a virtuous circle.
-Third, we showed that the statistical assumptions of local stationarity and compositionality made by the model are verified for text documents as long as the graph is well constructed.
+Future works will investigate two directions. On one hand, we will enhance the
+proposed framework with newly developed tools in GSP. On the other hand, we
+will explore applications of this generic model to important fields where the
+data naturally lies on graphs, which may then incorporate external information
+about the structure of the data rather than artificially created graphs which
+quality may vary as seen in the experiments. Another natural and future
+approach, pioneered in \cite{art:HenaffBrunaLeCun15DLgraphs}, would be to
+alternate the learning of the CNN parameters and the graph, as a virtuous
+circle.
 
-Future works will investigate two directions. On one hand, we will explore
-applications of this generic model to important fields where the data naturally
-lie on non-artificially constructed graphs such as social networks, biological
-networks, or telecommunication networks. It makes actually more sense to apply
-this generalized CNN framework to natural network-based data, rather than
-artificially created networks which quality may vary as seen in the experiments.
-On the other hand, we will also improve this framework with tools developed in GSP,
-including graph wavelet operators
-\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
-	pro:GavishNadlerCoifman10GraphHaar, pro:ChenChengMallat14deepHaar,
-	pro:RustamovGuibas14deepHaar}, spectral graph coarsening techniques
-	\cite{art:ShumanFarajiVandergheynst16PyramTrans}, and improved localization
-	properties \cite{pro:TsitsveroBarbarossa15Uncert,
-	pro:PasdeloupAlamiGriponRabbat15Uncert,
-art:PerraudinRicaudShumanVandergheynst16Uncert}.
+% including graph wavelet operators
+% \cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
+% 	%pro:ChenChengMallat14deepHaar, pro:RustamovGuibas14deepHaar,
+% 	pro:GavishNadlerCoifman10GraphHaar}, spectral graph coarsening techniques
+% 	\cite{art:ShumanFarajiVandergheynst16PyramTrans}, and improved localization
+% 	properties \cite{pro:TsitsveroBarbarossa15Uncert,
+% 	pro:PasdeloupAlamiGriponRabbat15Uncert,
+% art:PerraudinRicaudShumanVandergheynst16Uncert}.
+
+%In a future work, we will test our model on large-scale image datasets and
+%investigate if such an invariance is an advantage. We will also study the
+%possibility to learn anisotropic filters with anisotropic Laplacians.
+
+%In the future, we expect to further benefit from newly developed tools in GSP
+%to enhance our mathematical formulation of CNNs on graphs. This paper
+%introduces the mathematical and computational foundations of such models, while
+%future developments may consider to enhance the building blocks, namely graph
+%filtering and coarsening.
 
@@
 
-\newpage
+\clearpage
 \bibliographystyle{plain}
@@
 %}
+
+{
+	\setstretch{0}
+	\small
 	\bibliography{refs}
+}
 

### 2016-10-29 nips2016/images/runtime.py

@@ -1,19 +1,29 @@
 #!/bin/env python
 
 import numpy as np
-import matplotlib
 import matplotlib.pyplot as plt
-matplotlib.rcParams.update({'font.size': 18})
 
-fourier = np.array([10,23,68,139.5,231.5,351,495])/400*1000
-cheby = np.array([7.5,11.5,20,29.5,38,48.5,59])/400*1000
-x = np.array([1,2,4,6,8,10,12]) * 1000
+plt.rc('pdf', fonttype=42)
+plt.rc('ps', fonttype=42)
 
-plt.figure(figsize=(15,5))
-plt.plot(x, cheby, '.-', markersize=15, label='Chebyshev');
-plt.plot(x, fourier, '.--', markersize=15, label='Non-Param / Spline');
+SIZE = 15
+plt.rc('font', size=SIZE)         # controls default text sizes
+plt.rc('axes', titlesize=SIZE)    # fontsize of the axes title
+plt.rc('axes', labelsize=SIZE)    # fontsize of the x any y labels
+plt.rc('xtick', labelsize=SIZE)   # fontsize of the tick labels
+plt.rc('ytick', labelsize=SIZE)   # fontsize of the tick labels
+plt.rc('legend', fontsize=SIZE)   # legend fontsize
+plt.rc('figure', titlesize=SIZE)  # size of the figure title
+
+fourier = np.array([10, 23, 68, 139.5, 231.5, 351, 495]) / 400 * 1000
+cheby = np.array([7.5, 11.5, 20, 29.5, 38, 48.5, 59]) / 400 * 1000
+x = np.array([1, 2, 4, 6, 8, 10, 12]) * 1000
+
+plt.figure(figsize=(7, 3.2))
+plt.plot(x, cheby, '.-', markersize=15, label='Chebyshev')
+plt.plot(x, fourier, '.--', markersize=15, label='Non-Param / Spline')
 plt.legend(loc='upper left')
-plt.xlim(1000,12000)
+plt.xlim(1000, 12000)
 plt.xlabel('number of features (words)')
 plt.ylabel('time (ms)')
 plt.savefig('runtime.pdf')

### 2016-10-29 nips2016/nips2016.tex
tweaking for nips camera-ready, second submission (2016-10-28)

@@
 \usepackage{titlesec}		% space before / after headings
-\titlespacing\section{0pt}{6pt plus 1pt minus 2pt}{1pt plus 1pt minus 2pt}
-\titlespacing\subsection{0pt}{6pt plus 1pt minus 2pt}{1pt plus 1pt minus 2pt}
+\titlespacing\section{0pt}{5pt plus 1pt minus 2pt}{1.8pt plus 1pt minus 1pt}
+\titlespacing\subsection{0pt}{5pt plus 1pt minus 2pt}{2.7pt plus 1pt minus 2pt}
 %\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
-\setlength{\bibsep}{1.5pt} %comment for arxiv
 
-\usepackage{parskip} \setlength{\parskip}{0pt} \setlength{\parindent}{1em}
+\setlength{\bibsep}{1.52pt} %comment for arxiv
 
-\setlength{\textfloatsep}{7.2pt plus 1.0pt minus 2.0pt}
-\setlength{\floatsep}{7.2pt plus 1.0pt minus 2.0pt}
-\setlength{\intextsep}{7.2pt plus 1.0pt minus 2.0pt}
+\usepackage{parskip} \setlength{\parskip}{2.1pt} \setlength{\parindent}{1em}
+
+\setlength{\textfloatsep}{8pt plus 1.0pt minus 2.0pt}
+%\setlength{\floatsep}{8.2pt plus 1.0pt minus 2.0pt}
+%\setlength{\intextsep}{8.2pt plus 1.0pt minus 2.0pt}
 
@@
 
+\newcommand{\affe}{\textsuperscript{\dag}}
+\newcommand{\affs}{\textsuperscript{\ddag}}
+
 \author{
-  Michaël Defferrard \\ %\thanks{further info} \\
+	Michaël Defferrard\affe \hspace{1cm} Xavier Bresson\affs \hspace{1cm} Pierre Vandergheynst\affe %\thanks{further info} \\
 	%Department of Electrical Engineering \\
 	%École Polytechnique Fédérale de Lausanne (EPFL) \\
-  EPFL, Lausanne, Switzerland \\
-  \texttt{michael.defferrard@epfl.ch} \\
-  \And % \AND
-  Xavier Bresson \\
-  %School of Computer Science and Engineering \\
-  Nanyang Technological University, Singapore \\
-  \texttt{xavier.bresson@ntu.edu.sg} \\
-  \And % \AND
-  Pierre Vandergheynst \\
-  %Department of Electrical Engineering \\
-  EPFL, Lausanne, Switzerland \\
-  \texttt{pierre.vandergheynst@epfl.ch} \\
+	\vspace{0.2cm} \\
+	\affe EPFL, Lausanne, Switzerland \hspace{1cm} \affs Nanyang Technological University, Singapore \\
+	\texttt{\{michael.defferrard,pierre.vandergheynst\}@epfl.ch \hfill xavier.bresson@ntu.edu.sg} \\
 }
+
+%\author{
+%  Michaël Defferrard \\ %\thanks{further info} \\
+%  %Department of Electrical Engineering \\
+%  %École Polytechnique Fédérale de Lausanne (EPFL) \\
+%  EPFL, Lausanne, Switzerland \\
+%  \texttt{michael.defferrard@epfl.ch} \\
+%  \And % \AND
+%  Xavier Bresson \\
+%  %School of Computer Science and Engineering \\
+%  Nanyang Technological University, Singapore \\
+%  \texttt{xavier.bresson@ntu.edu.sg} \\
+%  \And % \AND
+%  Pierre Vandergheynst \\
+%  %Department of Electrical Engineering \\
+%  EPFL, Lausanne, Switzerland \\
+%  \texttt{pierre.vandergheynst@epfl.ch} \\
+%}
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.8\textwidth]{images/illustrationCNNgraphs}
-\caption{Architecture of the proposed CNN on graphs.
-	%Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at
-	%layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral
-	%filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the
-	%coarsening exponent, $n_c$ is the number of classes, $y$ is the output
-	%signal, and $\theta^l$ is the number of parameters to learn at $l$.
-}
-\label{fig_illustration}
-\end{figure}
-
 % Eventually, generic and efficient implementations of CNNs on graphs, as
@@
 \end{enumerate}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.87\textwidth]{images/illustrationCNNgraphs}
+\caption{Architecture of the proposed CNN on graphs.
+	%Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at
+	%layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral
+	%filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the
+	%coarsening exponent, $n_c$ is the number of classes, $y$ is the output
+	%signal, and $\theta^l$ is the number of parameters to learn at $l$.
+}
+\label{fig_illustration}
+\end{figure}
 
@@
 matrix of scaled eigenvalues that lie in $[-1,1]$.
-
 The filtering operation can then be written as $y = g_\theta(L) x
@@
 $\bar{x}_1 = \tilde{L}x$. The entire filtering operation
-\begin{equation}
-	y = g_\theta(L) x = [\bar{x}_0, \ldots, \bar{x}_{K-1}] \theta
-\end{equation}
+	$y = g_\theta(L) x = [\bar{x}_0, \ldots, \bar{x}_{K-1}] \theta$
 then costs $\bO(K|\E|)$ operations.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{images/training}
+\includegraphics[width=\textwidth]{images/training}
 \caption{Plots of validation accuracy and training loss for the first 2000

### 2016-10-31 nips2016/nips2016.tex
makefile: image creation
arxiv v2 (2016-10-31)

@@
 \documentclass{article}
+\pdfoutput=1  % for arXiv to use pdflatex
 
@@
 % \PassOptionsToPackage{numbers, compress}{natbib}
+\PassOptionsToPackage{numbers}{natbib}
 % before loading nips_2016
@@
 represented, to high-dimensional irregular domains, such as social networks,
-brain connectomes or words’ embedding, represented by graphs.
+brain connectomes or words' embedding, represented by graphs.
 We present a formulation of CNNs in the context of spectral graph theory, which
@@
 \centering
-\includegraphics[width=\textwidth]{images/coarsening_crop}
+\includegraphics[width=\textwidth]{images/coarsening}
 \caption{\textbf{Example of Graph Coarsening and Pooling.} Let us carry out a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
@@
 	\centering
-	\includegraphics[width=\textwidth]{images/runtime_crop}
+	\includegraphics[width=\textwidth]{images/runtime}
 	\captionof{figure}{Time to process a mini-batch of $S=100$ 20NEWS documents

### 2017-02-05 nips2016/nips2016.tex
rename folder: images -> figures
nips2016 final paper revision

@@
 \usepackage{graphicx}
+\graphicspath{{figures/}}
 %\usepackage{tabularx}
@@
 \usepackage{titlesec}		% space before / after headings
-\titlespacing\section{0pt}{5pt plus 1pt minus 2pt}{1.8pt plus 1pt minus 1pt}
-\titlespacing\subsection{0pt}{5pt plus 1pt minus 2pt}{2.7pt plus 1pt minus 2pt}
+\titlespacing\section{0pt}{3.3pt plus 1pt minus 2pt}{1.8pt plus 1pt minus 1pt}
+\titlespacing\subsection{0pt}{3.3pt plus 1pt minus 2pt}{2.7pt plus 1pt minus 2pt}
 %\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
@@
 
-\usepackage{parskip} \setlength{\parskip}{2.1pt} \setlength{\parindent}{1em}
+\usepackage{parskip} \setlength{\parskip}{3.9pt} %\setlength{\parindent}{1em}
 
@@
 \author{
-	Michaël Defferrard\affe \hspace{1cm} Xavier Bresson\affs \hspace{1cm} Pierre Vandergheynst\affe %\thanks{further info} \\
+	Michaël Defferrard \hspace{1cm} Xavier Bresson \hspace{1cm} Pierre Vandergheynst %\thanks{further info} \\
 	%Department of Electrical Engineering \\
@@
 	\vspace{0.2cm} \\
-	\affe EPFL, Lausanne, Switzerland \hspace{1cm} \affs Nanyang Technological University, Singapore \\
-	\texttt{\{michael.defferrard,pierre.vandergheynst\}@epfl.ch \hfill xavier.bresson@ntu.edu.sg} \\
+	%\affe EPFL, Lausanne, Switzerland \hspace{1cm} \affs Nanyang Technological University, Singapore \\
+	EPFL, Lausanne, Switzerland \\
+	\texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} \\
 }
@@
 data on telecommunication networks, or text documents on word embeddings are
-important examples of data lying on irregular or non-Euclidean domains which
+important examples of data lying on irregular or non-Euclidean domains that
 can be structured with graphs, which are universal representations of
@@
 \item \textbf{Spectral formulation.} A spectral graph theoretical formulation
-	of CNNs on graphs built on established tools in GSP
+	of CNNs on graphs built on established tools in graph signal processing (GSP).
 	\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
@@
 	have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor
-	(NN) graphs, leading to a linear complexity w.r.t the input data size.
+	(NN) graphs, leading to a linear complexity w.r.t the input data size $n$.
 	Moreover, this method avoids the Fourier basis altogether, thus the
@@
 	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. We also show that our graph
-	formulation performs similarly as a classical CNNs on MNIST and study the
-	impact of various graph constructions on classification performance. The
+	formulation performs similarly to a classical CNNs on MNIST and study the
+	impact of various graph constructions on performance. The
 	TensorFlow \cite{abadi_tensorflow_2016} code to reproduce our results and
@@
 	\centering
-\includegraphics[width=0.87\textwidth]{images/illustrationCNNgraphs}
-\caption{Architecture of the proposed CNN on graphs.
-	%Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at
-	%layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral
-	%filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the
-	%coarsening exponent, $n_c$ is the number of classes, $y$ is the output
-	%signal, and $\theta^l$ is the number of parameters to learn at $l$.
-}
-\label{fig_illustration}
+	\includegraphics[width=\textwidth]{architecture}
+	\caption{Architecture of a CNN on graphs and the four ingredients of a (graph) convolutional layer.}
+	\label{fig:architecture}
 \end{figure}
@@
 provide filter localization via the finite size of the kernel. However,
-although graph convolution directly in the spatial domain is conceivable, it
-also faces the challenge of matching local neighborhoods, as pointed out in
+although graph convolution in the spatial domain is conceivable, it
+faces the challenge of matching local neighborhoods, as pointed out in
 \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Consequently, there is no unique
 mathematical definition of translation on graphs from a spatial perspective. On
-the other side, a spectral approach provides a well-defined translation
+the other side, a spectral approach provides a well-defined localization
 operator on graphs via convolutions with a Kronecker delta implemented in the
@@
 defined on undirected and connected graphs $\G=(\V,\E,W)$, where $\V$ is a
-finite set of vertices with $|\V|=n$, $\E$ is a set of edges and $W$ is a
+finite set of $|\V|=n$ vertices, $\E$ is a set of edges and $W \in \R^{n \times n}$ is a
 weighted adjacency matrix encoding the connection weight between two vertices. A
@@
 \in \R^{n \times n}$ such that $L = U \Lambda U^T$ where $\Lambda =
-\diag([\lambda_0, \ldots, \lambda_{n-1}]) \in \R^{n \times n}$. We can now
-define the graph Fourier transform (GFT) of a spatial signal $x \in \R^n$ as
+\diag([\lambda_0, \ldots, \lambda_{n-1}]) \in \R^{n \times n}$.
+The graph Fourier transform of a signal $x \in \R^n$ is then defined as
 $\hat{x} = U^T x \in \R^n$, and its inverse as $x = U \hat{x}$
-\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. Similarly to the
-Fourier transform on Euclidean spaces, the GFT enables the formulation of
-fundamental operations such as filtering.
+\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. As on Euclidean
+spaces, that transform enables the formulation of fundamental
+operations such as filtering.
 
@@
 given by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
-(L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
+(L^k)_{i,j}$, where the kernel is localized via a convolution with a Kronecker
 delta function $\delta_i \in \R^n$. By \cite[Lemma
@@
 
-Recall that the Chebyshev polynomial $T_k(x)$ of order $k$ may be generated by
+Recall that the Chebyshev polynomial $T_k(x)$ of order $k$ may be computed by
 the stable recurrence relation $T_k(x) = 2x T_{k-1}(x) - T_{k-2}(x)$ with $T_0 =
@@
 / \sqrt{1-y^2})$, the Hilbert space of square integrable functions with respect
-to the measure $dy/\sqrt{1-y^2}$. A filter can thus be parametrized as a
-truncated expansion of order $K-1$ such that
+to the measure $dy/\sqrt{1-y^2}$. A filter can thus be parametrized as the
+truncated expansion
 \begin{equation} \label{eq:filt_cheby}
@@
 \end{equation}
+of order $K-1$,
 where the parameter $\theta \in \R^K$ is a vector of Chebyshev coefficients and
@@
 coarser versions of a given graph and is able to minimize several popular
-spectral clustering objectives. We chose the normalized cut
-\cite{art:ShiMalik00NCut}, which is an excellent clustering energy. Graclus'
-greedy rule consists, at any given coarsening level, in picking an unmarked
+spectral clustering objectives, from which we chose the normalized cut
+\cite{art:ShiMalik00NCut}. Graclus'
+greedy rule consists, at each coarsening level, in picking an unmarked
 vertex $i$ and matching it with one of its unmarked neighbors $j$ that
@@
 weights. The matching is repeated until all nodes have been explored. This is
-an extremely fast coarsening scheme which divides the number of nodes by
+an very fast coarsening scheme which divides the number of nodes by
 approximately two (there may exist a few singletons, non-matched nodes) from
@@
 \centering
-\includegraphics[width=\textwidth]{images/coarsening}
+\includegraphics[width=\textwidth]{coarsening}
 \caption{\textbf{Example of Graph Coarsening and Pooling.} Let us carry out a max pooling of size 4 (or two poolings of size 2) on a signal $x \in
@@
 possesses $n_0 = |\V_0| = 8$ vertices, arbitrarily ordered. For a pooling of
-size 4, two coarsenings of size 2 are needed: Graclus first gives $\G_1$ of size
+size 4, two coarsenings of size 2 are needed: let Graclus gives $\G_1$ of size
 $n_1 = |\V_1| = 5$, then $\G_2$ of size $n_2 = |\V_2| = 3$, the coarsest graph.
@@
 in a memory inefficient, slow, and hardly parallelizable implementation. It is
-however possible to structure the arrangement of the vertices such that a graph
+however possible to arrange the vertices such that a graph
 pooling operation becomes as efficient as a 1D pooling. We proceed in two
 steps: (i) create a balanced binary tree and (ii) rearrange the vertices.
-After coarsening, each node has either two parents, if it was matched at the
+After coarsening, each node has either two children, if it was matched at the
 finer level, or one, if it was not, i.e the node was a singleton. From the
@@
 regular nodes (e.g. level 1 vertex 0 in \figref{pooling}) or (ii) one singleton
-and a fake node as children (e.g. level 2 vertex 0 in \figref{pooling}), and
-(iii) fake nodes always have two fake nodes as children (e.g. level 1 vertex 1
-in \figref{pooling}). Input signals are initialized with a neutral value at the
+and a fake node as children (e.g. level 2 vertex 0), and
+(iii) fake nodes always have two fake nodes as children (e.g. level 1 vertex 1).
+Input signals are initialized with a neutral value at the
 fake nodes, e.g. 0 when using a ReLU activation with max pooling. Because these
@@
 rearranged graph signal is analog to pooling a regular 1D signal.
-\figref{pooling} shows an example of the whole coarsening and pooling process.
-This regular arrangement makes the pooling operation very efficient and
+\figref{pooling} shows an example of the whole process.
+This regular arrangement makes the operation very efficient and
 satisfies parallel architectures such as GPUs as memory accesses are local,
@@
 The authors of \cite{art:BrunaZarembaSzlamLeCun13DLgraphs} used this idea for
-their spatial formulation of graph CNNs. They use the weighted graph to define
+their spatial formulation of graph CNNs. They use a weighted graph to define
 the local neighborhood and compute a multiscale clustering of the graph for the
@@
 low-dimensional non-Euclidean spaces, was proposed in \cite{masci2015geodesic}.
-The authors used geodesic polar coordinates to define convolution operations on
+The authors used geodesic polar coordinates to define the convolution on
 mesh patches, and formulated a deep learning architecture which allows
 comparison across different manifolds. They obtained state-of-the-art results
-for 3D shape recognition tasks.
+for 3D shape recognition.
 
@@
 
-The first spectral formulation of graph CNN, proposed in
+The first spectral formulation of a graph CNN, proposed in
 \cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, defines a filter as
@@
 computational bottleneck already identified by the authors. Besides, as they
-rely on smoothness, via the spline parametrization, in the Fourier domain to
+rely on smoothness in the Fourier domain, via the spline parametrization, to
 bring localization in the vertex domain, their model does not provide a precise
@@
 dataset of 70,000 digits represented on a 2D grid of size $28 \times 28$. For
-our graph model, we construct a $8$-NN graph of the 2D grid which produces a
+our graph model, we construct an $8$-NN graph of the 2D grid which produces a
 graph of $n = |\V| = 976$ nodes ($28^2 = 784$ pixels and 192 fake nodes as
@@
 \begin{equation} \label{eq:knngraph}
-	W_{ij} = \exp \left( -\frac{\|z_i - z_j\|_2^2}{\sigma^2} \right)
+	W_{ij} = \exp \left( -\frac{\|z_i - z_j\|_2^2}{\sigma^2} \right),
 \end{equation}
@@
 probability of 0.5, regularization weight of $5\times10^{-4}$, initial learning
-rate of 0.03, learning rate decay of 0.95, momentum of 0.9. Convolutional layers
-have filters of size $5 \times 5$ and graph convolutional layers have the same
-support of $K = 5^2 = 25$. All models were trained for 20 epochs.
+rate of 0.03, learning rate decay of 0.95, momentum of 0.9. Filters are of size
+$5 \times 5$ and graph filters have the same support of $K = 25$. All models
+were trained for 20 epochs.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{images/runtime}
+	\includegraphics[width=\textwidth]{runtime}
 	\captionof{figure}{Time to process a mini-batch of $S=100$ 20NEWS documents
@@
 \centering
-\includegraphics[width=\textwidth]{images/training}
+\includegraphics[width=\textwidth]{training}
 \caption{Plots of validation accuracy and training loss for the first 2000
@@
 for text documents represented as bag-of-words. We investigate here three ways
-to represent a word $z$ when constructing a $16$-NN similarity graph with
-\eqnref{knngraph}: while the simplest option is to represent each word as its
-corresponding column in the bag-of-words matrix, another approach is to learn
-an embedding for each word with word2vec
+to represent a word $z$: the simplest option is to represent each word as its
+corresponding column in the bag-of-words matrix while, another approach is to
+learn an embedding for each word with word2vec
 \cite{pro:MikolovChenCorradoDean13word2vec} or to use the pre-learned
@@
 \cite{pro:BawaCondieGanesan05LSHForest} on the learned word2vec embeddings.
-\tabref{20news_quality} reports the classification results which highlight the
+\tabref{20news_quality} reports classification results which highlight the
 importance of a well constructed graph.
@@
 In this paper, we have introduced the mathematical and computational
-foundations of an efficient generalization of CNNs on graphs using tools from
+foundations of an efficient generalization of CNNs to graphs using tools from
 GSP. Experiments have shown the ability of the model to extract local and
@@
 approach, pioneered in \cite{art:HenaffBrunaLeCun15DLgraphs}, would be to
-alternate the learning of the CNN parameters and the graph, as a virtuous
-circle.
+alternate the learning of the CNN parameters and the graph.
 
@@
 
+%TODO \subsubsection{*\todo{Acknowledgment}}
+% Funding
+% NVIDIA for GPUs
+% Bruna & Scarceli for connection with GNN
+% NIPS anonymous reviewers
+
 \clearpage
@@
 {
+	\setlength{\parskip}{2.1pt}
 	\setstretch{0}

### 2019-10-04 review/rebuttal.txt
move paper to root

--- 2016-10-29 nips2016/review/rebuttal.txt
@@
-############
-ALL
-############
+Private to AC
+-------------
 
-4992
+The general consensus of the referees lean towards acceptance.
+Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful.
+The only concerns come from Referees 8,9 who acknowledge their low confidence in their understanding of the paper main contributions.
+
+Rebuttal to all referees
+------------------------
 
@@
 Referees 1,3,7
-While the isotropic nature of our filters (see answer to R1) and the lack of experience on architecture design are certainly explanations for the gap between classical CNNs and graph CNNs on MNIST, we observed that with 20 epochs (instead of 10 in our submission) for the same LeNet5 architectures, classical CNNs obtained 99.33% accuracy and graph CNNs 99.14%.
+While the isotropic nature of our filters (see answer to R1) and the lack of experience on architecture design are certainly explanations for the gap between classical CNNs and graph CNNs on MNIST, we observed that with 20 epochs (instead of 10 in our submission) for the same LeNet5 architectures, classical CNNs obtained an accuracy of 99.33% and graph CNNs obtained 99.14%.
 This reveals a slower convergence of graph CNNs, and the need to investigate better suited optimization or initialization strategies.
@@
 Referees 3,7,8
-The goal of Tables 3 and 5 is to show that the usefulness of the graph structure directly depends on the graph quality.
-Random graphs are used as a baseline to assess graph quality, as convolutional layers are almost useless in this setting.
+The goal of Tables 3,5 is to show that the usefulness of the graph structure directly depends on the graph quality.
+Random graphs are used as a lower bound of graph quality, as convolutional layers are almost useless in this setting.
 A well suited graph, like the grid for MNIST, is useful and allows our model to perform well.
 We did however not try to optimize the graph structure for 20NEWS: there may be better ways to construct higher-quality graphs.
-It was not our objective, which was to show how graph quality impacts performance.
-The goal of Table 4 is to demonstrate that a convolutional layer is also effective on graphs that are not regular grids, by comparing it to fully connected networks.
+Our only objective was to show how graph quality impacts performance.
+The goal of Table 4 is to demonstrate that a convolutional layer (CN32) is also effective on graphs that are not regular grids, by comparing it to fully connected networks (FC).
 We also compared against the effective Multinomial Naive Bayes and two baselines.
-Graph CNNs did not overcome Multinomial NB, likely because of the small size of the 20NEWS dataset (~19k documents).
+Graph CNNs did not overcome Multinomial NB, likely because of the small size of the 20NEWS dataset (19k documents).
 We plan to apply our technique to the larger RCV1 dataset (800k documents).
@@
 The spectral filters are indeed rotationally invariant.
-Whether this is a limitation or an advantage depends on the problem and should be verified empirically, as for any other invariance.
+Whether this is a limitation or an advantage depends on the problem and should be verified, as for any invariance.
 Our claim is motivated by the following experiment (not reported in the paper): on a rotated MNIST dataset, i.e.
@@
 Many data augmentation schemes have used rotated versions of (part of) images to force a network to learn this invariance.
-Spatial Transformer Networks have been developed to learn invariance to translation, scale, rotation and more generic warping.
+Models, like Spatial Transformer Networks, have been developed to learn this invariance.
 In a future work, we will test our model on large-scale image datasets and investigate if such an invariance is an advantage.
-We will also study the possibility to learn anisotropic filters with the help of anisotropic Laplacians.
+We will also study the possibility to learn anisotropic filters with anisotropic Laplacians.
 
 Referee 7
-We indeed consider a single graph in this work, which is already meaningful for multiple applications such as brain connectivity analysis, documents network analysis and social network analysis.
+We indeed consider a single graph in this work, which is already meaningful for multiple applications such as the analysis of brain connectivity, documents network and social network.
 Transferring graph features/filters from one graph to another is obviously of great interest, and can also be explored with spectral techniques like the one proposed here (e.g.
@@
 Referee 8
-We did recently improve the presentation of the paper. Expect the final version .
+We recently improved the paper presentation and will release it in the final version, along with the code used to produce our results.
 The performance analysis of classical CNNs is a central and open mathematical question.
 It is thus challenging to state anything about the theoretical performances of our model.
-The only performance study is experimental and summarized in Table 6, which shows that our model is indeed superior to [13].
+The only performance study is experimental and summarized in Table 6, which shows that our model compares favorably to [13].
 
 Referee 9
-Presentation: see our answer to R8.
+Presentation: see answer to R8.
 We did not introduce any natural graph experiments in this work.
@@
 Constructed graphs are a great way of demonstrating this.
-However, a data analysis project on the network of Wikipedia articles using this technique is ongoing.
-
-############
-Private to AC
-############
-
-The general consensus of the referees lean towards acceptance.
-Referees 1,3,6,7 agree that the model is novel, effective and potentially impactful.
-The only concerns come from Referees 8,9 who acknowledge their low confidence in their understanding of the paper main contributions.
-
-Please note that, to foster innovation, we are going to publish the code used to produce our results once the paper is published.
+However, a data analysis project on the (natural) network of Wikipedia articles using this technique is ongoing.
